---
title: "C14 hot incubation rates calculation"
author: "Denise Ong"
date: "4/2/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
  knitr::opts_chunk$set(message=FALSE,
                        warning=FALSE,
                        cache=TRUE,
                        tidy=FALSE)
```

```{r}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(broom)
```

# Denise edits

The following edits made to pico and syn data:
•	Exp 3: SUR: pico: vial A – remove. Decreasing LM value.
• Exp 2: DCM: pico vial A - only 1 value, cells sorted 1120. DPM is lower than the average dark vial. Remove
•	Add Exp 8: SUR: nano – for vial B and D, there is cells sorted for both 1000 and 2000. Move to syn and pico calculations for lm value
•	Exp 4: SUR: pico, syn – repeated vial B, take the average of the dpm value. For this experiment, vial D was lost. By mistake Domi started to sort again vial B so continued.

The following edits made to Nano data:
•	Exp 6: SUR: nano – repeated vial A, remove lower value. 20221019- 1307 vs 384. 384 value is very low.
•	Exp 7: SUR: nano – remove Dark incubation is higher than B. The DPM of B is lower than average D values. Remove.
•	Exp 7: DCM: nano – remove (Dark incubation is much higher than B, C). 20221019- has 1000 and 2000 cells sorted. Vial D is twice the value of vial B and C, but B and C have comparable values. For this will change, to only include 1000 cells sorted for B and C, and use average vial D values.
•	Exp 8: DCM: nano – remove. DPM range from 31-53, which is the same range for vial D.
•	Exp 1: SUR: nano – value for dpm for vial C and D swap
•	Exp 4: SUR: nano – repeated vial B, take the average of the dpm value. For this experiment, vial D was lost. By mistake Domi started to sort again vial B so continued.
• Added Exp 2: DCM: pico vial A- only 1 value, cells sorted 1120.
•	Exp 8: SUR: nano – for vial B and D, there is cells sorted for both 1000 and 2000. Move to syn and pico calculations. for lm value. vial C also has cells sorted for both 1000 and 2000 cells, but 2000 cells dpm value is too low. calculate for vials A,B,C for 1000 cells also with the other nano.
•	Exp 6: SUR: vial C, error in recording. Changed in excel sheet so excel sheet name used changed from "TAN1810_14C-Pico_25052020_QA.xlsx" to "TAN1810_14C-Pico_25052020_QA_1.0.xlsx"
•	exp 5: SUR. Vial D is high. Also marked in the excel sheet. Removed.


The calculations of primary production are based on single cell C14 uptake for cells sorted by flow cytometry. (https://vaulot.netlify.app/2018/05/20/compute-primary-production-based-on-single-cell-c14-uptake/)

To calculate the pp values, pico and syn had cell counts of 2000, 4000, 10000. The gradient of the lm is calculated by grouping vials A, B, C together to get the DPM per cell for each experiment. For nano, the cells were sorted mostly with one cell count of 1000. The DPM per cell is calculated by the average of vial A, B, C to get the value for each experiment. 

# 1. Calculations for pico and syn - Compute linear model. 

The following edits made to pico and syn data:
•	Exp 3: SUR: pico: vial A – remove. Decreasing LM value.
• Exp 2: DCM: pico vial A- only 1 value, cells sorted 1120. Moved to part 2 (nano calculations) for average value instead of lm value.
•	Add Exp 8: SUR: nano – for vial B and D, there is cells sorted for both 1000 and 2000. Move to syn and pico calculations for lm value.
•	Exp 4: SUR: pico, syn – repeated vial B, take the average of the dpm value

## Read table for pico and syn. Data points to edit are made here. 
```{r}
C14_DPM <- readxl::read_excel(here("0_data_used", "TAN1810_14C-Pico_25052020_QA_1.2.xlsx"), sheet = "Complete data TAN1810", skip = 11) %>%
  dplyr::select (`Cycle#`, `EXP#`, SAMPLE, `Vial code             (i - initials, A to C - light incubation, D - dark)`, `sorting population`, `Cells sorted`, DPM1, DIC, `SA_dpm_mL`) %>%   
     rename(cycle = `Cycle#`,
            exp = `EXP#`,
            sample = SAMPLE, 
            vial = `Vial code             (i - initials, A to C - light incubation, D - dark)`, 
            population = `sorting population`, 
            cells_sorted = `Cells sorted`, 
            dpm = DPM1,
            SA = `SA_dpm_mL`) %>%
    filter(population != "Nano" | (population == "Nano" & exp == 8 & sample == "SUR" & vial == "B") | (population == "Nano" & exp == 8 & sample == "SUR" & vial == "D")) %>% # cells sorted for both 1000 and 2000
    filter(!(exp == 3 & sample == "SUR" & vial == "A" & population == "Pico")) %>% # Decreasing LM value.
    filter(!(exp == 2 & sample == "DCM" & vial == "A" & population == "Pico")) %>% # DPM is lower than the average dark vial. Remove
     group_by(cycle, exp, sample, vial, population, cells_sorted) %>% 
     summarize(cells_sorted = mean(cells_sorted),
               dpm= mean(dpm),
               DIC = mean(DIC),
               SA = mean(SA)) %>%
    ungroup()

C14_DPM
```

## Subtract the dark DPM
Join tables based on common variables

```{r}
C14_DPM_Dark <- C14_DPM %>% 
  filter(vial == "D") %>% 
  rename(dpm_dark = dpm) %>% 
  dplyr::select(-vial) # Remove the vial column
C14_DPM_Dark
# Check the values of dpm dark with histogram. The values are normally distributed, concentrated around the mean.
hist(C14_DPM_Dark$dpm_dark)
mean(C14_DPM_Dark$dpm_dark)

#for checking. Not using.
#C14_dpm_dark_1000cells<- C14_DPM_Dark %>%
#  filter(cells_sorted == "2000") %>%
#  summarise(dpm_dark = mean(dpm_dark))
```

```{r}
C14_DPM_corrected <- left_join(C14_DPM,C14_DPM_Dark) %>% 
  filter (vial != "D") %>% 
  # DV - The next line replace dpm_dark to average dpm_dark if no value. Checked with histogram above.
  mutate(dpm_dark = case_when (is.na(dpm_dark) ~ mean(dpm_dark, na.rm = TRUE),
                               TRUE ~ dpm_dark))  %>% 
  # DV - The next line set dpm_correct to zero if negative, as negative values are all close to zero.
  mutate(dpm_corrected = case_when (dpm >= dpm_dark ~ dpm - dpm_dark,
                                    TRUE ~0)) 

C14_DPM_corrected  
```



## Method 2 - Compute lm for  ABC separately

### Plots
Do plots for each group. One regression line based on the EXP, Sample and Populations

```{r fig.height=15, fig.width=10}
 reg_plot_2<- ggplot(data = C14_DPM_corrected, aes(x=cells_sorted, y=dpm_corrected, color=population, shape=vial)) +
  geom_point() + stat_smooth(method="lm", se=FALSE) +
  # DV - Set the y_scale to free so that graph is more easy to read
  facet_grid(rows=vars(exp), cols=vars(sample), scales = "free_y") +
  scale_color_viridis_d()
# pdf("figures/C14_regression_method 2.pdf", height=15, width=10) ; plot(reg_plot_2) ; dev.off()
reg_plot_2
```



### Do linear model

y = ax + b

See: https://cran.r-project.org/web/packages/broom/vignettes/broom_and_dplyr.html

```{r}
 C14_DPM_model_2 <- C14_DPM_corrected %>%
  group_by(cycle, exp, sample, population, vial, DIC, SA) %>% 
  tidyr::nest() %>% 
  mutate(
    fit = purrr::map(data, ~ lm(dpm_corrected ~ cells_sorted, data = .x)),
    tidied = purrr::map(fit, tidy)
  ) %>% 
  unnest(tidied)

C14_DPM_model_output_2 <- C14_DPM_model_2 %>% 
  dplyr::select(exp:population, term, estimate) %>% 
  pivot_wider(names_from="term", values_from ="estimate" ) %>% 
  rename (slope = cells_sorted, intercept = `(Intercept)`)

C14_DPM_model_output_2
  
```


### Calculate PP value, based on Daniel's formula found here:  
https://vaulot.netlify.com/2018/05/20/compute-primary-production-based-on-single-cell-c14-uptake/

```{r}
pp_cal_2 <- mutate(C14_DPM_model_output_2, pp = DIC*slope*(1/(SA*24))*10^9*1.05) %>%
  ungroup() %>%
  filter(pp > 0) %>% # remove negative pp values 
  filter(population != "Nano") #remove the nano calculation to keep consistent.
  
pp_cal_2
```


# 2. Calculations for Nano (average)
The following edits made to Nano data:
•	Exp 6: SUR: nano – repeated vial A, remove lower value. 20221019- 1307 vs 384. 384 value is very low.
•	Exp 7: SUR: nano – remove Dark incubation is higher than B. The DPM of B is lower than average D values. Remove.
•	Exp 7: DCM: nano – Dark incubation is much higher than B, C. 20221019- has 1000 and 2000 cells sorted. Vial D is twice the value of vial B and C, but B and C have comparable values. For this will change, to only include 1000 cells sorted for B and C, and use average vial D values.
•	Exp 8: DCM: nano – remove. DPM range from 31-53, which is the same range for vial D.
•	Exp 1: SUR: nano – value for dpm for vial C and D swap
•	Exp 4: SUR: nano – repeated vial B, take the average of the dpm value
• Added Exp 2: DCM: pico vial A- only 1 value, cells sorted 1120.
•	Exp 8: SUR: nano – for vial B and D, there is cells sorted for both 1000 and 2000. Move to syn and pico calculations. for lm value. vial C also has cells sorted for both 1000 and 2000 cells, but 2000 cells dpm value is too low. calculate for vials A,B,C for 1000 cells also with the other nano. 
•	exp 5: SUR. Vial D is high. Also marked in the excel sheet. Removed.

## Read data
```{r}
# Changed the excel sheet read, with final dpm and DIC values
C14_DPM_nano <- readxl::read_excel(here("0_data_used", "TAN1810_14C-Pico_25052020_QA_1.2.xlsx"), sheet = "Complete data TAN1810", skip = 11)%>% 
  dplyr::select (`Cycle#`, `EXP#`, SAMPLE, `Vial code             (i - initials, A to C - light incubation, D - dark)`, `sorting population`, `Cells sorted`, DPM1, DIC, `SA_dpm_mL`) %>%   
     rename(cycle = `Cycle#`,
            exp = `EXP#`,
            sample = SAMPLE, 
            vial = `Vial code             (i - initials, A to C - light incubation, D - dark)`, 
            population = `sorting population`, 
            cells_sorted = `Cells sorted`, 
            dpm = DPM1,
            SA = `SA_dpm_mL`) %>%
    filter(population == "Nano") %>%
    filter(!(exp == 5 & sample == "SUR" & vial =="D")) %>%
    filter(!(exp == 6 & sample == "SUR" & vial == "A" & dpm == 384)) %>%
    filter(!(exp == 7 & sample == "SUR")) %>%
    filter(!(exp == 7 & sample == "DCM" & vial =="D")) %>%
    filter(!(exp == 7 & sample == "DCM" & cells_sorted == 2000)) %>%
    filter(!(exp == 8 & sample == "DCM")) %>%
    filter(!(exp == 8 & sample == "SUR" & cells_sorted == 2000)) %>%
    mutate(vial = case_when((exp == 1 & sample == "SUR" & vial == "C") ~ "D",
                            (exp == 1 & sample == "SUR" & vial == "D") ~ "C",
                            TRUE ~ vial)) %>%  #for swapping values of C and D
   group_by(cycle, exp, sample, vial, population, cells_sorted) %>% 
   summarize(cells_sorted = mean(cells_sorted),
             dpm= mean(dpm),
             DIC = mean(DIC),
             SA = mean(SA)) %>%
    ungroup()

C14_DPM_nano
```

## Substract the dark DPM
Join tables based on common variables

```{r}
 C14_DPM_Dark_nano <- C14_DPM_nano %>% 
  filter(vial == "D") %>% 
  rename(dpm_dark = dpm) %>% 
  dplyr::select(-vial)  # Remove the vial column
C14_DPM_Dark_nano
# Check the values of dpm dark with histogram.
hist(C14_DPM_Dark_nano$dpm_dark)
mean(C14_DPM_Dark_nano$dpm_dark)
```
```{r}
C14_DPM_corrected_nano <- left_join(C14_DPM_nano,C14_DPM_Dark_nano) %>% 
  filter (vial != "D") %>% 
  # DV - The next line replace dpm_dark to average dpm_dark if no value. Checked with histogram above.
  mutate(dpm_dark = case_when (is.na(dpm_dark) ~ mean(dpm_dark, na.rm = TRUE),
                               TRUE ~ dpm_dark))  %>% 
  # DV - The next line set dpm_correct to zero if negative, as negative values are all close to zero.
  mutate(dpm_corrected = case_when (dpm >= dpm_dark ~ dpm - dpm_dark,
                                    TRUE ~0)) 
C14_DPM_corrected_nano
```

## Method 2 - ABC vials are separate.

### Calculate PP value, based on Daniel's formula found here:  
https://vaulot.netlify.com/2018/05/20/compute-primary-production-based-on-single-cell-c14-uptake/

```{r} 
pp_cal_nano_2 <- mutate(C14_DPM_corrected_nano, pp = DIC*dpm_corrected*(1/(SA*24*cells_sorted))*10^9*1.05) %>% 
  ungroup() %>%
  filter(pp > 0) # remove negative pp values

# add missing rows for data analysis later
#pp_cal_nano_2 <- complete(pp_cal_nano_2, nesting(cycle, exp, sample), vial, population) %>%
#  ungroup()
  
pp_cal_nano_2
```


# 3. Join pico, syn and nano values together

## Method 2

```{r}
pp_cal_2_join<-full_join(pp_cal_2,pp_cal_nano_2) %>%
  select(cycle, exp:population, pp) %>%
  arrange(cycle, exp, sample, population) 
pp_cal_2_join
#pp_cal_2_join$cycle <- as.character(pp_cal_2_join$cycle) 
#pp_cal_2_join$exp <- as.character(pp_cal_2_join$exp)
writexl::write_xlsx(pp_cal_2_join,here("Output", "Raw data", "C14_rates per group_method2.xlsx"))
```


Compute means quickly.

```{r}
 
pp_cal_2_join %>% 
  group_by(population, sample) %>% 
  summarise(pp_mean = mean(pp, na.rm = TRUE))

pp_cal_2_join %>% 
  ggplot() +
  geom_point(aes(x= str_c(population,sample,  sep=" "), y = pp)) +
  xlab("Sample")

pp_cal_2_join %>% 
  ggplot() +
  geom_boxplot(aes(x= str_c(population,sample,  sep=" "), y = pp)) +
  xlab("Sample")
```