---
title: "Group rates figures 4.0"
author: "Denise Ong"
date: "2023-09-26"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(here)
source(here("0_init_files", "init.R")) #for libraries
source(here("0_init_files", "init_markdown.R")) #for markdown libraries
library(FSA)
library(ggeffects)
library(gghalves)
#Colours
colours_cycle <- readxl::read_excel("0_init_files/colours_cycle.xlsx", sheet = "cycle")
colours_cycle <- structure(colours_cycle$colour,.Names=colours_cycle$cycle)

colours_exp <- readxl::read_excel("0_init_files/colours_cycle.xlsx", sheet = "exp")
colours_exp <- structure(colours_exp$colour,.Names=colours_exp$exp)

# For consistency between scatterplot
set.seed(1)
```

# Read data

```{r}
cold_standard <- readxl::read_excel("Output/Raw data/NPP_cold_standard.xlsx") %>% 
  rename(depth_m = `depth`,
         depth = `sample`)
cold_standard$exp <- as.character(cold_standard$exp)

pp_rates <- readxl::read_excel("Output/Raw data/Rates_cell_npp_method2.xlsx") %>%
  mutate(cycle_name = dplyr::recode(cycle_name, "SA1" = "SA-Sc")) %>%
  mutate(cycle_name = dplyr::recode(cycle_name, "SA2" = "SA1")) %>%
  mutate(cycle_name = dplyr::recode(cycle_name, "SA3" = "SA2"))

#set factor levels
pp_rates$water_mass <- factor(pp_rates$water_mass, levels = c("ST", "SA"))
pp_rates$cycle_name <- factor(pp_rates$cycle_name, levels = c("ST1", "ST2", "SA-Sc", "SA1", "SA2"))
pp_rates$population <- factor(pp_rates$population, levels = c("Syn", "Pico", "Nano"))
pp_rates$depth <- factor(pp_rates$depth, levels = c("SUR", "DCM"))
```

# PP

## surface and DCM - t test to check

surface is not different from DCM for all groups not significantly
different, from here on combine surface and DCM

Syn: t-test t = -1.9885, df = 40, p-value = 0.05363

Pico: t = -1.3297, df = 41, p-value = 0.191

Nano: t = -0.28573, df = 21.247, p-value = 0.7778

```{r}
library("ggpubr")

# For syn and pico, t.test var.equal=TRUE. For nano, var.equal = FALSE
i = "Syn"
# i = "Pico"
# i = "Nano"

t_test_df<- pp_rates %>%
  filter(population == i) %>%
  select(depth, water_mass, pp)

# option to subsample to account for unbalanced sampling 
# set.seed(345)
# t_test_df<- pp_rates %>%
#   filter(population == i) %>%
#   select(depth, water_mass, pp) %>%
#   group_by(depth) %>%
#   sample_n(14)

summary <- t_test_df %>%
  group_by(depth) %>%
  summarise(count = n(),
    mean = mean(pp, na.rm = TRUE),
    sd = sd(pp, na.rm = TRUE))
print(summary)

 p <- t_test_df %>%
   ggplot(aes(x=depth, y=pp))  +geom_boxplot() + geom_point() +
    facet_grid(~depth, scales = "free") +
    scale_y_continuous(trans='log10') + annotation_logticks(sides = "l")  + ggtitle(i)
 print(p)

# checks, all groups meet assumptions of normality and homoskedasticity
hist(log10(t_test_df$pp))
qqnorm(log10(t_test_df$pp))
p<-with(t_test_df, shapiro.test(log10(pp[depth == "SUR"])))
print(p)
p<-with(t_test_df, shapiro.test(log10(pp[depth == "DCM"])))
print(p)

# check for variance. If variance sig different, use var.equal = FALSE
var.test(log10(pp) ~ depth, t_test_df)

res <- t.test(log10(pp) ~ depth, data =  t_test_df, var.equal = TRUE)
print(res)

# create lm and check residuals
lm <- lm(log10(pp) ~ depth, data=pp_rates)
plot(lm) 


# Calculate model estimated means and CIs
library(ggeffects)
lm_emm <- ggemmeans(lm, 
                    terms = c("depth"),
                    type = "fixed")

lm_emm
plot(lm_emm)

```

## ANOVA - population\*water mass

```{r}
# check between groups, not normally distributed.
hist(pp_rates$pp)
qqnorm(pp_rates$pp)
shapiro.test(pp_rates$pp)

# transform the data and check again. Use log transformed data.
hist(log10(pp_rates$pp))
qqnorm(log10(pp_rates$pp))
shapiro.test(log10(pp_rates$pp)) 
```

```{r, eval=FALSE}
# because of unbalanced design, subset 16 samples randomly for each group first 
# no difference when subset. Use all samples.
set.seed(74657)
# interaction plot population and water mass?
sub <- pp_rates %>%
  group_by(population, water_mass) %>%
  sample_n(16)

summary <- sub %>%
  group_by(population, water_mass) %>%
  summarize(#median = round(median(npp), 2),
            mean = round(mean(npp), 2),
            sd = round(sd(npp),2),
            n = n()) #%>%
   # unite("mean_sd", mean:sd, sep = " ± ") 
# print(summary)

lm1 <- lm(log10(pp) ~ population*water_mass, data=sub)
lm2 <- lm(log10(pp) ~ population+water_mass, data=sub)
# lm3 <- lm(log10(pp) ~ population, data=sub)
print(anova(lm1, lm2))
# plot(lm1)
# print(Anova(lm1, type = "3")) # p< 0.0001
# print(summary(lm1))

lm_emm <- ggemmeans(lm1,
                    terms = c("population", "water_mass"),
                    type = "fixed")

# print(lm_emm)
print(plot(lm_emm) + scale_y_continuous(trans='log10') +
      annotation_logticks(sides = "l", outside = TRUE))

#post-hoc test
# library(emmeans)
lm.emm <- emmeans(lm1, ~ population*water_mass)
print(contrast(lm.emm, "consec", simple = "each", combine = TRUE, adjust = "mvt"))


```

## test and plot

```{r}

lm1 <- lm(log10(pp) ~ population*water_mass, data=pp_rates)
# lm2 <- lm(log10(pp) ~ population+water_mass, data=pp_rates)
# lm3 <- lm(log10(pp) ~ population, data=pp_rates)
# anova(lm1, lm2) # significant difference between lm1 and lm2. Use lm1.
plot(lm1)
Anova(lm1, type = "3") # p< 0.0001
summary(lm1)

#post-hoc test
library(emmeans)
lm.emm <- emmeans(lm1, ~ population*water_mass)
emm_df <- as.data.frame(contrast(lm.emm, "consec", simple = "each", combine = TRUE, adjust = "mvt", type = "response"))

lm_emm <- ggemmeans(lm1,
                    terms = c("population", "water_mass"),
                    type = "fixed",
                    # back.transform = "FALSE"
                    )

lm_emm

emm.pp.df <- as.data.frame(lm_emm)

plot(lm_emm) + scale_y_continuous(trans='log10') +
      annotation_logticks(sides = "l", outside = TRUE)

pp_mixed <- ggplot() +
        # Raw data
        geom_half_violin(aes(x = population,
                            y = pp, split = water_mass, fill=water_mass),
                        nudge = 0.001,
                        alpha=0.5,
                        data = pp_rates) +
        # Line of model predicted means
        geom_pointrange(aes(x = x,
                            y = predicted,
                            ymin = conf.low,
                            ymax = conf.high,
                            colour = group),
                        size = 0.6,
                        lwd = 1.3, 
                        position = position_dodge(width = 0.25),
                        data = lm_emm) +
        labs(x = "Population",
             y = "Cell-specific carbon fixation rate (fgC cell-1 hr-1)",
             colour = "Water mass",
             shape = "Cycle") +
      scale_shape_manual(values=c(0,1, 15, 17, 19)) +
      scale_fill_manual(values = c("#C62828", "#1565C0"))+ 
      scale_colour_manual(values = c("#C62828", "#1565C0"))+
      scale_y_continuous(trans='log10') +
      annotation_logticks(sides = "l") +
  annotate("text", x = 1.5, y = 100, label = "ANOVA, R2=0.83,\n Population X Water mass: F(2,123) = 5.91,\n p = 0.004", size= 3)
      # annotate("text", x = 1.7, y = 10, label = "t-test; R2=0.48,\n t40=6.22, p<0.0001")
pp_mixed
```

# NPP

## surface and DCM - t test

not using wilcoxon test output, but all not significant.

syn: t-test t = -1.5679, df = 40, p-value = 0.1248 wilcoxon test W =
136, p-value = 0.1124

pico t = -1.4587, df = 41, p-value = 0.1523 W = 148, p-value = 0.1171

nano t = 0.92363, df = 19.711, p-value = 0.3668 W = 255, p-value =
0.4568

```{r}
library("ggpubr")
# For syn and pico, t.test var.equal=TRUE. For nano, var.equal = FALSE

# i = "Syn"
# i = "Pico"
i = "Nano"

t_test_df<- pp_rates %>%
  filter(population == i) %>%
  select(depth, water_mass, npp) 

# option to subsample to account for unbalanced sampling 
# set.seed(3245)
# t_test_df<- pp_rates %>%
#   filter(population == i) %>%
#   select(depth, water_mass, npp) %>%
#   group_by(depth) %>%
#   sample_n(14)

summary <- t_test_df %>%
  group_by(depth) %>%
  summarise(count = n(),
    mean = mean(npp, na.rm = TRUE),
    sd = sd(npp, na.rm = TRUE))
print(summary)

 p <- t_test_df %>%
   ggplot(aes(x=depth, y=npp))  +geom_boxplot() + geom_point() +
    facet_grid(~depth, scales = "free") +
    scale_y_continuous(trans='log10') + annotation_logticks(sides = "l")  + ggtitle(i)
 print(p)

# checks, all groups meet assumptions of normality and homoskedasticity
hist(log10(t_test_df$npp))
qqnorm(log10(t_test_df$npp))
p<-with(t_test_df, shapiro.test(log10(npp[depth == "SUR"])))
print(p)
p<-with(t_test_df, shapiro.test(log10(npp[depth == "DCM"])))
print(p)

lm <- lm(log10(npp) ~ depth, data=t_test_df)
plot(lm) 

# check for variance. If variance sig different, use var.equal = FALSE
var.test(log10(npp) ~ depth, t_test_df)

# parametric
res <- t.test(log10(npp) ~ depth, data =  t_test_df, var.equal = FALSE)
print(res)
```

## ANOVA- population + water mass

```{r}
hist(pp_rates$npp)
qqnorm(pp_rates$npp)
shapiro.test(pp_rates$npp)

hist(log10(pp_rates$npp))
qqnorm(log10(pp_rates$npp))
shapiro.test(log10(pp_rates$npp)) # significant. Meet assumptions, using ANOVA
```

```{r, eval=FALSE}
# because of unbalanced design, subset 16 samples randomly for each group first 
# no difference when subset. Use all samples.
set.seed(76876)
  sub <- pp_rates %>%
  group_by(population, water_mass) %>%
  sample_n(16)

# interaction plot population and water mass?
lm1 <- lm(log10(npp) ~ population*water_mass, data=sub)
lm2 <- lm(log10(npp) ~ population+water_mass, data=sub) # choose this model
# lm3 <- lm(log10(npp) ~ population, data=sub)
# lm4 <- lm(log10(npp) ~ water_mass, data=sub)

anova(lm1, lm2)
# choose lm2
# plot(lm2)
Anova(lm2, type = "2") # p< 0.0001
summary(lm2)

lm_emm <- ggemmeans(lm2,
                    terms = c("population", "water_mass"),
                    type = "fixed")

lm_emm
print(plot(lm_emm) + scale_y_continuous(trans='log10') +
      annotation_logticks(sides = "l", outside = TRUE))

#post-hoc test
# library(emmeans)
lm.emm <- emmeans(lm2, ~ population+water_mass)
print(contrast(lm.emm, "consec", simple = "each", combine = TRUE, adjust = "mvt"))
```

## test and plot

```{r}
# test interaction factors
# lm1 <- lm(log10(npp) ~ population*water_mass, data=pp_rates)
lm2 <- lm(log10(npp) ~ population+water_mass, data=pp_rates) # choose this model
# lm3 <- lm(log10(npp) ~ population, data=pp_rates)
# lm4 <- lm(log10(npp) ~ water_mass, data=pp_rates)

# anova(lm1, lm2)
# choose lm2

# plot(lm2)
Anova(lm2, type = "2") # because no interaction, use type 2
summary(lm2)

#post-hoc test
# library(emmeans)
lm.emm <- emmeans(lm2, ~ population+water_mass)
contrast(lm.emm, "consec", simple = "each", combine = TRUE, adjust = "mvt")

#plot predicted means and CI
lm_emm <- ggemmeans(lm2,
                    terms = c("population", "water_mass"),
                    type = "fixed")

lm_emm

emm.npp.df <- as.data.frame(lm_emm)

plot(lm_emm) + scale_y_continuous(trans='log10') +
      annotation_logticks(sides = "l", outside = TRUE)

npp_mixed <- ggplot() +
        # Raw data
        geom_half_violin(aes(x = population,
                            y = npp, split = water_mass, fill=water_mass),
                        nudge = 0.001,
                        alpha=0.5,
                        data = pp_rates) +
        # Line of model predicted means
        geom_pointrange(aes(x = x,
                            y = predicted,
                            ymin = conf.low,
                            ymax = conf.high,
                            colour = group),
                        size = 0.6,
                        lwd = 1.3, 
                        position = position_dodge(width = 0.25),
                        data = lm_emm) +
        labs(x = "Population",
             y = "Group-specific production (mgC m-3 day-1)",
             colour = "Water mass",
             shape = "Cycle") +
      scale_shape_manual(values=c(0,1, 15, 17, 19)) +
      scale_fill_manual(values = c("#C62828", "#1565C0"))+ 
      scale_colour_manual(values = c("#C62828", "#1565C0"))+
      scale_y_continuous(trans='log10') +
      annotation_logticks(sides = "l") +
      annotate("text", x = 2.75, y = 25, label = "ANOVA, R2=0.28;\n Population: F(2,125) = 20.32, p < 0.0001; \n Water mass: F(1,125) = 12.39, p < 0.0001", size= 3)
npp_mixed

p <- (pp_mixed|npp_mixed) + plot_annotation(tag_levels = 'A') +  plot_layout(guides = 'collect')  &
  theme(legend.position='right')  & theme(axis.text = element_text(size = 12), legend.text = element_text(size = 12), legend.title = element_text(size = 12), axis.title = element_text(size = 12))
p
pdf("pp npp interactions_mean and CI.pdf", height=5, width=10) ; plot(p) ; dev.off()
```

# Table 2 - summary for paper - pp, npp, cell conc

```{r}
# General summary
summary <- pp_rates %>%
  select(population, cycle_name, pp, npp, conc_ml) %>%
  pivot_longer(pp:conc_ml, names_to = "data", values_to = "value") %>%
  group_by(data, population, cycle_name) %>%
  summarize(mean = round(mean(value), 2),
            sd = round(sd(value),2), 
             # n = n()
            ) %>%
   unite("mean_sd", mean:sd, sep = " ± ") %>%
  pivot_wider(names_from = "cycle_name", values_from = "mean_sd") %>%
  arrange(factor(data, levels = c('pp', 'npp', 'conc_ml')))
summary

# writexl::write_xlsx(summary, path = "group_summary.xlsx")

# cell conc summary - only change for SA2
summary <- pp_rates %>%
  filter(exp != "5") %>%
  select(water_mass, population, cycle_name, conc_ml) %>%
  distinct() %>%
  pivot_longer(conc_ml, names_to = "data", values_to = "value") %>%
  group_by(data, population, water_mass) %>%
  summarize(mean = round(mean(value), 0),
            sd = round(sd(value),0), 
            n = n(),
            min = round(min(value), 0),
            max = round(max(value), 0)
            ) %>%
   unite("mean_sd", mean:sd, sep = " ± ") %>%
  unite("range", min:max, sep = "-")

summary

# Small phyto NPP
summary <- pp_rates %>%
  select(population, cycle_name, npp, exp, depth) %>%
  group_by(population, cycle_name, exp, depth) %>%
  summarize(mean = round(mean(npp), 2)) %>%
  group_by(cycle_name, exp,depth) %>%
  summarise(sum = sum(mean)) %>%
  group_by(cycle_name) %>%
  summarise(mean = round(mean(sum), 2),
             sd = round(sd(sum),2)
            ) %>%
   unite("mean_sd", mean:sd, sep = " ± ") %>%
  pivot_wider(names_from = "cycle_name", values_from = "mean_sd")
summary
```

Mean of pp and npp between water mass and depth

```{r}
summary <- pp_rates %>%
  group_by(population, water_mass) %>%
  summarize(mean = round(mean(pp), 2),
            sd = round(sd(pp),2), n = n()) %>%
   unite("mean_sd", mean:sd, sep = "±") 
summary

summary <- pp_rates %>%
  group_by(population, water_mass) %>%
  summarize(#median = round(median(npp), 2),
            mean = round(mean(npp), 2),
            sd = round(sd(npp),2),
            n = n()) #%>%
   # unite("mean_sd", mean:sd, sep = " ± ") 
summary


summary <- pp_rates %>%
  group_by(population, depth) %>%
  summarize(mean = round(mean(pp), 2),
            sd = round(sd(pp),2), n = n()) %>%
   unite("mean_sd", mean:sd, sep = "±") 
summary
```

# Figure 4 - initial pp and npp boxplot included in  main text. I will probably remove from figures

```{r}


# water mass by population
p1 <- pp_rates %>%
    ggplot() +
      geom_jitter(aes(x=cycle_name, y=pp, colour = depth), width = 0.2, alpha = 0.5, size = 1.5)+
      geom_boxplot(aes(x=water_mass, y= pp, fill=water_mass))+
      scale_colour_manual(values =  c("#009292", "#DB6D00"))+
      scale_fill_manual(values = c("#C62828", "#1565C0")) +
      facet_grid(population~water_mass, scales = "free", space = "free_x") +
      theme(strip.text.y.right = element_blank(),
            legend.position = "none",
            panel.spacing = unit(1, "lines"))+
      scale_y_continuous(trans='log10') +
      annotation_logticks(sides = "l", outside = TRUE)+
      coord_cartesian(clip = "off")+
      ylab("Carbon fixation rate (fgC/cell/hr)") +
      xlab("Cycle")
p1

p2 <- pp_rates %>%
  ggplot() +
    geom_jitter(aes(x=cycle_name, y=npp, colour = depth), width = 0.2, alpha = 0.5, size = 1.5)+
    geom_boxplot(aes(x=water_mass, y= npp, fill=water_mass))+
    scale_colour_manual(values =  c("#009292", "#DB6D00"))+
    scale_fill_manual(values = c("#C62828", "#1565C0")) +
    facet_grid(population~water_mass, scales = "free", space = "free_x") +
    theme(panel.spacing = unit(1,"lines"),
          legend.position = "top",
          strip.text.y.right = element_text(angle = 0))+
    scale_y_continuous(trans='log10') +
    annotation_logticks(sides = "l", outside = TRUE)+
    coord_cartesian(clip = "off") +
    ylab("NPP (mgC/m3/day)")+
    xlab("Cycle")
p2

library(patchwork)
plot <- p1 + p2 +  plot_annotation(tag_levels = 'A') +  plot_layout(guides = 'collect')  &
  theme(legend.position='top')
plot
# pdf("pp_npp v2.pdf", height=7, width=8) ; plot(plot) ; dev.off()

```

# Supplementary figures - Figures S6, S7, S8.

```{r}

# Figure S6 - PP by experiment
  p <- pp_rates %>%
  ggplot() +
    geom_jitter(aes(x=exp, y=pp, shape = depth), width = 0.2, alpha = 0.5) +
    scale_fill_manual (values = colours_cycle)+
    facet_grid(population~cycle_name, scales = "free", space = "free_x") +
        theme(panel.spacing.x = unit(0,"lines"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              strip.background = element_blank(),
              strip.text.y.right = element_text(angle = 0),
              panel.border = element_rect(colour = "black", fill = NA),
              legend.position = "top")+
    scale_y_continuous(trans='log10') +
    annotation_logticks(sides = "l")+
    ylab("Carbon fixation rate (fgC/cell/hr)")
p
pdf("figures/1_group rates/202231130_pp_group_exp.pdf", height=6, width=6) ; plot(p) ; dev.off()



# Figure S7 - cell conc by experiment
  p <- pp_rates %>%
  select(cycle_name, exp, conc_ml, population, depth) %>%
  filter(exp!="5") %>%
  distinct() %>%
  ggplot() +
    # geom_point(aes(x=exp, y=npp), alpha = 0.5)+
    geom_jitter(aes(x=exp, y=conc_ml, shape = depth), width = 0.2, alpha = 0.5) +
    facet_grid(population~cycle_name, scales = "free", space = "free_x") +
        theme_minimal()+
        theme(panel.spacing.x = unit(0,"lines"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              strip.background = element_blank(),
              strip.text.y.right = element_text(angle = 0),
              panel.border = element_rect(colour = "black", fill = NA),
              legend.position = "top")+
    scale_y_continuous(trans='log10') +
    annotation_logticks(sides = "l")+
    ylab("Cell concentration (mL)") +
  xlab("Experiment") +
  labs(shape = "Depth")
p
pdf("figures/1_group rates/supp_cell conc_group_exp.pdf", height=6, width=6) ; plot(p) ; dev.off()


# Figure S8- npp by experiment
  p <- pp_rates %>%
  ggplot() +
    # geom_point(aes(x=exp, y=npp), alpha = 0.5)+
    geom_jitter(aes(x=exp, y=npp, shape = depth), width = 0.2, alpha = 0.5) +
    facet_grid(population~cycle_name, scales = "free", space = "free_x") +
        theme_minimal()+
        theme(panel.spacing.x = unit(0,"lines"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              strip.background = element_blank(),
              strip.text.y.right = element_text(angle = 0),
              panel.border = element_rect(colour = "black", fill = NA),
              legend.position = "top")+
    scale_y_continuous(trans='log10') +
    annotation_logticks(sides = "l")+
    ylab("NPP (mgC/m3/day)")
p
pdf("figures/1_group rates/202231130_npp_group_exp.pdf", height=6, width=6) ; plot(p) ; dev.off()


```
