---
title: "C14 CTD_nutrients_cell abund"
author: "Denise Ong"
date: "2/28/2021"
output: pdf_document
---
Plots and tables for CTD, nutrients and cell abundance. 
- all obtained from morning CTD- 9-12pm
- remove all SA-sc-b cycles
- select only days that correspond to incubation experiments

```{r knitr_init, echo=FALSE, cache=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=TRUE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE,
               tidy = FALSE,
               fig.height=8, 
               fig.width=8,
               results = 'asis')
opts_knit$set(width=75)
```


# Load libraries

```{r, message=FALSE, warning=FALSE}
library(here)
source(here("0_init_files", "init.R")) #for libraries
source(here("0_init_files", "init_markdown.R")) #for maerkdown libraries

colours_cycle <- readxl::read_excel(here("0_init_files", "colours_cycle.xlsx"))
colours_cycle <- structure(colours_cycle$colour,.Names=colours_cycle$cycle)
```


#CTD

- depth was selected until 150 m
- cycle 1/subantarctic 1 - 24 Oct to 30 Oct
- cycle 2/subantarctic 2 - 2 Nov to 6 Nov
- cycle 3/subtropical 1 - 8 to 10 Nov
- cycle 4/ subtropical 2 - 12 to 14 Nov
- cycle 5/ subantarctic 3 - 16 to 17 Nov

##Read table
```{r}
CTD_raw <- readxl::read_excel(here("0_data_used", "TK_CTD Downcasts.xlsx")) %>%
  dplyr::rename(CTD = `cast`) %>%
  mutate(datetime =  dmy_hms(datetime)) %>%
  filter(!is.na(t)) %>%
  mutate(date = date(datetime)) %>%
  mutate(time = hour(datetime)) %>%
  mutate(julian_day = julian(date, origin = date("2018-01-01"))) %>%
  # filter(time %in% (0:3)) %>% 
  filter(time %in% (9:12) ) %>%# only morning CTD
  mutate(t_mean = (t+t2)/2) %>% # each CTD has 2 readings. use the average of the two readings.
  mutate(s_mean = (s+s2)/2) %>%
  mutate(os_mean = (os+os2)/2) %>%
  select(-t, -t2, -s, -s2, -os, -os2) %>%
  filter(p <= 150)
CTD_raw
#check CTD,  date, time
# check <- readxl::read_excel(here("0_data_used", "TK_CTD Downcasts.xlsx")) %>%
#   dplyr::rename(CTD = `cast`) %>%
#   mutate(datetime =  dmy_hms(datetime)) %>%
#   filter(!is.na(t)) %>%
#   mutate(date = date(datetime)) %>%
#   mutate(time = hour(datetime))  %>%
CTD_raw %>%
  select(date,time) %>%
  distinct()
```

```{r}
CTD <- CTD_raw %>%
  mutate(cycle = NA)

CTD$julian_day <- as.character(CTD$julian_day)
CTD$cycle[grep("297", CTD$julian_day)] = "1"
CTD$cycle[grep("298", CTD$julian_day)] = "1"
CTD$cycle[grep("299", CTD$julian_day)] = "1"
CTD$cycle[grep("300", CTD$julian_day)] = "1"
CTD$cycle[grep("301", CTD$julian_day)] = "1"
CTD$cycle[grep("302", CTD$julian_day)] = "1"
CTD$cycle[grep("305", CTD$julian_day)] = "2"
CTD$cycle[grep("306", CTD$julian_day)] = "2"
CTD$cycle[grep("307", CTD$julian_day)] = "2"
CTD$cycle[grep("308", CTD$julian_day)] = "2"
CTD$cycle[grep("309", CTD$julian_day)] = "2"
CTD$cycle[grep("311", CTD$julian_day)] = "3"
CTD$cycle[grep("312", CTD$julian_day)] = "3"
CTD$cycle[grep("313", CTD$julian_day)] = "3"
CTD$cycle[grep("315", CTD$julian_day)] = "4"
CTD$cycle[grep("316", CTD$julian_day)] = "4"
CTD$cycle[grep("317", CTD$julian_day)] = "4"
CTD$cycle[grep("319", CTD$julian_day)] = "5"
CTD$cycle[grep("320", CTD$julian_day)] = "5"

CTD <- CTD %>% mutate(cycle_name = dplyr::recode(cycle, "1" = "SA-Sc",
                                    "2" = "SA1",
                                    "3" = "ST1",
                                    "4" = "ST2",
                                    "5" = "SA2")) %>%
  mutate(CTD = as.character(CTD)) %>%
  mutate(cycle_name_2 = case_when(julian_day == "297" ~ "SA-Sc-A",
                                  julian_day == "298" ~ "SA-Sc-A",
                                  julian_day == "299" ~ "SA-Sc-A",
                                  julian_day == "300" ~ "SA-Sc-B",
                                  julian_day == "301" ~ "SA-Sc-B",
                                  julian_day == "302" ~ "SA-Sc-B",
                                  TRUE ~ as.character(cycle_name))) %>%
  filter(cycle_name_2 != "SA-Sc-B") %>% #  SA1 separated into two sub cycles. C14 exp only includes SA-Sc-A. remove SA-Sc-B
  filter(!is.na(cycle)) %>% #remove rows without cycle number
  mutate(cycle_name = fct_relevel(cycle_name, "ST1", "ST2", "SA-Sc", "SA1", "SA2")) %>%
  filter(julian_day %in% c("297", "298", "305", "306", "307", "311", "312", "315", "316", "319")) # only days corresponding to exp, except 5th and 6th nov, without afternoon CTD.

#check the dates
check <- CTD  %>% 
  select(date, cycle, time, CTD, julian_day) %>%
  distinct()
check
```
## Plots

## Main Figure 1 - T-S

```{r}
ts_plot <- CTD %>%
  ggplot(aes(x=s_mean, y=t_mean, colour=cycle_name)) + 
   # geom_errorbar(aes(xmin=s_mean-se, xmax=s_mean+se), width=.1, position=pd) +
    #geom_path() +
  geom_point(size=0.75) +
  #geom_smooth(orientation = "y", method = "gam") +
  #scale_y_reverse()+
  ylab(expression(paste("Temperature (",degree,"C)")))+
  xlab("Salinity (PSS)")+
  scale_colour_manual(values = colours_cycle) +
  guides(colour=guide_legend(title="Cycle"))

ts_plot

```

## Supplementary
```{r}

##Plot temperature
temp_plot <- CTD %>%
  group_by(cycle_name, CTD) %>%
  group_modify(~ add_row(.x,.before=0)) %>% # add gaps in the dataframe for geom_path linebreaks
  ggplot(aes(x=t_mean, y=p, colour=cycle_name)) +
 geom_point(size=0.02, alpha = 0.6) +
 # geom_path(size=0.1)+
  geom_smooth(orientation = "y", method = "loess", se = FALSE) +
  scale_y_reverse() +
  ylab("Depth (m)")+
  xlab(expression(paste("Temperature (",degree,"C)")))+
  labs(colour = "Cycle") +
  scale_colour_manual(values = colours_cycle)
 
temp_plot

##Plot salinity
s_plot <- CTD %>%
  group_by(cycle_name, CTD) %>%
  group_modify(~ add_row(.x,.before=0)) %>% # add gaps in the dataframe for geom_path linebreaks
  ggplot(aes(x=s_mean, y=p, colour=cycle_name)) + 
 #   geom_path(size=0.1) +
  geom_point(size=0.02, alpha = 0.6) +
  geom_smooth(orientation = "y", method = "loess", se = FALSE) +
  scale_y_reverse() +
  ylab(NULL)+
  xlab("Salinity (PSS)") +
  labs(colour = "Cycle") +
  scale_colour_manual(values = colours_cycle)

s_plot

#PAR
par_plot <- CTD %>%
  group_by(cycle_name, CTD) %>%
  group_modify(~ add_row(.x,.before=0)) %>% # add gaps in the dataframe for geom_path linebreaks
  # filter(p <= 100) %>%
  filter(par <=1500) %>%
  ggplot(aes(x=par, y=p, colour=cycle_name)) + 
   # geom_path(size=0.1) +
 geom_point(size=0.02, alpha = 0.6) +
  geom_smooth(orientation = "y", se = FALSE, method = "loess") +
  scale_y_reverse() +
  ylab(NULL)+
  xlab("PAR") +
  labs(colour = "Cycle") +
  scale_colour_manual(values = colours_cycle)

par_plot
```



#Nutrients
## Read table 
combine with C14 hot for the cycle, exp and station number.
```{r}
nutrients_raw <- readxl::read_excel(here("0_data_used", "TAN1810_nutrients.xls")) %>%
  dplyr::rename(CTD = `U_Cast`,
                depth =`Depth`)

CTD_raw <- readxl::read_excel(here("0_data_used", "TK_CTD Downcasts.xlsx")) %>%
  dplyr::rename(CTD = `cast`) %>%
  mutate(datetime =  dmy_hms(datetime)) %>%
  filter(!is.na(t)) %>%
  mutate(date = date(datetime)) %>%
  mutate(time = hour(datetime)) %>%
  mutate(julian_day = julian(date, origin = date("2018-01-01"))) %>%
  select(CTD, datetime, date, time, julian_day) %>%
  mutate(CTD=as.character(CTD)) %>%
  distinct()

nutrients_raw <- full_join(nutrients_raw,CTD_raw) %>%
  filter(!is.na(depth)) %>%
  filter(!is.na(julian_day)) #how to insert date for the missing dates. Corresponding CTD no. is 9167, 9171, 9169, 9168

nutrients_raw
```

```{r}
nutrients <- nutrients_raw %>%
  mutate(cycle = NA)

nutrients$julian_day <- as.character(nutrients$julian_day)
nutrients$cycle[grep("296", nutrients$julian_day)] = "1"
nutrients$cycle[grep("297", nutrients$julian_day)] = "1"
nutrients$cycle[grep("298", nutrients$julian_day)] = "1"
nutrients$cycle[grep("299", nutrients$julian_day)] = "1"
nutrients$cycle[grep("300", nutrients$julian_day)] = "1"
nutrients$cycle[grep("301", nutrients$julian_day)] = "1"
nutrients$cycle[grep("302", nutrients$julian_day)] = "1"
nutrients$cycle[grep("305", nutrients$julian_day)] = "2"
nutrients$cycle[grep("306", nutrients$julian_day)] = "2"
nutrients$cycle[grep("307", nutrients$julian_day)] = "2"
nutrients$cycle[grep("308", nutrients$julian_day)] = "2"
nutrients$cycle[grep("309", nutrients$julian_day)] = "2"
nutrients$cycle[grep("311", nutrients$julian_day)] = "3"
nutrients$cycle[grep("312", nutrients$julian_day)] = "3"
nutrients$cycle[grep("313", nutrients$julian_day)] = "3"
nutrients$cycle[grep("315", nutrients$julian_day)] = "4"
nutrients$cycle[grep("316", nutrients$julian_day)] = "4"
nutrients$cycle[grep("317", nutrients$julian_day)] = "4"
nutrients$cycle[grep("319", nutrients$julian_day)] = "5"
nutrients$cycle[grep("320", nutrients$julian_day)] = "5"

nutrients <- nutrients %>%
  filter(!is.na(cycle)) %>%
  mutate(cycle_name = dplyr::recode(cycle, "1" = "SA-Sc",
                                    "2" = "SA1",
                                    "3" = "ST1",
                                    "4" = "ST2",
                                    "5" = "SA2"))  %>%
  mutate(cycle_name_2 = case_when(julian_day == "296" ~ "SA-Sc-A",
                                  julian_day == "297" ~ "SA-Sc-A",
                                  julian_day == "298" ~ "SA-Sc-A",
                                  julian_day == "299" ~ "SA-Sc-A",
                                  julian_day == "300" ~ "SA-Sc-B",
                                  julian_day == "301" ~ "SA-Sc-B",
                                  julian_day == "302" ~ "SA-Sc-B",
                                  TRUE ~ as.character(cycle_name))) %>%
  filter(cycle_name_2 != "SA-Sc-B") %>% #  SA1 separated into two sub cycles. C14 exp only includes SA1-A.
  mutate(cycle_name = fct_relevel(cycle_name, "ST1", "ST2", "SA-Sc", "SA1", "SA2")) %>%
  filter(time %in% (10:12) ) %>% # only afternoon CTD
  filter(julian_day %in% c("297", "298", "305", "306", "307", "311", "312", "315", "316", "319"))


date_check <- nutrients  %>% 
  select(CTD, date, time, cycle_name_2, julian_day) %>%
  distinct() %>%
  arrange(date)
date_check


depth_check <- nutrients %>%
  select(cycle_name, depth) %>%
  distinct() %>%
  group_by(cycle_name) %>%
  arrange(cycle_name,depth) 

depth_check
```

# Table 1 - Summary table SST and nutrients
```{r}
summary_sst <-  CTD %>%
  filter(p <= 12) %>%
  pivot_longer(cols = c("t_mean"), names_to = "ctd", values_to = "value") %>%
  group_by(cycle_name, ctd) %>%
  summarise(mean = round(mean(value), 2),
            sd= round(sd(value),2)) %>%
  unite("SST", mean:sd, sep = " ± ") %>%
  ungroup() %>%
  select(cycle_name,  "SST")
summary_sst
```
nutrients from surface and dcm only.
```{r}
nutrients_sub <- nutrients %>%
  mutate(depth_cat = case_when(depth == "10" ~ "SUR",
                              (depth == "25" & cycle_name == "ST1" & CTD == "9142") ~ "DCM",
                              (depth == "40" & cycle_name == "ST1" & CTD == "9145") ~ "DCM",
                               (depth == "30" & cycle_name == "ST2" & CTD == "9150") ~ "DCM",
                               (depth == "40" & cycle_name == "ST2" & CTD == "9153") ~ "DCM",
                               (depth == "40" & cycle_name == "SA-Sc") ~ "DCM",
                               (depth == "40" & cycle_name == "SA1") ~ "DCM",
                               (depth == "70" & cycle_name == "SA2") ~ "DCM",
                                  TRUE ~as.character(cycle_name))) %>%
  filter(depth_cat %in% c("SUR", "DCM"))
summary <-   nutrients_sub %>%
  pivot_longer(cols = c("DRP_uM", "NH4_uM", "NO3_uM", "DRSi_uM"), names_to = "nutrients", values_to = "conc") %>%
  group_by(cycle_name, nutrients) %>%
  summarise(conc_mean = round(mean(conc), 2),
            conc_sd= round(sd(conc),2)) %>%
  unite("mean_sd", conc_mean:conc_sd, sep = " ± ") %>%
  pivot_wider(names_from = nutrients,  values_from = mean_sd) %>%
  ungroup() %>%
  select(cycle_name,  "NH4_uM", "NO3_uM","DRP_uM","DRSi_uM")
summary

summary_all <- full_join(summary_sst, summary)
summary_all
writexl::write_xlsx(summary_all, "summary_ctdnut_4.0.xlsx")
```


## Plots
supplementary
```{r}

nitrate_plot <- nutrients %>%
    arrange(cycle_name, CTD, depth)%>%
  group_by(cycle_name, CTD) %>%
  group_modify(~ add_row(.x,.before=0)) %>% # add gaps in the dataframe for geom_path linebreaks
  ggplot(aes(x=NO3_uM, y=depth, colour=cycle_name)) + 
  geom_point(size=0.2, alpha = 0.6) +
  # geom_path(size=0.2)+
  geom_smooth(orientation = "y",
             se = FALSE
              ) +
  scale_y_reverse(limits = c(100,0))+
  # scale_y_reverse()+
  ylab(NULL)+
  xlab(expression(paste("Nitrate (uM)")))+
  labs(colour = "Cycle name") +
  scale_colour_manual(values = colours_cycle)
 
nitrate_plot


ammonia_plot <- nutrients %>%
    arrange(cycle_name, CTD, depth)%>%
  group_by(cycle_name, CTD) %>%
  group_modify(~ add_row(.x,.before=0)) %>% # add gaps in the dataframe for geom_path linebreaks
  ggplot(aes(x=NH4_uM, y=depth, colour=cycle_name)) + 
  geom_point(size=0.1, alpha = 0.6) +
  geom_smooth(orientation = "y",
             se = FALSE
              ) +
  scale_y_reverse(limits = c(100,0))+
  # scale_y_reverse()+
  ylab("Depth (m)")+
  xlab(expression(paste("Ammonia (uM)")))+
  labs(colour = "Cycle name") +
  scale_colour_manual(values = colours_cycle)
 
ammonia_plot

p_plot <- nutrients %>%
  arrange(cycle_name, CTD, depth)%>%
  group_by(cycle_name, CTD) %>%
  group_modify(~ add_row(.x,.before=0)) %>% # add gaps in the dataframe for geom_path linebreaks
  ggplot(aes(x=DRP_uM, y=depth, colour=cycle_name)) + 
  geom_point(size=0.1, alpha = 0.6) +
  # geom_path(size=0.2)+
  geom_smooth(orientation = "y",
              se = FALSE
              ) +
  scale_y_reverse(limits = c(100,0))+
  ylab(NULL)+
  xlab(expression(paste("Phosphorus (uM)")))+
  labs(colour = "Cycle name")+
  scale_colour_manual(values = colours_cycle)
 
p_plot

si_plot <- nutrients %>%
  ggplot(aes(x=DRSi_uM, y=depth, colour=cycle_name)) + 
  geom_point(size=0.1, alpha = 0.6) +
  geom_smooth(orientation = "y",
              se = FALSE,
              fill = "lightgrey"
              ) +
  scale_y_reverse(limits = c(100,0))+
  ylab(NULL)+
  xlab(expression(paste("Silicate (uM)")))+
  labs(colour = "Cycle name")  +
  scale_colour_manual(values = colours_cycle)
 
si_plot
```

# cell abundance profile
Cell abundance values
- error in TAN1810_14C-Pico_25052020_QA_1.0.xlsx, U1941 CHANGED TO U9141. excel sheet change to 1.2
- station U9144 has no counts of FCM at 40 m, so using the one at 35 m
- Missing U9133 - cycle 2, experiment 5. julian day 309, 5 nov 2018, station 176. Checked the FCM CTD cast number against Moira "ALL_CTDS_BottleData", missing FCM counts for this day. I have decided to take the average counts of for cycle 2
## Read data
```{r}
#use the "complete data TAN1810" sheet for the experiment details.
C14_hot <- readxl::read_excel(here("0_data_used", "TAN1810_14C-Pico_25052020_QA_1.3.xlsx"), sheet = "Complete data TAN1810") %>%
  dplyr::select (`Cycle#`, `EXP#`, `SAMPLE`, `STN`, `DEPTH`, `CTD#`) %>%   
     dplyr::rename(cycle = `Cycle#`,
            exp = `EXP#`,
            depth = `SAMPLE`, 
            station = `STN`,
            depth_m = `DEPTH`,
            ctd = `CTD#`)%>%
    distinct()%>%
  drop_na()

# cell abund values
fcm_raw <- readxl::read_xlsx(here("0_data_used", "MD FCM DATA.xlsx"), sheet= "INITS PROFILES_CYCLE", skip = 2) %>%
  select(MD, CTD, CYCLE, DEPTH_C14, `SYN/ML`, `PICO/mL`, `NANO/mL`) %>%
  dplyr::rename(cycle =`CYCLE`,
                ctd = `CTD`,
                depth_m = `DEPTH_C14`, # changed U9144 depth
                Syn = `SYN/ML`,
                Pico = `PICO/mL`,
                Nano = `NANO/mL`) %>% 
  pivot_longer(Syn:Nano, 
               names_to = "population",
               values_to = "conc_ml") %>%
  filter(!is.na(conc_ml))
```

## plot profile
```{r}
ctd_list <- C14_hot %>%
  select(ctd) %>%
  unique()

df_fcm_profile <-inner_join(fcm_raw,ctd_list)  %>%
  filter(ctd != "U9133") %>%# this entry is an average
  mutate(cycle_name = dplyr::recode(cycle, "1" = "SA-Sc",
                                    "2" = "SA1",
                                    "3" = "ST1",
                                    "4" = "ST2",
                                    "5" = "SA2")) %>%
  mutate(cycle_name = fct_relevel(cycle_name, "ST1", "ST2", "SA-Sc", "SA1", "SA2")) %>%
  mutate(conc_ml = round(conc_ml)) %>%
  mutate(conc_trans = conc_ml /10^3)

cell_abund_plot <- function (df, pop) {
  p <- df %>%
  filter(population == pop) %>%
  arrange(cycle_name, ctd, depth_m)%>%
  group_by(cycle_name, ctd) %>%
  group_modify(~ add_row(.x,.before=0)) %>% # add gaps in the dataframe for geom_path linebreaks
  ggplot(aes(x=conc_trans, y=depth_m, colour=cycle_name)) + 
  geom_point(size=0.4, alpha = 0.8) +
  geom_smooth(orientation = "y",
             se = FALSE
              ) +
  scale_y_reverse(limits = c(100,0))+
  xlab(expression(paste("Cell abundance (x10^3 mL-1)")))+
  labs(colour = "Cycle name") +
  scale_colour_manual(values = colours_cycle)
  return(p)
}

syn_plot <- cell_abund_plot(df_fcm_profile, "Syn") +ylab("Depth (m)") +ggtitle("Synechococcus")+theme(legend.position="none")
pico_plot <- cell_abund_plot(df_fcm_profile, "Pico") +ylab(NULL) +ggtitle("Picoeukaryotes")+theme(legend.position="none")
nano_plot <- cell_abund_plot(df_fcm_profile, "Nano") +ylab(NULL) +ggtitle("Nanoeukaryotes")
```
## Figure S1
```{r}
library(patchwork)
plot <- (temp_plot | s_plot |par_plot | plot_spacer()) / 
  (ammonia_plot | nitrate_plot | p_plot | si_plot)/  
  (syn_plot | pico_plot |nano_plot | plot_spacer()) +  
  plot_annotation(tag_levels = 'A') & 
  theme(legend.position="none")
plot
pdf("ctd nut cell abund_v5.pdf", height=10, width=9) ; plot(plot) ; dev.off()

```


