---
title: "Cold vs group rates compare"
author: "Denise Ong"
date: "10/25/2022"
output: html_document
---
npp units mgC/m3/day.
Group calculations from method 2.

```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(here)
source(here("init_files", "init.R")) #for libraries
source(here("init_files", "init_markdown.R")) #for markdown libraries
```

# Read data
```{r}
cold_standard <- readxl::read_excel("group_rates_output/NPP_cold_standard.xlsx") %>% 
  rename(depth_m = `depth`,
         depth = `sample`)
cold_standard$exp <- as.character(cold_standard$exp)
group_rates <- readxl::read_excel("group_rates_output/Rates_cell_npp_method2.xlsx")
```

#Compute NPP for EACH vial by adding the NPP values of the groups
some of the rows have NA. leave for now.
how about average the NPP across the vials before adding?

```{r}
group_npp <- group_rates %>%
  select(cycle, exp, depth, vial, population, cycle_name, station, depth_m, ctd, npp) %>%
  pivot_wider(names_from = population, values_from = npp) %>%
  mutate(npp_group = Nano + Pico + Syn) %>%
  rename(replicate =`vial`)
```


# Join group and cold rates
For column replicates, the replicates for cold, standard and group are not from the same source. ie. replicate A is not the same across the different types. Just a way to number and identify the replicates.
```{r}
cold_group <- full_join(cold_standard, group_npp) %>%
  unite(cycle_exp, c(cycle_name, exp), sep = "_", remove = FALSE)

writexl::write_xlsx(cold_group, path = "group_rates_output/NPP_cold_standard_group.xlsx")

summary<- cold_group %>%
  group_by(cycle, cycle_name, station, exp, cycle_exp, depth) %>%
  dplyr::summarise(cold_mean = mean(cold_corrected, na.rm = TRUE),
                   cold_sd = sd(cold_corrected, na.rm = TRUE),
                   standard_mean = mean(standard_corrected, na.rm = TRUE),
                   standard_sd = sd(standard_corrected, na.rm = TRUE),
                   group_mean = mean(npp_group, na.rm = TRUE),
                   group_sd = sd(npp_group, na.rm = TRUE)) %>%
  mutate(diff_standard_cold = standard_mean - cold_mean) %>%
  mutate(diff_cold_group = cold_mean - group_mean) %>%
  mutate(standard_cold_ratio = standard_mean/cold_mean) %>%
  ungroup() %>%
  arrange(cycle_exp)
writexl::write_xlsx(summary, "group_rates_output/NPP_cold_standard_group_summary.xlsx")
```

#summary for cold rates
```{r}
summary<- cold_group %>%
  select(cycle_name, exp, depth, cold_corrected) %>%
  filter(!(is.na(cold_corrected))) %>%
  group_by(cycle_name, exp, depth) %>%
  summarise(value = mean(cold_corrected)) %>%
  group_by(cycle_name) %>%
  summarise(mean = round(mean(value), 2),
            # sd = round(sd(value),2)
             ) #%>%
   # unite("mean_sd", mean:sd, sep = " ± ") %>%
  # pivot_wider(names_from = "cycle_name", values_from = "mean_sd")
summary
```


#plot all- check variation between replicates
```{r}
cold_group_long <- cold_group %>%
  rename(standard = `standard_corrected`,
         cold = `cold_corrected`,
         group = `npp_group`) %>%
  select(cycle:depth_m, cycle_name, replicate_incubation:standard, group) %>%
  pivot_longer(cols = cold:group,
               values_to = "NPP",
               names_to = "type") %>%
  mutate(depth = fct_relevel(depth, "SUR", "DCM")) %>%
  mutate(cycle_name = fct_relevel(cycle_name, "ST1", "ST2", "SA1", "SA2", "SA3"))
p <- cold_group_long %>%
  ggplot(aes(x=exp, y = NPP, colour =  type, fill = type)) +
  geom_jitter(width = 0.1, alpha = 5/10) +
  # geom_bar(position = "dodge", stat = "summary", fun.y = "mean") + #to visualise mean. Not using.
 # geom_point(position=position_jitterdodge(0.01), alpha = 5/10) +
  geom_errorbar(data = plyr::ddply(cold_group_long, c("cycle_name", "depth", "exp", "type"), summarize, mean = mean(NPP, na.rm=TRUE)), aes(y=mean, ymin = mean, ymax = mean), width = 0.5) +
  facet_grid(depth~cycle_name, scales = "free_x", space = "free") +
  scale_y_continuous(trans='log10') +
  annotation_logticks(sides = "l") +
  ylab("NPP (mgC/m3/day)") 
  
p

pdf("group_rates_output/group_cold_standard_compare.pdf", height=4, width=6) ; plot(p) ; dev.off()

p <- cold_group_long %>%
  ggplot(aes(x=cycle_name, y = NPP, colour =  type, fill = type)) +
  geom_jitter(width = 0.1, alpha = 5/10) +
  # geom_bar(position = "dodge", stat = "summary", fun.y = "mean") + #to visualise mean. Not using.
 # geom_point(position=position_jitterdodge(0.01), alpha = 5/10) +
  geom_errorbar(data = plyr::ddply(cold_group_long, c("cycle_name", "depth", "type"), summarize, mean = mean(NPP, na.rm=TRUE)), aes(y=mean, ymin = mean, ymax = mean), width = 0.5) +
  facet_grid(depth~cycle_name, scales = "free_x", space = "free") +
  scale_y_continuous(trans='log10') +
  annotation_logticks(sides = "l") +
  ylab("NPP (mgC/m3/day)") 
  
p
```
```{r}
cold_group_long_sub <- cold_group_long %>%
  filter(!(type =="standard"))

p <- cold_group_long_sub %>%
  ggplot(aes(x=exp, y = NPP, colour =  type, fill = type)) +
  #geom_jitter(width = 0.1) +
  # geom_bar(position = "dodge", stat = "summary", fun.y = "mean") + #to visualise mean. Not using.
  geom_point(position=position_jitterdodge(0.15), alpha = 5/10) +
  geom_errorbar(data = plyr::ddply(cold_group_long_sub, c("cycle_name", "depth", "exp", "type"), summarize, mean = mean(NPP, na.rm=TRUE)), aes(y=mean, ymin = mean, ymax = mean), width = 0.5) +
  facet_grid(depth~cycle_name, scales = "free_x", space = "free") +
  scale_y_continuous(trans='log10') +
  annotation_logticks(sides = "l") +
  ylab("NPP (mgC/m3/day)") 
  
p
pdf("group_rates_output/group_cold_compare.pdf", height=4, width=6) ; plot(p) ; dev.off()

```

contribution % of small phytoplankton to total C fixation - ignore
```{r}
group_npp <- group_rates %>%
  select(cycle, exp, depth, vial, population, cycle_name, station, depth_m, ctd, npp) 
cold <- cold_standard %>%
  # select(-standard_corrected) %>%
  # rename(npp=cold_corrected) %>%
  select(-cold_corrected) %>%
  rename(npp=standard_corrected) %>%
  mutate(population = "cold")
cold_group <- full_join(cold, group_npp) %>%
  unite(cycle_exp, c(cycle_name, exp), sep = "_", remove = FALSE) %>%
  group_by(cycle_name, population) %>%
  summarise(npp = mean(npp, na.rm=TRUE)) %>%
  pivot_wider(names_from = population, values_from = npp) %>%
  mutate(micro = cold-Nano-Pico-Syn) %>%
  select(-cold) %>%
  pivot_longer(!cycle_name, names_to = "population",
               values_to = "npp") %>%
  mutate(cycle_name = fct_relevel(cycle_name, "ST1", "ST2", "SA1", "SA2", "SA3"))

p <- cold_group %>%
  # filter(population !="micro") %>%
  # mutate(population = fct_relevel(population, "20", "2", "0.2"))
  ggplot(aes(x=cycle_name, y=npp, fill = population)) +
  geom_bar(stat="identity", position = "fill") +
  # geom_bar(stat="identity") +
  scale_fill_viridis_d() +
  xlab(NULL) +
  ylab("Proportion of npp (%)") +
  theme(legend.position="top") 
p
```

```{r}
group_npp <- group_rates %>%
  select(cycle, exp, depth, vial, population, cycle_name, station, depth_m, ctd, npp)  %>%
  group_by(cycle_name, exp, depth, population) %>%
  summarise(npp = mean(npp))
standard <- cold_standard %>%
  select(-cold_corrected) %>%
  rename(npp=standard_corrected) %>%
  mutate(population = "standard")%>%
  filter(!(is.na(npp))) %>%
  filter(!(npp <= 0)) %>%
  group_by(cycle_name, exp, depth, population) %>%
  summarise(npp = mean(npp))

cold_group <- full_join(standard, group_npp) %>%
  # filter(!(cycle_name == "ST1" & depth == "SUR")) %>%
  # unite(cycle_exp, c(cycle_name, exp), sep = "_", remove = FALSE) %>%
  # group_by(cycle_name, population) %>%
  # summarise(npp = mean(npp, na.rm=TRUE)) %>%
  pivot_wider(names_from = population, values_from = npp) %>%
  mutate(others = standard-Nano-Pico-Syn) %>%
  mutate(sum = Nano + Pico + Syn) %>%
  select(-standard) %>%
  pivot_longer(!cycle_name, names_to = "population",
               values_to = "npp") %>%
  mutate(cycle_name = fct_relevel(cycle_name, "ST1", "ST2", "SA1", "SA2", "SA3"))

p <- cold_group %>%
  filter(population !="others") %>%
  # mutate(population = fct_relevel(population, "20", "2", "0.2"))
  ggplot(aes(x=cycle_name, y=npp, fill = population)) +
  # geom_bar(stat="identity", position = "fill") +
  geom_bar(stat="identity") +
  scale_fill_viridis_d() +
  xlab(NULL) +
  ylab("Proportion of npp (%)") +
  theme(legend.position="top") 
 # scale_y_continuous(trans='log10') + annotation_logticks(sides = "l") 
p

  

```

