---
title: "C14 cold data"
author: "Denise Ong"
date: "2/28/2021"
output: pdf_document
---

Compare the readings from the cold incubation vs standard incubation. The units is (mgC/m3/day).
Calculations for NPP_cold and NPP_standard have been done to this formula: Pvol = DIC x (DPMvol/(SA x V)) x 1.05 x 1000
Cold incubation - small vol, high C14
- 9 mL, 5 mCi/ml
Standard incubation - large vol, low C14


If standard is too high compared to cold, could be because it is on the high size fraction, increase variability.

to check:
- U9161: 70 m npp cold values are measured twice?

24 Oct 2022 notes:
- ST1, exp 7 (station 207). U9144 DCM. both cold and standard NPPs negligible primary production - however Chla concentrations at this depth close to 2 mgChla m-3 - very high ? Spike forgotten? Problem with 14C? Senescent phytoplankton ? I would probably leave this experiment out when it comes to estimate ‘average’ production or interpret more general trends. Explaining that given the relatively high chla and FCM cell counts it seems odd to have such low carbon uptake.
- U9141: I swapped surface and DCM in the cold incubation - based on chla profiles, NPP standard profiles,…that have 2-fold higher chla concentration at surface compared to 25 m
- U9164 and U9167 are in cycle 5 but have no group-specific rate measurements. 
- Cycle 1 exp 1 has no corresponding rate for cold and standard NPP measurements.

- U9161: 70 m npp cold values are measured twice?

Denise edits:
- U9141: remove 25 m DCM dark vial reading. very high compared to other dark vial readings. 
- U9133: remove NPP_cold, replicate B. leak during processing
- U9136: 40 m incubation remove NPP cold. very low readings, possibly not spiked.
- U9167: 60 m vial B, remove NPP cold measurement. Lower than dark measurement. X in comments.\



```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(here)
source(here("init_files", "init.R")) #for libraries
source(here("init_files", "init_markdown.R")) #for markdown libraries
```


# Read table for C14 cold values, merge for cycle and experiment number. Add new column to indicate dark. Added missing cycle and experiment information based on voyage report.
```{r}
C14_cold_raw <- readxl::read_excel("data_used/TAN1810_14C-Pico_25052020_QA_1.2.xlsx", sheet = "Cold 14C - Pico Results", skip = 5,  n_max = 93) %>%
  select (`Unique code or sample identifier`,`Station`,`CTD`, `Depth m`, `Sample ID`, `DPM to full volume`, `SA_dpm_mL`, `DIC`, `Replicate`,`replicate_incubation`,  `NPP_cold`, `NPP_standard`)  %>%
  dplyr::rename(name = `Unique code or sample identifier`,
            station = `Station`,
            depth = `Depth m`,
            sample_id = `Sample ID`,
            replicate = `Replicate`,
            DPM_full = `DPM to full volume`,
            SA = `SA_dpm_mL`) %>%
   add_column(dark = NA) %>% #mutate(dark = case_when (str_detect(name,"D") ~ TRUE, TRUE~FALSE))
  filter(!(depth == 200))


#use the "complete data TAN1810" sheet for the experiment details.
C14_hot <- readxl::read_excel("data_used/TAN1810_14C-Pico_25052020_QA_1.2.xlsx", sheet = "Complete data TAN1810", skip = 11) %>%
  dplyr::select (`Cycle#`, `EXP#`, `CTD#`, `SAMPLE`, `STN`, `DEPTH`) %>%   
     dplyr::rename(cycle = `Cycle#`,
            exp = `EXP#`,
            CTD = `CTD#`,
            sample = `SAMPLE`, 
            station = `STN`,
            depth = `DEPTH`)%>%
    distinct() %>%
  filter(!is.na(cycle))

C14_cold <- right_join(C14_hot, C14_cold_raw) %>%
  na_if("missing") %>% # some labels for DCM are not correct
  mutate(sample = case_when(depth <= 12 ~ "SUR",
                            depth >12 ~ "DCM"))

C14_cold$cycle <- as.character(C14_cold$cycle) 
C14_cold$exp <- as.character(C14_cold$exp)
C14_cold$NPP_standard <- as.double(C14_cold$NPP_standard)

# Dark column. Create empty col then use grep to fill in.
C14_cold$dark <- as.character(C14_cold$dark)
C14_cold$dark[grep("D", C14_cold$name)] = "dark"
C14_cold$dark <- ifelse(is.na(C14_cold$dark), 
             'light', C14_cold$dark)
#C14_cold$dar[is.na(C14_cold$dar)] <- "light"

#Add missing cycle and sample details, based on the voyage report.
C14_cold$cycle[grep("U9164", C14_cold$name)] = "5" 
C14_cold$cycle[grep("U9167", C14_cold$name)] = "5"
C14_cold$cycle[grep("U9136", C14_cold$name)] = "2"
C14_cold$exp[grep("U9136", C14_cold$name)] = "6"

```

# remove few values based on comments in excel
- U9141: remove 25 m DCM dark vial reading. very high compared to other dark vial readings. 
- U9133: remove NPP_cold, replicate B. leak during processing
- U9136: 40 m incubation remove NPP cold. very low readings, possibly not spiked.
- U9167: 60 m vial B, remove NPP cold values. Lower than dark measurement. X in comments. Why?
- U9144: remove 40 m for both NPP standard and cold. both cold and standard NPPs negligible primary production - however Chla concentrations at this depth close to 2 mgChla m-3

```{r}
C14_cold_edit <- C14_cold %>%
  pivot_longer(cols = NPP_cold:NPP_standard, 
               names_to = "type",
               values_to = "NPP") %>%
  filter(!(CTD == "U9141"& depth == '25' & replicate == "D" & type == "NPP_cold")) %>%
  filter(!(CTD == "U9133" & replicate == "B" & type == "NPP_cold")) %>%
  filter(!(CTD == "U9136" & depth == '40' & type == "NPP_cold")) %>%
  filter(!(CTD == "U9167" & depth == '60' & replicate == "B" & type == "NPP_cold")) %>%
  filter(!(CTD == "U9144" & depth == '40')) %>%
  pivot_wider(names_from = "type",
              values_from = "NPP")
```


# remove dark values
```{r}
dark_NPP <- C14_cold_edit %>% 
  filter(dark == "dark") %>%
  dplyr::rename(NPP_cold_dark = NPP_cold)  %>%
  dplyr::rename(NPP_standard_dark = NPP_standard)%>%
  select(cycle, exp, CTD, sample, replicate_incubation, NPP_cold_dark, NPP_standard_dark)
dark_NPP
# Check the values of dpm dark with histogram. The values are normally distributed, concentrated around the mean.
hist(dark_NPP$NPP_cold_dark)
hist(dark_NPP$NPP_standard_dark)
mean(dark_NPP$NPP_cold_dark, na.rm = TRUE)
mean(dark_NPP$NPP_standard_dark, na.rm = TRUE) #na.rm function to ignore NA values
```

# Corrected values = light - dark (same method as C14 hot calculations)

```{r}
cold_corrected$exp <- as.double(cold_corrected$exp)
cold_corrected <- full_join(C14_cold_edit,dark_NPP) %>% 
  filter (dark == "light") %>% 
  # DV - The next line replace dpm_dark to average dpm_dark if no value. Checked with histogram above.
  mutate(NPP_cold_dark = case_when (is.na(NPP_cold_dark) ~ mean(NPP_cold_dark, na.rm = TRUE),
                               TRUE ~ NPP_cold_dark)) %>%
 
  mutate(NPP_standard_dark = case_when (is.na(NPP_standard_dark) ~ mean(NPP_standard_dark, na.rm = TRUE), TRUE ~ NPP_standard_dark))  %>%
  mutate(cold_corrected = NPP_cold - NPP_cold_dark)%>%
  mutate(standard_corrected = NPP_standard - NPP_standard_dark) %>%
  select(cycle, exp, sample,  station, depth, name, replicate_incubation, replicate, cold_corrected, standard_corrected) %>%
  mutate(exp = case_when(is.na(exp) ~ station, TRUE ~ as.double(exp))) %>% # add station to exp in cycle 5 with no corresponding exp number
  mutate(cycle_name = recode(cycle, "1" = "SA1",
                                    "2" = "SA2",
                                    "3" = "ST1",
                                    "4" = "ST2",
                                    "5" = "SA3")) %>% 
  unite(cycle_exp, c(cycle_name, exp), sep = "_", remove = FALSE) %>%
  arrange(exp) %>%
  mutate(sample = fct_relevel(sample, "SUR", "DCM")) %>%
  mutate(cycle_name = fct_relevel(cycle_name, "ST1", "ST2", "SA1", "SA2", "SA3"))
cold_corrected$exp <- as.character(cold_corrected$exp)
cold_corrected
writexl::write_xlsx(cold_corrected, "group_rates_output/NPP_cold_standard.xlsx")
```


# Summary using the mean of replicates. The NPP of each experiment is the mean of the replicates.
```{r}
summary<- cold_corrected %>%
  group_by(cycle, cycle_name, station, exp, cycle_exp,sample) %>%
  dplyr::summarise(cold_mean = mean(cold_corrected, na.rm = TRUE),
                   cold_sd = sd(cold_corrected, na.rm = TRUE),
                   standard_mean = mean(standard_corrected, na.rm = TRUE),
                   standard_sd = sd(standard_corrected, na.rm = TRUE)) %>%
  mutate(difference = standard_mean - cold_mean) %>%
  mutate(standard_cold_ratio = standard_mean/cold_mean) %>%
  ungroup() %>%
  arrange(cycle_exp)
summary
writexl::write_xlsx(summary, "group_rates_output/NPP_cold_standard_summary.xlsx")
```

#plot all- check variation between replicates
```{r}
cold_corrected_long <- cold_corrected %>%
  pivot_longer(cols = cold_corrected:standard_corrected,
               values_to = "NPP",
               names_to = "type")
p <- cold_corrected_long %>%
  ggplot(aes(x=exp, y = NPP, colour =  type, fill = type)) +
  #geom_jitter(width = 0.1) +
  # geom_bar(position = "dodge", stat = "summary", fun.y = "mean") + #to visualise mean. Not using.
  geom_point(position=position_jitterdodge(0.15), alpha = 5/10) +
  geom_errorbar(data = plyr::ddply(cold_corrected_long, c("cycle_name", "sample", "exp", "type"), summarize, mean = mean(NPP, na.rm=TRUE)), aes(y=mean, ymin = mean, ymax = mean)) +
  facet_grid(sample~cycle_name, scales = "free_x", space = "free") +
  scale_y_continuous(trans='log10') +
  annotation_logticks(sides = "l") +
  ylab("NPP (mgC/m3/day)") 
  
p

#pdf("group_rates_output/cold_standard_compare.pdf", height=4, width=6) ; plot(p) ; dev.off()
```


#compare cold readings against the total sum npp
```{r, eval=FALSE}
summary_table <- cold_mean %>%
  dplyr::group_by(cycle, exp, sample) %>%
  dplyr::summarise(standard_mean = mean(standard), cold_mean = mean(cold) )
            
summary_table
            
group_npp <- readxl::read_excel("../data_used/C14_sum npp per group.xlsx")
group_npp

npp_compare <- full_join(group_npp, summary_table)
npp_compare
```
