---
title: "Group rates figure"
author: "Denise Ong"
date: "3/11/2021"
output: html_document
---


```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(here)
source(here("init_files", "init.R")) #for libraries
source(here("init_files", "init_markdown.R")) #for markdown libraries

#Colours
colours_cycle <- readxl::read_excel("init_files/colours_cycle.xlsx")
colours_cycle <- structure(colours_cycle$colour,.Names=colours_cycle$cycle)

colours_exp <- readxl::read_excel("init_files/colours_cycle.xlsx", sheet = "exp")
colours_exp <- structure(colours_exp$colour,.Names=colours_exp$exp)
```

# Read data
```{r}
cold_standard <- readxl::read_excel("group_rates_output/NPP_cold_standard.xlsx") %>% 
  rename(depth_m = `depth`,
         depth = `sample`)
cold_standard$exp <- as.character(cold_standard$exp)

pp_rates <- readxl::read_excel("group_rates_output/Rates_cell_npp_method2.xlsx")

#set factor levels
pp_rates$water_mass <- factor(pp_rates$water_mass, levels = c("ST", "SA"))
pp_rates$cycle_name <- factor(pp_rates$cycle_name, levels = c("ST1", "ST2", "SA1", "SA2", "SA3"))
pp_rates$population <- factor(pp_rates$population, levels = c("Syn", "Pico", "Nano"))
pp_rates$depth <- factor(pp_rates$depth, levels = c("SUR", "DCM"))
```

# summary for paper - pp, npp, cell conc
```{r}
summary <- pp_rates %>%
  select(population, cycle_name, pp, npp, conc_ml) %>%
  pivot_longer(pp:conc_ml, names_to = "data", values_to = "value") %>%
  group_by(data, population, cycle_name) %>%
  summarize(mean = round(mean(value), 2),
            sd = round(sd(value),2), 
             n = n()
            ) %>%
   unite("mean_sd", mean:sd, sep = " ± ") %>%
  pivot_wider(names_from = "cycle_name", values_from = "mean_sd") %>%
  arrange(factor(data, levels = c('pp', 'npp', 'conc_ml')))
summary

writexl::write_xlsx(summary, path = "group_summary.xlsx")


# sum of group sp NPP.
summary <- pp_rates %>%
  select(population, cycle_name, npp, exp, depth) %>%
  group_by(population, cycle_name, exp, depth) %>%
  summarize(mean = round(mean(npp), 2)) %>%
  group_by(cycle_name, exp,depth) %>%
  summarise(sum = sum(mean)) %>%
  group_by(cycle_name) %>%
  summarise(mean = round(mean(sum), 2),
            # sd = round(sd(sum),2)
            ) #%>%
  
   unite("mean_sd", mean:sd, sep = " ± ") %>%
  pivot_wider(names_from = "cycle_name", values_from = "mean_sd")
summary

```

#PP group rates

## Summary stats
```{r}
summary <- pp_rates %>%
  group_by(population, water_mass) %>%
  summarize(mean = round(mean(pp), 2),
            sd = round(sd(pp),2), n = n()) %>%
   unite("mean_sd", mean:sd, sep = "±") 
summary

summary <- pp_rates %>%
  group_by(population, water_mass) %>%
  summarize(median = round(median(npp), 2),
            mean = round(mean(npp), 2),
            sd = round(sd(npp),2), n = n()) %>%
   unite("mean_sd", mean:sd, sep = " ± ") 
summary

```
## Overall between groups
not normally distributed, using kruskal test. 
```{r}
hist(pp_rates$pp)
qqnorm(pp_rates$pp)
shapiro.test(pp_rates$pp)

hist(log10(pp_rates$pp))
qqnorm(log10(pp_rates$pp))
shapiro.test(log10(pp_rates$pp)) # still not significant

# lm <- lm(log10(pp) ~ population, data=pp_rates)
# Anova(lm, type = "3")
# pairwise.t.test(log10(pp_rates$pp), rates_data$population, p.adj='bonferroni')

kruskal.test(pp~population, data = pp_rates)
library(FSA)
dunnTest(pp~population, data = pp_rates,
         method="bonferroni")
```
## check difference between surface and DCM
### plot
```{r}
for (i in c("Syn", "Pico", "Nano")) {
  p <- pp_rates %>%
  filter(population == i) %>%
    ungroup() %>%
    complete(exp, depth) %>%
  ggplot() +
    geom_point(aes(x=depth, y=pp))+
    facet_grid(.~exp, scales = "free_x", space = "free") +
    #scale_colour_manual(values = colours_exp)+
    #scale_fill_manual (values = colours_cycle)+
    scale_y_continuous(trans='log10') +
    annotation_logticks(sides = "l")+
    ylab("PP rate (fgC/cell/hr)") +
    ggtitle(i)
  print(p)
  pdf(here(str_c("pp_depth_", i, ".pdf")), height=5, width=8) ; plot(p) ; dev.off()
}

```

### t test to check if surface is different from DCM
surface is not different from DCM for all groups
not significantly different, from here on combine surface and DCM
```{r}
library("ggpubr")

for (i in c("Syn", "Pico", "Nano")) {
t_test_df<- pp_rates %>%
  filter(population == i) %>%
  select(depth, water_mass, pp)

summary <- t_test_df %>%
  group_by(depth) %>%
  summarise(count = n(),
    mean = mean(pp, na.rm = TRUE),
    sd = sd(pp, na.rm = TRUE))
print(summary)
 p <- t_test_df %>%
   ggplot(aes(x=depth, y=pp)) + geom_point() +geom_boxplot()+ facet_grid(water_mass~depth, scales = "free") + scale_y_continuous(trans='log10') + annotation_logticks(sides = "l")  +ggtitle(i)
 print(p)

#Because p value <0.05, data is not normally distributed. Use the wilcoxon test
p<-with(t_test_df, shapiro.test(pp[depth == "SUR"]))
print(p)
p<-with(t_test_df, shapiro.test(pp[depth == "DCM"]))
print(p)
res <- wilcox.test(pp ~ depth, data = t_test_df,
                   exact = FALSE)
print(res)
}
```

## within group - ST vs SA
```{r}
for (i in c("Syn", "Pico", "Nano")) {
  p <- pp_rates %>%
  filter(population == i) %>%
  ggplot() +
    geom_point(aes(x=exp, y=pp, fill=exp), pch = 21, colour = "black", alpha = 0.5)+
    scale_fill_manual (values = colours_exp)+
    #geom_jitter(aes(x=exp, y=pp, shape = depth, fill=exp), width = 0.1, alpha = 0.5) +
    geom_boxplot(aes(x=cycle_name, y= pp, fill=cycle_name))+
    scale_fill_manual (values = colours_cycle)+
        scale_y_continuous(trans='log10') +
    annotation_logticks(sides = "l")+
    facet_grid(.~cycle_name, scales = "free_x", space = "free") +
        theme(panel.spacing.x = unit(0,"lines"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              strip.background = element_blank(),
              panel.border = element_rect(colour = "black", fill = NA))+

    ylab("PP rate (fgC/cell/hr)") +
    ggtitle(i)
  print(p)
}

```
### t test to check if ST is different from SA
ST is different from SA for syn and pico.
```{r}
library("ggpubr")

#within each group, ST VS SA.
for (i in c("Syn", "Pico", "Nano")) {
t_test_df<- pp_rates %>%
  filter(population == i) %>%
  select(water_mass, pp)

summary <- t_test_df %>%
  group_by(water_mass) %>%
  summarise(count = n(),
    mean = mean(pp, na.rm = TRUE),
    sd = sd(pp, na.rm = TRUE))
print(summary)
 p <- t_test_df %>%
   ggplot(aes(x=water_mass, y=pp)) + geom_point() +geom_boxplot()+ scale_y_continuous(trans='log10') + annotation_logticks(sides = "l")  
 print(p)

#Because p value <0.05, data is not normally distributed. Use the wilcoxon test
p<-with(t_test_df, shapiro.test(pp[water_mass == "ST"]))
print(p)
p<-with(t_test_df, shapiro.test(pp[water_mass == "SA"]))
print(p)
res <- wilcox.test(pp ~ water_mass, data = t_test_df,
                   exact = FALSE)
print(res)
}


```
In each water mass compare between groups.
```{r}
#ST is not normally distributed. Use kruskal wallis test.
for (i in c("ST")) {
  df <- pp_rates %>%
    filter(water_mass == i)
  print(kruskal.test(pp~population, data = df)) # not significant
  print(
  dunnTest(pp~population, data = df,
         method="bonferroni")
  )
}

#SA is not normally distributed. Use kruskal wallis test.
for (i in c("SA")) {
  df <- pp_rates %>%
    filter(water_mass == i)
  print(kruskal.test(pp~population, data = df)) # not significant
  print(
  dunnTest(pp~population, data = df,
         method="bonferroni")
  )
}
```

# NPP
## Overall between groups
not normally distributed, using kruskal test. 
```{r}
hist(pp_rates$npp)
qqnorm(pp_rates$npp)
shapiro.test(pp_rates$npp)

hist(log10(pp_rates$npp))
qqnorm(log10(pp_rates$npp))
shapiro.test(log10(pp_rates$npp)) # still not significant

lm <- lm(log10(npp) ~ population, data=pp_rates)
Anova(lm, type = "3")
pairwise.t.test(log10(pp_rates$pp), pp_rates$population, p.adj='bonferroni')


```

## check diff between surface and DCM
```{r}
for (i in c("Syn", "Pico", "Nano")) {
  p <- pp_rates %>%
  filter(population == i) %>%
  ggplot() +
    geom_point(aes(x=depth, y=npp, shape = depth, colour=exp))+
    facet_grid(.~exp, scales = "free_x", space = "free") +
    scale_colour_manual(values = colours_exp)+
    scale_fill_manual (values = colours_cycle)+
    scale_y_continuous(trans='log10') +
    annotation_logticks(sides = "l")+
    #ylab("PP rate (fgC/cell/hr)") +
    ggtitle(i)
  print(p)
}

```
### t test to check if surface is different from DCM
```{r}
library("ggpubr")

for (i in c("Syn", "Pico", "Nano")) {
t_test_df<- pp_rates %>%
  filter(population == i) %>%
  select(depth, npp)

summary <- t_test_df %>%
  group_by(depth) %>%
  summarise(count = n(),
    mean = mean(npp, na.rm = TRUE),
    sd = sd(npp, na.rm = TRUE))
print(summary)
 p <- t_test_df %>%
   ggplot(aes(x=depth, y=npp)) + geom_point() +geom_boxplot()+ scale_y_continuous(trans='log10') + annotation_logticks(sides = "l")  
 print(p)

#Because p value <0.05, data is not normally distributed. Use the wilcoxon test
p<-with(t_test_df, shapiro.test(npp[depth == "SUR"]))
print(p)
p<-with(t_test_df, shapiro.test(npp[depth == "DCM"]))
print(p)
res <- wilcox.test(npp ~ depth, data = t_test_df,
                   exact = FALSE)
print(res)
}
```

### Within each water mass check between groups
```{r}

for (i in c("SA")) {
  df <- pp_rates %>%
    filter(water_mass == i)
  hist(log10(df$npp))
  qqnorm(log10(df$npp))
  print(shapiro.test(log10(df$npp)))
  
  lm <- lm(log10(npp) ~ population, data=df)
  print(Anova(lm, type = "3"))
  print(pairwise.t.test(log10(df$npp), df$population, p.adj='bonferroni'))
}

#ST is not normally distributed. Use kruskal wallis test.
for (i in c("ST")) {
  df <- pp_rates %>%
    filter(water_mass == i)
  print(kruskal.test(pp~population, data = df)) # not significant
  print(
  dunnTest(pp~population, data = df, #perform Dunn's Test with Bonferroni correction for p-values
         method="bonferroni")
  )
}
```

# Plots for paper
##PP
```{r}
set.seed(1)
#this plot goes to the supplementary. Remove box plot
  p <- pp_rates %>%
  ggplot() +
    geom_jitter(aes(x=exp, y=pp, shape = depth), width = 0.2, alpha = 0.5) +
   # geom_boxplot(aes(x=cycle_name, y= pp, fill=cycle_name), alpha = 0.9)+
    scale_fill_manual (values = colours_cycle)+
    facet_grid(population~cycle_name, scales = "free", space = "free_x") +
        theme(panel.spacing.x = unit(0,"lines"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              strip.background = element_blank(),
              strip.text.y.right = element_text(angle = 0),
              panel.border = element_rect(colour = "black", fill = NA),
              legend.position = "top")+
    scale_y_continuous(trans='log10') +
    annotation_logticks(sides = "l")+
    ylab("Carbon fixation rate (fgC/cell/hr)")
p
pdf("figures/1_group rates/202231130_pp_group_exp.pdf", height=6, width=6) ; plot(p) ; dev.off()


# water mass by population
p1 <- pp_rates %>%
  #filter(cycle_name!= "SA1") %>%
  ggplot() +
    geom_jitter(aes(x=cycle_name, y=pp), width = 0.2, alpha = 0.4)+
    #geom_jitter(aes(x=exp, y=pp, shape = depth, fill=exp), width = 0.1, alpha = 0.5) +
    geom_boxplot(aes(x=water_mass, y= pp, fill=water_mass))+
  #  scale_fill_manual (values = colours_cycle) +
  scale_fill_manual(values = c("#C62828", "#1565C0")) +
    facet_grid(population~water_mass, scales = "free", space = "free_x") +
        theme(strip.text.y.right = element_blank(),
              legend.position = "none",
              panel.spacing = unit(1, "lines"))+
   # scale_x_discrete(expand=c(.08)) +
    scale_y_continuous(trans='log10') +
    annotation_logticks(sides = "l", outside = TRUE)+
  coord_cartesian(clip = "off")+
    ylab("Carbon fixation rate (fgC/cell/hr)") +
  xlab("Cycle")
p1

# population by water mass
p <- pp_rates %>%
  filter(cycle_name!= "SA1") %>%
  ggplot() +
    geom_jitter(aes(x=cycle_name, y=pp), width = 0.1, alpha = 0.4)+
    #geom_jitter(aes(x=exp, y=pp, shape = depth, fill=exp), width = 0.1, alpha = 0.5) +
    geom_boxplot(aes(x=water_mass, y= pp, fill=water_mass))+
  #  scale_fill_manual (values = colours_cycle) +
  scale_fill_manual(values = c("#C62828", "#1565C0")) +
    facet_grid(water_mass~population, scales = "free", space = "free_x") +
        theme(panel.spacing.x = unit(0,"lines"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              strip.background = element_blank(),
              strip.text.y.right = element_blank(),
              panel.border = element_rect(colour = "black", fill = NA),
              legend.position = "none")+
   # scale_x_discrete(expand=c(.08)) +
    scale_y_continuous(trans='log10') +
    annotation_logticks(sides = "l")+
    ylab("PP rate (fgC/cell/hr)") +
  xlab("Cycle")
p

```
## NPP
```{r}
# cell conc supp
  p <- pp_rates %>%
  select(cycle_name, exp, conc_ml, population, depth) %>%
  distinct() %>%
  ggplot() +
    # geom_point(aes(x=exp, y=npp), alpha = 0.5)+
    geom_jitter(aes(x=exp, y=conc_ml, shape = depth), width = 0.2, alpha = 0.5) +
    facet_grid(population~cycle_name, scales = "free", space = "free_x") +
        theme_minimal()+
        theme(panel.spacing.x = unit(0,"lines"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              strip.background = element_blank(),
              strip.text.y.right = element_text(angle = 0),
              panel.border = element_rect(colour = "black", fill = NA),
              legend.position = "top")+
    scale_y_continuous(trans='log10') +
    annotation_logticks(sides = "l")+
    ylab("Cell concentration (ml)")
p
pdf("figures/1_group rates/202231130_cell conc_group_exp.pdf", height=6, width=6) ; plot(p) ; dev.off()


#supp
  p <- pp_rates %>%
  ggplot() +
    # geom_point(aes(x=exp, y=npp), alpha = 0.5)+
    geom_jitter(aes(x=exp, y=npp, shape = depth), width = 0.2, alpha = 0.5) +
    facet_grid(population~cycle_name, scales = "free", space = "free_x") +
        theme_minimal()+
        theme(panel.spacing.x = unit(0,"lines"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              strip.background = element_blank(),
              strip.text.y.right = element_text(angle = 0),
              panel.border = element_rect(colour = "black", fill = NA),
              legend.position = "top")+
    scale_y_continuous(trans='log10') +
    annotation_logticks(sides = "l")+
    ylab("NPP (mgC/m3/day)")
p
pdf("figures/1_group rates/202231130_npp_group_exp.pdf", height=6, width=6) ; plot(p) ; dev.off()


p2 <- pp_rates %>%
  ggplot() +
    geom_jitter(aes(x=cycle_name, y=npp), width = 0.2, alpha = 0.4)+
    geom_boxplot(aes(x=water_mass, y= npp, fill=water_mass))+
    scale_fill_manual(values = c("#C62828", "#1565C0")) +
    facet_grid(population~water_mass, scales = "free", space = "free_x") +
    theme(panel.spacing = unit(1,"lines"),
          legend.position = "none",
          strip.text.y.right = element_text(angle = 0))+
    scale_y_continuous(trans='log10') +
    annotation_logticks(sides = "l", outside = TRUE)+
    coord_cartesian(clip = "off") +
    ylab("NPP (mgC/m3/day)")+
    xlab("Cycle")
p2

library(patchwork)
plot <- p1 + p2 +  plot_annotation(tag_levels = 'A')
plot
pdf("figures/1_group rates/202231213_pp_npp.pdf", height=6, width=8) ; plot(plot) ; dev.off()

```
## ignore these plots
```{r}
p <- pp_rates %>%
    ggplot(aes(x=population, y=pp)) +
    geom_boxplot()+
    geom_jitter(aes(shape =depth, colour=exp), width = 0.1) +
    scale_colour_manual(values = colours_exp)+
    facet_grid(.~cycle_name)+
    scale_y_continuous(trans='log10') +
    annotation_logticks(sides = "l") 
p

p <- pp_rates %>%
    ggplot(aes(x=cycle_name, y=pp)) +
    geom_boxplot()+
    geom_jitter(aes(shape =depth, colour=exp), width = 0.1) +
    scale_colour_manual(values = colours_exp)+
    facet_grid(.~population)+
    scale_y_continuous(trans='log10') +
    annotation_logticks(sides = "l")  +
    ylab("PP rate (fgC/cell/hr)")
p
#pdf("figures/1_group rates/pp_group.pdf", height=4, width=6) ; plot(p) ; dev.off()
```


## plot
```{r}
for (i in c("Syn", "Pico", "Nano")) {
  p <- pp_rates %>%
  filter(population == i) %>%
  ggplot() +
    geom_point(aes(x=exp, y=npp, fill=exp), pch = 21, colour = "black", alpha = 0.5)+
    scale_fill_manual (values = colours_exp)+
    #geom_jitter(aes(x=exp, y=pp, shape = depth, fill=exp), width = 0.1, alpha = 0.5) +
    geom_boxplot(aes(x=cycle_name, y= npp, fill=cycle_name))+
    scale_fill_manual (values = colours_cycle)+
    facet_grid(.~cycle_name, scales = "free_x", space = "free") +
        theme(panel.spacing.x = unit(0,"lines"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              strip.background = element_blank(),
              panel.border = element_rect(colour = "black", fill = NA))+
    scale_y_continuous(trans='log10') +
    annotation_logticks(sides = "l")+
    ylab("NPP") +
    ggtitle(i)
  print(p)
}

```
```{r}
  p <- pp_rates %>%
  ggplot() +
    geom_point(aes(x=exp, y=npp), alpha = 0.5)+
    #geom_jitter(aes(x=exp, y=pp, shape = depth, fill=exp), width = 0.1, alpha = 0.5) +
    geom_boxplot(aes(x=cycle_name, y= npp, fill=cycle_name), alpha = 0.90)+
    scale_fill_manual (values = colours_cycle)+
    facet_grid(population~cycle_name, scales = "free", space = "free_x") +
        theme(panel.spacing.x = unit(0,"lines"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              strip.background = element_blank(),
              strip.text.y.right = element_text(angle = 0),
              panel.border = element_rect(colour = "black", fill = NA),
              legend.position = "none")+
   # scale_x_discrete(expand=c(.08)) +
    scale_y_continuous(trans='log10') +
    annotation_logticks(sides = "l")+
    ylab("NPP (mgC/m3/day")
p
#pdf("NPP_group.pdf", height=8, width=6) ; plot(p) ; dev.off()

p2 <- pp_rates %>%
  ggplot() +
    geom_jitter(aes(x=cycle_name, y=npp), width = 0.1, alpha = 0.4)+
    #geom_jitter(aes(x=exp, y=pp, shape = depth, fill=exp), width = 0.1, alpha = 0.5) +
    geom_boxplot(aes(x=water_mass, y= npp, fill=water_mass))+
  #  scale_fill_manual (values = colours_cycle) +
  scale_fill_manual(values = c("#C62828", "#1565C0")) +
    facet_grid(population~water_mass, scales = "free", space = "free_x") +
        theme(panel.spacing.x = unit(0,"lines"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              strip.background = element_blank(),
              strip.text.y.right = element_text(angle = 0),
              panel.border = element_rect(colour = "black", fill = NA),
              legend.position = "none")+
   # scale_x_discrete(expand=c(.08)) +
    scale_y_continuous(trans='log10') +
    annotation_logticks(sides = "l")+
    ylab("NPP (mgC/m3/day)")+
  xlab("Cycle")
p2
```
```{r}
library(patchwork)
plot <- p+p2
plot
pdf("pp_npp_group_watermass.pdf", height=6, width=7) ; plot(plot) ; dev.off()
```
```{r}
# population by water mass
p <- pp_rates %>%
 # filter(cycle_name!= "SA1") %>%
  ggplot() +
    geom_jitter(aes(x=cycle_name, y=npp), width = 0.1, alpha = 0.4)+
    #geom_jitter(aes(x=exp, y=pp, shape = depth, fill=exp), width = 0.1, alpha = 0.5) +
    geom_boxplot(aes(x=water_mass, y= npp, fill=water_mass))+
  #  scale_fill_manual (values = colours_cycle) +
  scale_fill_manual(values = c("#C62828", "#1565C0")) +
    facet_grid(water_mass~population, scales = "free", space = "free_x") +
        theme(panel.spacing.x = unit(0,"lines"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              strip.background = element_blank(),
              strip.text.y.right = element_blank(),
              panel.border = element_rect(colour = "black", fill = NA),
              legend.position = "none")+
   # scale_x_discrete(expand=c(.08)) +
    scale_y_continuous(trans='log10') +
    annotation_logticks(sides = "l")+
   # ylab("PP rate (fgC/cell/hr)") +
  xlab("Cycle")
p
```


## ignore
```{r}
p <- pp_rates %>%
    ggplot(aes(x=population, y=npp)) +
    geom_boxplot()+
    geom_jitter(aes(shape =depth, colour=exp), width = 0.1) +
    scale_colour_manual(values = colours_exp)+
    facet_grid(.~cycle_name)+
    scale_y_continuous(trans='log10') +
    annotation_logticks(sides = "l") 
p

p <- pp_rates %>%
    ggplot(aes(x=cycle_name, y=npp)) +
    geom_boxplot()+
    geom_jitter(aes(shape =depth, colour=exp), width = 0.1) +
    scale_colour_manual(values = colours_exp)+
    facet_grid(.~population)+
    scale_y_continuous(trans='log10') +
    annotation_logticks(sides = "l") +
    ylab("NPP rate (mgC/m3/day)")
p
#pdf("figures/1_group rates/npp_group.pdf", height=4, width=6) ; plot(p) ; dev.off()
```


```{r}
p <- pp_rates %>%
    ggplot(aes(x=population, y=conc_ml)) +
    geom_boxplot()+
    geom_jitter(aes(shape =depth, colour=exp), width = 0.1) +
    scale_colour_manual(values = colours_exp)+
    facet_grid(.~cycle_name)+
    scale_y_continuous(trans='log10') +
    annotation_logticks(sides = "l") 
p

p <- pp_rates %>%
    ggplot(aes(x=cycle_name, y=conc_ml)) +
    geom_boxplot()+
    geom_jitter(aes(shape =depth, colour=exp), width = 0.1) +
    scale_colour_manual(values = colours_exp)+
    facet_grid(.~population)+
    scale_y_continuous(trans='log10') +
    annotation_logticks(sides = "l") +
    ylab("Cell conc (ml)")
p
#pdf("figures/1_group rates/cell conc.pdf", height=4, width=6) ; plot(p) ; dev.off()

```
