---
title: "chla_npp 2.0"
author: "Denise Ong"
date: "11/28/2022"
output: html_document
---

For manuscript figures and data.
remove stations that correspond to cycles SA1-B

# initialise

```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(here)
source(here("0_init_files", "init.R")) #for libraries
source(here("0_init_files", "init_markdown.R")) #for markdown libraries
#Colours
colours_cycle <- readxl::read_excel(here("0_init_files", "colours_cycle.xlsx"), sheet = "cycle")
colours_cycle <- structure(colours_cycle$colour,.Names=colours_cycle$cycle)

```

# NPP data

```{r}
npp_raw <- readxl::read_excel(here("0_data_used", "TAN1810_NPP_areal_2.0.xlsx"))%>%
  select(Cycle, Station, Depth, U_Cast, mean_NPP, sd_NPP, areal_PP) %>%
  distinct() %>%
  mutate(cycle_name = dplyr::recode(Cycle, "1" = "SA-Sc",
                                    "2" = "SA1",
                                    "3" = "ST1",
                                    "4" = "ST2",
                                    "5" = "SA2")) %>%
  rename(cycle = `Cycle`,
         station = `Station`,
         depth = `Depth`,
         CTD = `U_Cast`)
npp_raw$CTD <- sub("^", "U", npp_raw$CTD)

#use the "complete data TAN1810" sheet for the experiment details.
C14_hot <- readxl::read_excel(here("0_data_used", "TAN1810_14C-Pico_25052020_QA_1.2.xlsx"), sheet = "Complete data TAN1810", skip = 11) %>%
  dplyr::select (`Cycle#`, `EXP#`, `CTD#`,`STN`) %>%   
     dplyr::rename(cycle = `Cycle#`,
            exp = `EXP#`,
            CTD = `CTD#`, 
            station = `STN`)%>%
    distinct() %>%
  filter(!is.na(cycle))
C14_hot$cycle <- as.character(C14_hot$cycle)
C14_hot$exp <- as.character(C14_hot$exp)

npp <- full_join(npp_raw, C14_hot) %>%
  unite(station_exp, c(station, exp), sep = "_", remove = FALSE) %>%
  mutate(cycle_name = fct_relevel(cycle_name, "ST1", "ST2", "SA-Sc", "SA1", "SA2")) %>%
  filter(!(station %in% c("51", "69","90" ))) # remove stations that correspond to cycle SA1-B

depth_check <- npp %>%
  select(cycle_name, depth) %>%
  distinct() %>%
  group_by(cycle_name) %>%
  arrange(cycle_name,depth) 

depth_check
```

for depth group incubations.

```{r}
C14_hot_depth <- readxl::read_excel(here("0_data_used", "TAN1810_14C-Pico_25052020_QA_1.2.xlsx"), sheet = "Complete data TAN1810", skip = 11) %>%
  dplyr::select (`Cycle#`, `EXP#`, `CTD#`, `SAMPLE`, `STN`, `DEPTH`) %>%   
     dplyr::rename(cycle = `Cycle#`,
            exp = `EXP#`,
            CTD = `CTD#`,
            sample = `SAMPLE`, 
            station = `STN`,
            depth = `DEPTH`)%>%
    distinct() %>%
  filter(!is.na(cycle))
C14_hot_depth$cycle <- as.character(C14_hot_depth$cycle)
C14_hot_depth$exp <- as.character(C14_hot_depth$exp)
```

## summary surface and DCM

```{r}
npp_sum <- full_join(npp, C14_hot_depth) %>%
  filter(!(is.na(sample))) %>%
  mutate(mean_NPP = case_when (is.na(mean_NPP) ~ 0.14045846, # Missing one value from station 223, using mean NPP from depth at 30 m instead of 40 m
                               TRUE ~ mean_NPP)) 
summary <- npp_sum %>%
  group_by(cycle_name) %>%
  summarize(mean = round(mean(mean_NPP), 2),
             sd = round(sd(mean_NPP),2)
            ) %>%
  unite("mean_sd", mean:sd, sep = " ± ") %>%
  pivot_wider(names_from = "cycle_name", values_from = "mean_sd")
summary
```

# Chlorophyll data

Notes:

-   Only using PM chla as an indicator for phytoplankton biomass.
-   the depth for U9104, exp 1 does not correspond to the depth usually collected. I have changed to what I think is the right depth. values are similar to the other ctds for the same cycle, using depth_corrected column for depths instead.
-   AD145 - 2 depths measured at 30 m. one should be 20 and one should be 30m. using depth_corrected values
-   Based on fluoresence, separate SA-Sc-A and SA-Sc-B. SA-Sc-B has lower fluorescence. As the exps for C14 only come from SA-Sc-A, use SA-Sc-A only for SA-Sc.
-   no corresponding chla values for the same CTD for experiment 1,2,3.

## Read data

```{r}
chl <- readxl::read_excel(here("0_data_used", "TAN1810_SFChla_2.0.xls"), sheet = "chla_int") %>%
  dplyr::select(NIWA_ID, U_Cast, cycle, cycle_2, time, exp, depth_corrected, Chla_0.2:Chla_20_pct)%>%
  filter(!is.na(cycle)) %>%
  rename(depth = `depth_corrected`,
         CTD = `U_Cast`) %>%
  mutate(cycle_name = dplyr::recode(cycle, "1" = "SA-Sc",
                                    "2" = "SA1",
                                    "3" = "ST1",
                                    "4" = "ST2",
                                    "5" = "SA2")) %>%
  mutate(cycle_name_2 = dplyr::recode(cycle_2, "1a" = "SA-Sc-a",
                                    "1b" = "SA-Sc-b",
                                    "2" = "SA1",
                                    "3" = "ST1",
                                    "4" = "ST2",
                                    "5" = "SA2")) %>%
  mutate(cycle_name = fct_relevel(cycle_name, "ST1", "ST2", "SA-Sc", "SA1", "SA2")) %>%
  mutate(cycle_name_2 = fct_relevel(cycle_name_2, "ST1", "ST2", "SA-Sc-a", "SA-Sc-b", "SA1", "SA2")) %>%
  mutate(across(where(is.numeric), ~na_if(., 0))) %>% # replace 0 as NA as these values were not measured.
  drop_na(Chla_0.2,Chla_2, Chla_20)  %>% #remove rows where there are some missing measurements for the SF. I think this is better so that the values are more accurate. 
  filter(cycle_name_2 != "SA-Sc-b") %>%
  filter(time == "noon")

chl$cycle <- as.character(chl$cycle)
chl$CTD <- sub("^", "U", chl$CTD)
chl$exp <- as.character(chl$exp)

depth_check <- chl %>%
  dplyr::select(cycle_name, depth) %>%
  distinct() %>%
  group_by(cycle_name) %>%
  arrange(cycle_name,depth) 

depth_check
```

#summary table

```{r}
#by water mass
chl_summary <- chl %>%
  select(cycle_name, areal_chl) %>%
  distinct() %>%
  mutate(water_mass = dplyr::recode(cycle_name, 
                             "SA-Sc" = "SA",
                             "SA1" = "SA",
                             "SA2" = "SA",
                             "ST1" = "ST",
                             "ST2" = "ST"
                             )) %>%
  group_by(water_mass) %>%
  summarise(mean = mean(areal_chl),
            sd = sd(areal_chl),
            min = min(areal_chl),
            max = max(areal_chl))
            
chl_summary

npp_summary <- npp %>%
  select(cycle_name, areal_PP) %>%
  distinct() %>%
  mutate(water_mass = dplyr::recode(cycle_name, 
                             "SA-Sc" = "SA",
                             "SA1" = "SA",
                             "SA2" = "SA",
                             "ST1" = "ST",
                             "ST2" = "ST"
                             )) %>%
  group_by(water_mass) %>%
  summarise(mean = mean(areal_PP,  na.rm=TRUE),
            sd = sd(areal_PP,  na.rm=TRUE),
            min = min(areal_PP,  na.rm=TRUE),
            max = max(areal_PP,  na.rm=TRUE)
            )
npp_summary

# by cycle
chl_summary <- chl %>%
  filter(depth %in% c("10", "12")) %>%
  select(cycle_name, SumSFChla) %>%
  distinct() %>%
  mutate(water_mass = dplyr::recode(cycle_name, 
                             "SA-Sc" = "SA",
                             "SA1" = "SA",
                             "SA2" = "SA",
                             "ST1" = "ST",
                             "ST2" = "ST"
                             )) %>%
  group_by(cycle_name) %>%
  summarise(mean = round(mean(SumSFChla), 2),
            sd = round(sd(SumSFChla),2)) %>%
  unite("rate", mean:sd, sep=" ± ", remove = TRUE)
chl_summary

npp_summary <- npp %>%
  mutate(water_mass = dplyr::recode(cycle_name, 
                             "SA-Sc" = "SA",
                             "SA1" = "SA",
                             "SA2" = "SA",
                             "ST1" = "ST",
                             "ST2" = "ST"
                             )) %>%
  mutate(depth_cat = case_when(depth == "12" ~ "SUR",
                              (depth == "40" & cycle_name == "ST1") ~ "DCM",
                               (depth == "40" & cycle_name == "ST2") ~ "DCM",
                               (depth == "40" & cycle_name == "SA-Sc") ~ "DCM",
                               (depth == "40" & cycle_name == "SA1") ~ "DCM",
                               (depth == "70" & cycle_name == "SA2") ~ "DCM",
                                  TRUE ~as.character(cycle_name))) %>%
  filter(depth_cat %in% c("SUR", "DCM")) %>%
  select(cycle_name, mean_NPP) %>%
  distinct() %>%
  group_by(cycle_name) %>%
  summarise(mean = mean(mean_NPP,  na.rm=TRUE),
            sd = sd(mean_NPP,  na.rm=TRUE),
            min = min(mean_NPP,  na.rm=TRUE),
            max = max(mean_NPP,  na.rm=TRUE)
            )
npp_summary

```

#integrated chla and integrated npp plot

## prepare dataframe

```{r}
#list of integrated chla and max depth sampled. 
depth <- chl %>%
  group_by(cycle_name, exp, CTD) %>%
  count(CTD)

check <- chl %>%
  group_by(cycle_name, exp, CTD) %>%
  slice_tail() %>%
  dplyr::select(CTD, cycle_name, exp, depth, areal_chl) %>%
  distinct() %>%
  full_join(depth)

# to add size fraction. calculate mean pct per cycle -> calculate mean int chla per cycle
sf_mean <- chl %>%
  group_by(cycle_name) %>%
  summarise("0.2" = mean(Chla_0.2_pct),
            "2" = mean(Chla_2_pct),
            "20" = mean(Chla_20_pct)) %>%
  pivot_longer(cols = "0.2":"20", names_to = "size", values_to = "mean")
sf_sd <- chl %>%
  group_by(cycle_name) %>%
  summarise("0.2" = sd(Chla_0.2_pct),
            "2" = sd(Chla_2_pct),
            "20" = sd(Chla_20_pct)) %>%
  pivot_longer(cols = "0.2":"20", names_to = "size", values_to = "sd")

sf_pct <- full_join(sf_mean, sf_sd)

int_chl <- chl %>%
  dplyr::select(cycle_name, CTD, areal_chl) %>%
  distinct() %>%
  group_by(cycle_name) %>%
  summarise(mean_int = mean(areal_chl),
            sd_int = sd(areal_chl))
sf_chl <- full_join(sf_pct, int_chl) %>%
  mutate(sf_int = mean_int*mean/100) %>%
  mutate(size = fct_relevel(size, "20", "2", "0.2")) 
```

```{r, eval=FALSE}
#checks
sf_summary <- sf_chl %>%
  filter(size == "20") %>%
  dplyr::select(sf_int, cycle_name, size) %>%
  distinct() %>%
  filter(cycle_name == "SA-Sc") %>%
  summarise(sf_int = mean(sf_int))
sf_summary
```

## Figure 3 - depth integrated Chl a and NPP

```{r}
theme_set(theme_classic())
# two separate graphs for int chla and npp, combine tgt
p1 <- sf_chl %>%
  ggplot(aes(x=cycle_name, y=sf_int, fill = size)) +
  geom_bar(stat="identity") +
  geom_errorbar(aes(ymin=mean_int, ymax=mean_int+sd_int), width=.1)+
  scale_fill_viridis_d() +
  xlab(NULL) +
  ylab("Integrated chla (mg/m2)") +
  theme(legend.position="top") 
p1

p2<- npp %>%
  select(cycle_name, areal_PP) %>%
  distinct() %>%
  filter(!is.na(areal_PP)) %>%
  ggplot(aes(x=cycle_name, y = areal_PP)) +
  stat_summary(fun.data=mean_cl_boot, geom = "pointrange") +
  xlab(NULL) +
  ylab("Integrated NPP (mgC/m2/day)") 
p2
library(patchwork)
plot <- p1/p2 +  plot_annotation(tag_levels = 'A')
plot
# pdf("figures/2_NPP chla/20221129_int chla npp combine_v2.pdf", height=7, width=5) ; plot(plot) ; dev.off()
```

## Figure S4 - proportion of size fraction

```{r}
#proportion of sf
p <- sf_chl %>%
  ggplot(aes(x=cycle_name, y=sf_int, fill = size)) +
  geom_bar(stat="identity", position = "fill") +
  scale_fill_viridis_d() +
  xlab(NULL) +
  ylab("Proportion of chla (%)") +
  theme(legend.position="top") 

p
# pdf("figures/2_NPP chla/20221129_chla proportion_v2.pdf", height=7, width=6) ; plot(p) ; dev.off()
```

# chla and npp by depth

## Figure S3 - chla stacked area chart - combine sf chla and chla by depth

```{r}
library(viridis)
p <- chl %>%
  pivot_longer(Chla_0.2:Chla_20, names_to = "size_fraction", values_to = "chla") %>%
  group_by(cycle_name, depth, size_fraction) %>%
  summarise(mean_chla = mean(chla)) %>%
  # summarise(mean_chla = list(as_tibble(approx(mean_chla, depth, n = 1)))) %>% 
  # unnest() %>% 
  mutate(size_fraction = fct_relevel(size_fraction, "Chla_20", "Chla_2", "Chla_0.2")) %>%
  ggplot(aes(x=mean_chla, y=depth, fill=size_fraction)) +
  geom_area(size=0.25, colour = "white", orientation = "y") +
  facet_grid(cycle_name~., scales = "free", space = "free") +
  scale_y_reverse(limits = c(100,0)) +
  xlim(0,1.5) +
  ylab("Depth (m)") +
  xlab("Average chla (mg/m3)") +
  theme(strip.text.y.right = element_text(angle = 0),
        legend.position="top",
        strip.background = element_blank()) +
  scale_fill_viridis_d()

p
# pdf("figures/2_NPP chla/chla_sf_depth_v2.pdf", height=7, width=6) ; plot(p) ; dev.off()

```

## Figure S5 - NPP and Chl a by depth

```{r}

# plot all cycles together
p1 <- chl %>%
  group_by(cycle_name, CTD) %>%
  group_modify(~ add_row(.x,.before=0)) %>% # add gaps in the dataframe for geom_path linebreaks
  ggplot(aes(x=SumSFChla, y= depth, colour = cycle_name)) + 
  # geom_path(size=0.1) +
  geom_point(size=0.15, alpha=0.7) +
  geom_smooth(method = "loess", orientation = "y", se= FALSE) +
  scale_y_reverse()+
   # scale_x_continuous(trans='log10') +
   #  annotation_logticks(sides = "b")+
   ylim(100,0)+
  xlim(0,NA) +
  ylab("Depth (m)")+
  xlab("Chla (mg/m3)") +
   scale_colour_manual(values = colours_cycle) +
  theme(legend.position="none")
p1

# pdf("chla_am.pdf", height=5, width=4) ; plot(p2) ; dev.off()

p2 <- npp %>%
  filter(station !="15") %>%
    filter(!is.na(mean_NPP)) %>% #so that the lines can connect over NA values.
  group_by(cycle_name, station) %>%
  group_modify(~ add_row(.x,.before=0)) %>% # add gaps in the dataframe for geom_path linebreaks
  ggplot(aes(x=mean_NPP, y= depth, colour = cycle_name)) + 
  # geom_path(size=0.1) +
  geom_smooth(method = "loess", orientation = "y", se= FALSE) +
  geom_point(size=0.15, alpha=0.7) +
  scale_y_reverse()+
  # scale_x_continuous(trans='log10') +
  #   annotation_logticks(sides = "b")+
  ylim(100,0)+
  xlim(0,NA)+
  ylab(NULL)+
  xlab("NPP (mgC/m3/day)")+
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))#+
  scale_color_manual(values = colours_cycle) +
  theme(legend.position="bottom")

p2
#pdf("figures/NPP chla/npp_depth_avg.pdf", height=4, width=6) ; plot(p) ; dev.off()

library(patchwork)
plot <- (p1+p2) +  plot_annotation(tag_levels = 'A') + plot_layout (guides = "collect") & theme (legend.position = "top") 
plot
# pdf("figures/2_NPP chla/20221213_depth chla npp combine_v2.pdf", height=6, width=6) ; plot(plot) ; dev.off()

```
