---
title: "C14 cold data"
author: "Denise Ong"
date: "2/28/2021"
output: pdf_document
---

Compare the readings from the cold incubation vs standard incubation. The units is (mgC/m3/day).
Calculations for NPP_cold and NPP_standard have been done to this formula: Pvol = DIC x (DPMvol/(SA x V)) x 1.05 x 1000
NPP_cold - small vol, high C14
NPP_standard - large vol, low C14

```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(here)
source(here("0_init_files", "init.R")) #for libraries
source(here("0_init_files", "init_markdown.R")) #for markdown libraries
```


# Read table for C14 cold values, merge for cycle and experiment number. Add new column to indicate dark. Added missing cycle and experiment information based on voyage report.
```{r}
C14_cold_raw <- readxl::read_excel(here("0_data_used", "TAN1810_14C-Pico_25052020_QA_1.2.xlsx"), sheet = "Cold 14C - Pico Results", skip = 5,  n_max = 93) %>%
  select (`Unique code or sample identifier`,`Station`,`CTD`, `Depth m`, `Sample ID`, `DPM to full volume`, `SA_dpm_mL`, `DIC`, `Replicate`,`replicate_incubation`,  `NPP_cold`, `NPP_standard`)  %>%
  dplyr::rename(name = `Unique code or sample identifier`,
            station = `Station`,
            depth = `Depth m`,
            sample_id = `Sample ID`,
            replicate = `Replicate`,
            DPM_full = `DPM to full volume`,
            SA = `SA_dpm_mL`) %>%
   add_column(dark = NA) %>% #mutate(dark = case_when (str_detect(name,"D") ~ TRUE, TRUE~FALSE))
  filter(!(depth == 200))


#use the "complete data TAN1810" sheet for the experiment details.
C14_hot <- readxl::read_excel(here("0_data_used", "TAN1810_14C-Pico_25052020_QA_1.2.xlsx"), sheet = "Complete data TAN1810", skip = 11) %>%
  dplyr::select (`Cycle#`, `EXP#`, `CTD#`, `SAMPLE`, `STN`, `DEPTH`) %>%   
     dplyr::rename(cycle = `Cycle#`,
            exp = `EXP#`,
            CTD = `CTD#`,
            sample = `SAMPLE`, 
            station = `STN`,
            depth = `DEPTH`)%>%
    distinct() %>%
  filter(!is.na(cycle))

C14_cold <- right_join(C14_hot, C14_cold_raw) %>%
  mutate(NPP_standard = na_if(NPP_standard, "missing")) %>% 
  mutate(sample = case_when(depth <= 12 ~ "SUR",
                            depth >12 ~ "DCM")) # some labels for DCM are not correct

C14_cold$cycle <- as.character(C14_cold$cycle) 
C14_cold$exp <- as.character(C14_cold$exp)
C14_cold$NPP_standard <- as.double(C14_cold$NPP_standard)

# Dark column. Create empty col then use grep to fill in.
C14_cold$dark <- as.character(C14_cold$dark)
C14_cold$dark[grep("D", C14_cold$name)] = "dark"
C14_cold$dark <- ifelse(is.na(C14_cold$dark), 
             'light', C14_cold$dark)
#C14_cold$dar[is.na(C14_cold$dar)] <- "light"

#Add missing cycle and sample details, based on the voyage report.
C14_cold$cycle[grep("U9164", C14_cold$name)] = "5" 
C14_cold$cycle[grep("U9167", C14_cold$name)] = "5"
C14_cold$cycle[grep("U9136", C14_cold$name)] = "2"
C14_cold$exp[grep("U9136", C14_cold$name)] = "6"

```

# remove few values based on comments in excel
- U9141: remove 25 m DCM dark vial reading. very high compared to other dark vial readings. 
- U9133: remove NPP_cold, replicate B. leak during processing
- U9136: 40 m incubation remove NPP cold. very low readings, possibly not spiked.
- U9167: 60 m vial B, remove NPP cold values. Lower than dark measurement. X in comments. Why?
- U9144: remove 40 m for both NPP standard and cold. both cold and standard NPPs negligible primary production - however Chla concentrations at this depth close to 2 mgChla m-3

```{r}
C14_cold_edit <- C14_cold %>%
  pivot_longer(cols = NPP_cold:NPP_standard, 
               names_to = "type",
               values_to = "NPP") %>%
  filter(!(CTD == "U9141"& depth == '12' & replicate == "D" & type == "NPP_cold")) %>%
  filter(!(CTD == "U9133" & replicate == "B" & type == "NPP_cold")) %>%
  filter(!(CTD == "U9136" & depth == '40' & type == "NPP_cold")) %>%
  filter(!(CTD == "U9167" & depth == '60' & replicate == "B" & type == "NPP_cold")) %>%
  filter(!(CTD == "U9144" & depth == '40')) %>%
  pivot_wider(names_from = "type",
              values_from = "NPP")
```


# remove dark values
```{r}
dark_NPP <- C14_cold_edit %>% 
  filter(dark == "dark") %>%
  dplyr::rename(NPP_cold_dark = NPP_cold)  %>%
  dplyr::rename(NPP_standard_dark = NPP_standard)%>%
  select(cycle, exp, CTD, sample, replicate_incubation, NPP_cold_dark, NPP_standard_dark)
dark_NPP
# Check the values of dpm dark with histogram. The values are normally distributed, concentrated around the mean.
hist(dark_NPP$NPP_cold_dark)
hist(dark_NPP$NPP_standard_dark)
mean(dark_NPP$NPP_cold_dark, na.rm = TRUE)
mean(dark_NPP$NPP_standard_dark, na.rm = TRUE) #na.rm function to ignore NA values
```

# Corrected values = light - dark (same method as C14 hot calculations)

```{r}

cold_corrected <- full_join(C14_cold_edit,dark_NPP) %>% 
  filter (dark == "light") %>% 
  # DV - The next line replace dpm_dark to average dpm_dark if no value. Checked with histogram above.
  mutate(NPP_cold_dark = case_when (is.na(NPP_cold_dark) ~ mean(NPP_cold_dark, na.rm = TRUE),
                               TRUE ~ NPP_cold_dark)) %>%
 
  mutate(NPP_standard_dark = case_when (is.na(NPP_standard_dark) ~ mean(NPP_standard_dark, na.rm = TRUE), TRUE ~ NPP_standard_dark))  %>%
  mutate(cold_corrected = NPP_cold - NPP_cold_dark)%>%
  mutate(standard_corrected = NPP_standard - NPP_standard_dark) %>%
  select(cycle, exp, sample,  station, depth, name, replicate_incubation, replicate, cold_corrected, standard_corrected) %>%
  mutate(exp = case_when(is.na(exp) ~ station, TRUE ~ as.double(exp))) %>% # add station to exp in cycle 5 with no corresponding exp number
  mutate(cycle_name = dplyr::recode(cycle, "1" = "SA-Sc",
                                    "2" = "SA1",
                                    "3" = "ST1",
                                    "4" = "ST2",
                                    "5" = "SA2")) %>% 
  unite(cycle_exp, c(cycle_name, exp), sep = "_", remove = FALSE) %>%
  arrange(exp) %>%
  mutate(sample = fct_relevel(sample, "SUR", "DCM")) %>%
  mutate(cycle_name = fct_relevel(cycle_name, "ST1", "ST2", "SA-Sc", "SA1", "SA2"))
cold_corrected$exp <- as.character(cold_corrected$exp)
cold_corrected$exp <- as.double(cold_corrected$exp)
cold_corrected
writexl::write_xlsx(cold_corrected, here("Output", "Raw data", "NPP_cold_standard.xlsx"))
```


# Summary using the mean of replicates. The NPP of each experiment is the mean of the replicates.
```{r}
summary<- cold_corrected %>%
  group_by(cycle, cycle_name, station, exp, cycle_exp,sample) %>%
  dplyr::summarise(cold_mean = mean(cold_corrected, na.rm = TRUE),
                   cold_sd = sd(cold_corrected, na.rm = TRUE),
                   standard_mean = mean(standard_corrected, na.rm = TRUE),
                   standard_sd = sd(standard_corrected, na.rm = TRUE)) %>%
  mutate(difference = standard_mean - cold_mean) %>%
  mutate(standard_cold_ratio = standard_mean/cold_mean) %>%
  ungroup() %>%
  arrange(cycle_exp)
summary
writexl::write_xlsx(summary, here("Output", "Raw data", "NPP_cold_standard_summary.xlsx"))
```


