---
title: "Cold vs group rates compare"
author: "Denise Ong"
date: "10/25/2022"
output: html_document
---
npp units mgC/m3/day.
Group calculations from method 2.

Standard rates: NPP from standard volume incubation
Cold rates: NPP from small volume incubation
Group rates: NPP from summation of Syn, Pico and Nano NPP.

```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(here)
source(here("0_init_files", "init.R")) #for libraries
source(here("0_init_files", "init_markdown.R")) #for markdown libraries
```

# Read data
```{r}
cold_standard <- readxl::read_excel(here("Output", "Raw data", "NPP_cold_standard.xlsx")) %>% 
  rename(depth_m = `depth`,
         depth = `sample`)
cold_standard$exp <- as.character(cold_standard$exp)
group_rates <- readxl::read_excel(here("Output", "Raw data","Rates_cell_npp_method2.xlsx"))
```

#Compute NPP for EACH vial by adding the NPP values of group-specific rates.
```{r}
group_npp <- group_rates %>%
  select(cycle, exp, depth, vial, population, cycle_name, station, depth_m, ctd, npp) %>%
  pivot_wider(names_from = population, values_from = npp) %>%
  mutate(npp_group = Nano + Pico + Syn) %>%
  rename(replicate =`vial`)
```


# Join group and cold rates
For column replicates, the replicates for cold, standard and group are not from the same source. ie. replicate A is not the same across the different types. Just a way to number and identify the replicates.
```{r}
cold_group <- full_join(cold_standard, group_npp) %>%
  unite(cycle_exp, c(cycle_name, exp), sep = "_", remove = FALSE)

#checks
check<- cold_group %>%
  group_by(cycle, cycle_name, station, exp, cycle_exp, depth) %>%
  dplyr::summarise(cold_mean = mean(cold_corrected, na.rm = TRUE),
                   cold_sd = sd(cold_corrected, na.rm = TRUE),
                   standard_mean = mean(standard_corrected, na.rm = TRUE),
                   standard_sd = sd(standard_corrected, na.rm = TRUE),
                   group_mean = mean(npp_group, na.rm = TRUE),
                   group_sd = sd(npp_group, na.rm = TRUE)) %>%
  mutate(diff_standard_cold = standard_mean - cold_mean) %>%
  mutate(diff_cold_group = cold_mean - group_mean) %>%
  mutate(standard_cold_ratio = standard_mean/cold_mean) %>%
  ungroup() %>%
  arrange(cycle_exp)
check
```

#summary for cold rates
```{r}
summary<- cold_group %>%
  select(cycle_name, exp, depth, cold_corrected) %>%
  filter(!(is.na(cold_corrected))) %>%
  group_by(cycle_name, exp, depth) %>%
  summarise(value = mean(cold_corrected)) %>%
  group_by(cycle_name) %>%
  summarise(mean = round(mean(value), 2),
            # sd = round(sd(value),2)
             ) #%>%
   # unite("mean_sd", mean:sd, sep = " ± ") %>%
  # pivot_wider(names_from = "cycle_name", values_from = "mean_sd")
summary

```

# Statistical test to check difference between standard and small volume NPP incubation.
No significant difference.
```{r}
t_test_df<- cold_group %>%
  select(cycle_name, exp, depth, cold_corrected, standard_corrected) %>%
  filter(!(is.na(cold_corrected))) %>%
  filter(!(is.na(standard_corrected))) %>%
  filter(standard_corrected >0) 
hist(t_test_df$standard_corrected)

 p <- t_test_df %>%
   ggplot(aes(x=type, y=npp)) + geom_point() +geom_boxplot()
 print(p)
 
#Because p value <0.05, data is not normally distributed. Use the wilcoxon test
p<-with(t_test_df, shapiro.test(cold_corrected))
print(p)
p<-with(t_test_df, shapiro.test(standard_corrected))
print(p)

# wilcoxon test
t_test_df <- t_test_df %>%
  pivot_longer(cold_corrected:standard_corrected, names_to = "type", values_to = "npp")
res <- wilcox.test(npp ~ type, data = t_test_df,
                   exact = FALSE)
print(res)

```


#prepare dataframe
```{r}
cold_group_long <- cold_group %>%
  rename(standard = `standard_corrected`,
         small = `cold_corrected`,
         group = `npp_group`) %>%
  select(cycle:depth_m, cycle_name, replicate_incubation:standard, group) %>%
  pivot_longer(cols = small:group,
               values_to = "NPP",
               names_to = "type") %>%
  mutate(depth = fct_relevel(depth, "SUR", "DCM")) %>%
  mutate(cycle_name = fct_relevel(cycle_name, "ST1", "ST2", "SA-Sc", "SA1", "SA2")) %>%
  filter(type!= "group") %>%
  filter(exp!="1")
```

# Figure S2 - compare small and standard vol incubation.
```{r}
p <- cold_group_long %>%
  ggplot(aes(x=exp, y = NPP, colour =  type, fill = type)) +
  geom_jitter(width = 0.1, alpha = 5/10) +
  # geom_bar(position = "dodge", stat = "summary", fun.y = "mean") + #to visualise mean. Not using.
 # geom_point(position=position_jitterdodge(0.01), alpha = 5/10) +
  geom_errorbar(data = plyr::ddply(cold_group_long, c("cycle_name", "depth", "exp", "type"), summarize, mean = mean(NPP, na.rm=TRUE)), aes(y=mean, ymin = mean, ymax = mean), width = 0.5) +
  facet_grid(depth~cycle_name, scales = "free_x", space = "free") +
  # scale_y_continuous(trans='log10') +
  # annotation_logticks(sides = "l") +
  ylab("NPP (mgC/m3/day)") +
        theme(panel.spacing.x = unit(0,"lines"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              strip.background = element_blank(),
              strip.text.y.right = element_text(angle = 0),
              panel.border = element_rect(colour = "black", fill = NA),
              legend.position = "top")+
  scale_color_brewer(palette = "Dark2")+
  xlab("Experiment/Station")
  
p

pdf("cold_standard_compare.pdf", height=7, width=8) ; plot(plot) ; dev.off()


```

