---
title: "18S barchart"
author: "Denise Ong"
date: "11/6/2022"
output: html_document
---

Plotting a median range chart and heatmap using median relative abundance. 
Only using light incubated samples. Initials and dark are controls. already checked previously.
# Initialize
```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(here)
source(here("3_Phyloseq output and figures", "init_files", "init.R")) #for libraries
source(here("3_Phyloseq output and figures", "init_files", "init_markdown.R")) #for markdown libraries
source(here("3_Phyloseq output and figures", "init_files", "define_functions.R")) #functions to 1.transform to dataframe, 2.normalise and 3.create treemaps 
source(here("3_Phyloseq output and figures", "init_files", "colours_18sv4.R")) # predetermined colours
#source(here("phyloseq", "init_files", "colours_random.R")) #random colours
source(here("3_Phyloseq output and figures", "trophic_mode", "trophic_function.R"))  # Function to assign trophic function for phyloseq objects. 2 
# source(here("3_Phyloseq output and figures", "init_files", "top_taxa_func.R")) #Function to select top taxa
library(ggtext)
# Colours
# colours_cycle <- readxl::read_excel(here("phyloseq", "init_files","colours_cycle.xlsx"))
# colours_cycle <- structure(colours_cycle$colour,.Names=colours_cycle$cycle)
```

# Sorted pico
## Read data
remove heterotrophs and 2 samples 
only include light incubated samples
```{r}
source(here("3_Phyloseq output and figures", "init_files", "read_pico.R"))

#Assign trophic mode and remove heterotrophics that are not dinophyceae
ps_pico <- phyloseq_assign_trophic(ps_pico) %>%
    subset_taxa(trophic_mode_2 != "heterotrophic")

# Do normalizations and transformation
ps_pico <- phyloseq_normalize_median(ps_pico)
long_pico <- phyloseq_transform_to_long(ps_pico) %>%
    filter(!(file_code %in% c("pico-01-30", "pico-15"))) %>% #filter problem samples
    filter(!(cycle_exp =="SA1_2" & SAMPLE == "DCM")) %>% #only intials in these experiments
    filter(!(cycle_exp == "SA1_1" & SAMPLE == "DCM")) 

#create levels for factors
long_pico <- long_pico %>%
    mutate(cycle_exp = fct_relevel(cycle_exp, "ST1_7", "ST1_8", "ST2_9", "ST2_10", "SA-Sc_1", "SA-Sc_2", "SA1_3", "SA1_4", "SA1_5", "SA1_6", "SA2_11")) %>%
    mutate(trophic_mode_2 = fct_relevel(trophic_mode_2, "photosynthetic", "mixotrophic", "dinophyceae")) %>%
    mutate(cycle_name_2 = fct_relevel(cycle_name_2, "ST1", "ST2", "SA-Sc", "SA1", "SA2")) %>%
    mutate(vial_type = fct_relevel(vial_type, "initial", "light", "dark")) %>%
    mutate(depth_cat = fct_relevel(SAMPLE, "SUR", "DCM"))

```

## dataframe prep
```{r}

long_pico_1 <- long_pico %>%
     filter(vial_type=="light")
check <- long_pico_1 %>%
    select(asv_code:species) %>%
    unique()

# pivot longer dataframe
otu_rel_abund <- long_pico_1 %>%
    group_by(file_code) %>%
    mutate(rel_abund = n_reads / sum(n_reads)) %>%
    ungroup() %>%
    select(-n_reads) %>%
    pivot_longer(cols = asv_code:species,
                 names_to = "level",
                 values_to = "taxon") 
# for log scale
# nseqs_per_sample <- otu_rel_abund %>%
#     group_by(file_code) %>%
#     summarise(N=sum(rel_abund)) %>%
#     count(N) %>%
#     pull(N)
# 
# lod <- 1/nseqs_per_sample

```


## Class - median for each cycle more than >5%
```{r}
taxon_rel_abund <- otu_rel_abund %>%
    filter(level == "class") %>% #choose level here 
    group_by(water_mass, cycle_name_2, file_code, taxon) %>%
    summarise(rel_abund = 100*sum(rel_abund)) %>%
    pivot_wider(names_from = taxon, values_from = rel_abund) %>% #from here is to replace any file_code with NA as 0
    replace(., is.na(.), 0) %>%
    pivot_longer(cols= !water_mass:file_code,
                 names_to = "taxon",
                 values_to = "rel_abund")

## summary for results
summary <- taxon_rel_abund %>%
  filter(taxon %in% c("Mamiellophyceae", "Chrysophyceae","Pelagophyceae","Prymnesiophyceae", "Chloropicophyceae")) %>%
  filter(cycle_name_2 != "SA1") %>%
  group_by(taxon, water_mass) %>%
  summarise(median = median(rel_abund),
            mean = mean(rel_abund))
summary
```

```{r}
taxon_pool <- taxon_rel_abund %>%
    group_by(cycle_name_2, taxon) %>%
    summarise(median = median(rel_abund)) %>% #calculate median proportion of reads per sp per cycle
    group_by(taxon) %>%
    summarise(pool = max(median) <= 5, #species with a max median proportion of reads in any cycle < 5% is under Others
              median = max(median),
              .groups = "drop")

 #heatmap
 p1 <- inner_join(taxon_rel_abund, taxon_pool, by="taxon") %>%
    mutate(taxon=if_else(pool, "Other", taxon)) %>%
    group_by(file_code, cycle_name_2, taxon) %>%
    summarise(rel_abund = sum(rel_abund), 
              .groups = "drop") %>%
  group_by(cycle_name_2, taxon) %>%
  summarise(rel_abund = median(rel_abund)) %>% # for median reads per taxon per cycle
     mutate(taxon=fct_relevel(taxon,
                             "Other",
                             "Chrysophyceae",
                             "Pelagophyceae",
                             "Prymnesiophyceae",
                             "Mamiellophyceae")) %>%
  ggplot(aes(x=cycle_name_2, y = taxon, fill =rel_abund)) +
    geom_tile() +
    geom_text(aes(label=format(round(rel_abund,1), nsmall = 1)), size=3)+ # for number labels
   scale_fill_distiller(type = "seq", 
                        palette = "BuGn",
                        name = "Median<br>Relative<br>Abund (%)",
                       expand = c(0,0),
                       trans = "log",
        na.value = "#F7FCFD",
        direction = 1,
        breaks = c(0, 0.1, 1, 10, 100),
        limits = c(NA,100))+
  theme_minimal() +
  theme(text=element_text(size=10),
              strip.text.y.left = element_text(angle = 0),
              panel.spacing.x = unit(0,"lines"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              strip.background = element_blank(),
          axis.line = element_blank(),
          axis.ticks = element_blank(),
          legend.title = element_markdown()) +
  scale_x_discrete(position = "top") +
  xlab(NULL)+
  ylab(NULL)
p1
# pdf("20221130_pico_class_heatmap.pdf", height=5, width=5) ; plot(p) ; dev.off()

```


## species - select median >5% for any cycle
```{r}
taxon_rel_abund <- otu_rel_abund %>%
    filter(level == "species") %>% #choose level here 
    group_by(cycle_name_2, file_code, taxon) %>%
    summarise(rel_abund = 100*sum(rel_abund)) %>%
    pivot_wider(names_from = taxon, values_from = rel_abund) %>% #from here is to replace any file_code with NA as 0
    replace(., is.na(.), 0) %>%
    pivot_longer(cols= !cycle_name_2:file_code,
                 names_to = "taxon",
                 values_to = "rel_abund")


## summary for results
summary <- taxon_rel_abund %>%
  filter(taxon %in% c("Micromonas_commoda_A2", "Bathycoccus_prasinos","Ostreococcus_lucimarinus")) %>%
  filter(cycle_name_2 %in% c("SA2", "SA3")) %>%
  group_by(taxon) %>%
  summarise(median = median(rel_abund),
            mean = mean(rel_abund))
summary
```

```{r}
taxon_pool <- taxon_rel_abund %>%
    group_by(cycle_name_2, taxon) %>%
    summarise(median = median(rel_abund)) %>% #calculate median proportion of reads per sp per cycle
    group_by(taxon) %>%
    summarise(pool = max(median) <= 5, #species with a max median proportion of reads in any cycle < 5% is under Others
              median = max(median),
              .groups = "drop")


#get class name for facet grid grouping with class
class_df <- long_pico_1 %>%
    select(class, species) %>%
    distinct()

#heatmap
p2 <- inner_join(taxon_rel_abund, taxon_pool, by="taxon") %>%
    mutate(taxon=if_else(pool, "Other", taxon)) %>% 
    group_by(file_code, cycle_name_2, taxon) %>%
      summarise(rel_abund = sum(rel_abund), #add Others taxon together
              .groups = "drop") %>%
    group_by(cycle_name_2, taxon) %>%
   summarise(rel_abund = median(rel_abund)) %>% # for median reads per taxon per cycle
   left_join(class_df , by=c("taxon" = "species")) %>% # get class levels for sp.
      mutate(class = fct_relevel(class,
                               "Mamiellophyceae",
                               "Prymnesiophyceae",
                               "Pelagophyceae",
                               "Chrysophyceae",
                              # "Chloropicophyceae",
                               "NA")) %>%
    filter(taxon!="Other") %>%
     ggplot(aes(x=cycle_name_2, y = taxon, fill =rel_abund)) +
    geom_tile() +
  # geom_text(aes(label=format(round(rel_abund,1), nsmall = 1)), size=3)+
   scale_fill_distiller(type = "seq", 
                        palette = "BuGn",
                        name = "Median<br>Relative<br>Abund (%)",
                       expand = c(0,0),
                       trans = "log",
        na.value = "#F7FCFD",
        direction = 1,
        breaks = c(0, 0.1, 1, 10, 100),
        limits = c(NA,100))+
  theme_classic()+
    theme(text=element_text(size=10),
              strip.text.y.left = element_text(angle = 0),
              panel.spacing.y = unit(0,"lines"),
              # panel.grid.major = element_blank(),
              # panel.grid.minor = element_blank(),
              strip.background = element_blank(),
          axis.line = element_blank(),
          axis.ticks = element_blank(),
          legend.title = element_markdown(),
          strip.placement = "outside") +
  scale_x_discrete(position = "top") +
  xlab(NULL)+
  ylab(NULL) +
    facet_grid(class~., scales = "free_y", space = "free", switch = "y") +
    ggtitle("Picoeukaryotes - species level")
    
p2
# pdf("20221130_pico_species_heatmap.pdf", height=5, width=7) ; plot(p) ; dev.off()

```

```{r}
#no. of samples per cycle
sample_no <- long_pico_1 %>%
  select(cycle_name_2, file_code) %>%
  distinct() %>%
  group_by(cycle_name_2) %>%
  summarise(n = n())
sample_no
```


##order 
```{r}
taxon_rel_abund <- otu_rel_abund %>%
    filter(level == "order") %>% #choose level here 
    group_by(water_mass, cycle_name_2, file_code, taxon) %>%
    summarise(rel_abund = 100*sum(rel_abund)) %>%
    pivot_wider(names_from = taxon, values_from = rel_abund) %>% #from here is to replace any file_code with NA as 0
    replace(., is.na(.), 0) %>%
    pivot_longer(cols= !water_mass:file_code,
                 names_to = "taxon",
                 values_to = "rel_abund")

## summary for results
# summary <- taxon_rel_abund %>%
#   filter(taxon %in% c("Mamiellophyceae", "Chrysophyceae","Pelagophyceae","Prymnesiophyceae", "Chloropicophyceae")) %>%
#   filter(cycle_name_2 != "SA1") %>%
#   group_by(taxon, water_mass) %>%
#   summarise(median = median(rel_abund),
#             mean = mean(rel_abund))
# summary
```

```{r}
taxon_pool <- taxon_rel_abund %>%
    group_by(cycle_name_2, taxon) %>%
    summarise(median = median(rel_abund)) %>% #calculate median proportion of reads per sp per cycle
    group_by(taxon) %>%
    summarise(pool = max(median) <= 5, #species with a max median proportion of reads in any cycle < 5% is under Others
              median = max(median),
              .groups = "drop")
#get class name for facet grid grouping with class
class_df <- long_pico_1 %>%
    select(class, order) %>%
    distinct()
 #heatmap
 p1 <- inner_join(taxon_rel_abund, taxon_pool, by="taxon") %>%
     mutate(taxon=if_else(pool, "Other", taxon)) %>%
     group_by(file_code, cycle_name_2, taxon) %>%
     summarise(rel_abund = sum(rel_abund), 
                  .groups = "drop") %>%
     group_by(cycle_name_2, taxon) %>%
     summarise(rel_abund = median(rel_abund)) %>% # for median reads per taxon per cycle
     left_join(class_df , by=c("taxon" = "order")) %>% # get class levels for sp.
     filter(taxon != "Other") %>%
     mutate(class=fct_relevel(class,
                            "Mamiellophyceae",
                             "Prymnesiophyceae",
                             "Pelagophyceae",
                             "Chrysophyceae")) %>%
     ggplot(aes(x=cycle_name_2, y = taxon, fill =rel_abund)) +
     geom_tile() +
    # geom_text(aes(label=format(round(rel_abund,1), nsmall = 1)), size=3)+ # for number labels
    scale_fill_distiller(type = "seq", 
                        palette = "BuGn",
                        name = "Median<br>Relative<br>Abund (%)",
                       expand = c(0,0),
                       # trans = "log",
                        # na.value = "#F7FCFD",
                        direction = 1,
                        # breaks = c(0, 0.1, 1, 10, 100),
                        limits = c(NA,100))+
     theme_minimal() +
     theme(text=element_text(size=10),
          strip.text.y.left = element_text(angle = 0),
          panel.spacing.x = unit(0,"lines"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          strip.background = element_blank(),
          axis.line = element_blank(),
          axis.ticks = element_blank(),
          legend.title = element_markdown()) +
      scale_x_discrete(position = "top") +
      xlab(NULL)+
      ylab(NULL) +
     facet_grid(class~., scales = "free_y", space = "free", switch = "y") +
     ggtitle("Picoeukaryotes - order level (not log scale)")

p1
# pdf("pico_order_heatmap.pdf", height=5, width=5) ; plot(p) ; dev.off()

 #heatmap
 p1 <- inner_join(taxon_rel_abund, taxon_pool, by="taxon") %>%
     mutate(taxon=if_else(pool, "Other", taxon)) %>%
     group_by(file_code, cycle_name_2, taxon) %>%
     summarise(rel_abund = sum(rel_abund), 
                  .groups = "drop") %>%
     group_by(cycle_name_2, taxon) %>%
     summarise(rel_abund = median(rel_abund)) %>% # for median reads per taxon per cycle
     left_join(class_df , by=c("taxon" = "order")) %>% # get class levels for sp.
     filter(taxon != "Other") %>%
     mutate(class=fct_relevel(class,
                            "Mamiellophyceae",
                             "Prymnesiophyceae",
                             "Pelagophyceae",
                             "Chrysophyceae")) %>%
     ggplot(aes(x=cycle_name_2, y = taxon, fill =rel_abund)) +
     geom_tile() +
    # geom_text(aes(label=format(round(rel_abund,1), nsmall = 1)), size=3)+ # for number labels
    scale_fill_distiller(type = "seq", 
                        palette = "BuGn",
                        name = "Median<br>Relative<br>Abund (%)",
                       expand = c(0,0),
                       trans = "log",
                       na.value = "#F7FCFD",
                        direction = 1,
                        breaks = c(0, 0.1, 1, 10, 100),
                        limits = c(NA,100))+
     theme_minimal() +
     theme(text=element_text(size=10),
          strip.text.y.left = element_text(angle = 0),
          panel.spacing.x = unit(0,"lines"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          strip.background = element_blank(),
          axis.line = element_blank(),
          axis.ticks = element_blank(),
          legend.title = element_markdown()) +
      scale_x_discrete(position = "top") +
      xlab(NULL)+
      ylab(NULL) +
     facet_grid(class~., scales = "free_y", space = "free", switch = "y") +
     ggtitle("Picoeukaryotes - order level")

p1
```

# Sorted nano
## Read data
```{r}
source(here("3_Phyloseq output and figures", "init_files", "read_nano.R"))

#assign trophic mode and remove heterotrophics that are not dinophyceae
ps_nano <- phyloseq_assign_trophic(ps_nano)%>%
    subset_taxa(trophic_mode_2 != "heterotrophic")

# Do normalizations and long form, remove few samples.
ps_nano <- phyloseq_normalize_median(ps_nano)
long_nano <- phyloseq_transform_to_long(ps_nano) %>%
    filter(!(file_code %in% c("nano-12"))) 

#order all the factors
long_nano <- long_nano %>%
  mutate(cycle_exp = fct_relevel(cycle_exp, "ST1_7", "ST1_8", "ST2_9", "ST2_10", "SA-Sc_1", "SA-Sc_2", "SA1_3", "SA1_4", "SA1_5", "SA1_6", "SA2_11")) %>%
  mutate(trophic_mode = fct_relevel(trophic_mode, "photosynthetic", "mixotrophic"))  %>%
  mutate(trophic_mode_2 = fct_relevel(trophic_mode_2, "photosynthetic", "mixotrophic",  "dinophyceae"))  %>%
  mutate(cycle_name_2 =  fct_relevel(cycle_name_2,"ST1", "ST2", "SA-Sc", "SA1", "SA2"))  %>%
  mutate(vial_type = fct_relevel(vial_type, "initial", "light", "dark"))
```
## prepare data frame
```{r}

long_nano_1 <- long_nano %>%
   #  ps_top_asv(10) %>% #subset top 10 ASVs per sample
     filter(vial_type=="light")

check <- long_nano_1 %>%
    select(asv_code:species) %>%
    unique()

# pivot longer dataframe
otu_rel_abund <- long_nano_1 %>%
    group_by(file_code) %>%
    mutate(rel_abund = n_reads / sum(n_reads)) %>%
    ungroup() %>%
    select(-n_reads) %>%
    pivot_longer(cols = asv_code:species,
                 names_to = "level",
                 values_to = "taxon") 
```



## Order - select above 5%
```{r}
taxon_rel_abund <- otu_rel_abund %>%
    filter(level == "order") %>% #choose level here 
    group_by(water_mass, cycle_name_2, file_code, taxon) %>%
    summarise(rel_abund = 100*sum(rel_abund)) %>%
    pivot_wider(names_from = taxon, values_from = rel_abund) %>% #from here is to replace any file_code with NA as 0
    replace(., is.na(.), 0) %>%
    pivot_longer(cols= !water_mass:file_code,
                 names_to = "taxon",
                 values_to = "rel_abund")

summary <- taxon_rel_abund %>%
  # filter(taxon %in% c("Bacillariophyta_X", "Cryptomonadales",  "Haptophyta_Clade_HAP3_X")) %>%
  group_by(taxon, water_mass) %>%
  summarise(median = median(rel_abund),
            mean = mean(rel_abund))
summary

```

```{r}
taxon_pool <- taxon_rel_abund %>%
    group_by(cycle_name_2, taxon) %>%
    summarise(median = median(rel_abund)) %>% #calculate median proportion of reads per sp per cycle
    group_by(taxon) %>%
    summarise(pool = max(median) <= 5, #species with a max median proportion of reads in any cycle < 5% is under Others
              median = max(median),
              .groups = "drop")

#get class name for facet grid grouping with class
class_df <- long_nano_1 %>%
    select(class, order) %>%
    distinct()

#heatmap
p3 <- inner_join(taxon_rel_abund, taxon_pool, by="taxon") %>%
    mutate(taxon=if_else(pool, "Other", taxon)) %>% 
    group_by(cycle_name_2, file_code, taxon) %>%
    summarise(rel_abund = sum(rel_abund), #add Others taxon together
              .groups = "drop") %>%
    group_by(cycle_name_2, taxon) %>%
   summarise(rel_abund = median(rel_abund)) %>% # for median reads per taxon per cycle
   left_join(class_df , by=c("taxon" = "order")) %>% # get class levels for sp.
       filter(taxon != "Other") %>%
     ggplot(aes(x=cycle_name_2, y = taxon, fill =rel_abund)) +
    geom_tile() +
  # geom_text(aes(label=format(round(rel_abund,1), nsmall = 1)), size=3)+ # for number labels
  scale_fill_distiller(palette = "Purples",
                       name = "Median<br>Relative<br>Abund (%)",
                       expand = c(0,0),
                       breaks = c(0, 0.1, 1, 10,100),
                       na.value = "white",
                       direction = 1,
                       trans = "log",
                       # breaks = c(0, 0.1, 1, 10, 100),
                       # limits = c(NA,100)
                       )+
    theme_classic()+
    theme(text=element_text(size=10),
          strip.text.y.left = element_text(angle = 0),
          panel.spacing = unit(0,"lines"),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          strip.background = element_blank(),
          axis.line = element_blank(),
          axis.ticks = element_blank(),
          legend.title = element_markdown(),
          strip.placement = "outside") +
    scale_x_discrete(position = "top") +
    xlab(NULL)+
    ylab(NULL) +
    facet_grid(class~., scales = "free_y", space = "free", switch = "y") +ggtitle("Nanoeukaryotes - order (log scale)")
    
p3
# pdf("nano order_heatmap.pdf", height=4, width=7) ; plot(p3) ; dev.off()
```

## Figure S13 - Species - median  select 5% minimum
```{r}
taxon_rel_abund <- otu_rel_abund %>%
    filter(level == "species") %>% #choose level here 
    group_by(cycle_name_2, file_code, taxon) %>%
    summarise(rel_abund = 100*sum(rel_abund)) %>%
    pivot_wider(names_from = taxon, values_from = rel_abund) %>% #from here is to replace any file_code with NA as 0
    replace(., is.na(.), 0) %>%
    pivot_longer(cols= !cycle_name_2:file_code,
                 names_to = "taxon",
                 values_to = "rel_abund")

taxon_pool <- taxon_rel_abund %>%
    group_by(cycle_name_2, taxon) %>%
    summarise(median = mean(rel_abund)) %>% #calculate median proportion of reads per sp per cycle
    group_by(taxon) %>%
    summarise(pool = max(median) <= 5, #species with a max median proportion of reads in any cycle < 5% is under Others
              .groups = "drop")

#get class name for facet grid grouping with class
class_df <- long_nano_1 %>%
    select(class, order, species) %>%
    distinct()

#heatmap
p4 <- inner_join(taxon_rel_abund, taxon_pool, by="taxon") %>%
    mutate(taxon=if_else(pool, "Other", taxon)) %>% 
    group_by(cycle_name_2, file_code, taxon) %>%
    summarise(rel_abund = sum(rel_abund), #add Others taxon together
              .groups = "drop") %>%
    group_by(cycle_name_2, taxon) %>%
   summarise(rel_abund = mean(rel_abund)) %>% # for median reads per taxon per cycle
   left_join(class_df , by=c("taxon" = "species")) %>% # get class levels for sp.
     ggplot(aes(x=cycle_name_2, y = taxon, fill =rel_abund)) +
    geom_tile() +
  geom_text(aes(label=format(round(rel_abund,1), nsmall = 1)), size=3)+
  scale_fill_distiller(name = "Median<br>Relative<br>Abund (%)",
                       palette = "Purples",
                       expand = c(0,0),
                      na.value = "white",
                      direction = 1,
                      trans = "log")+
  theme_classic()+
    theme(text=element_text(size=10),
              strip.text.y.left = element_text(angle = 0),
              panel.spacing.x = unit(0,"lines"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              strip.background = element_blank(),
          axis.line = element_blank(),
          axis.ticks = element_blank(),
          legend.title = element_markdown(),
          strip.placement = "outside") +
  scale_x_discrete(position = "top") +
  xlab(NULL)+
  ylab(NULL) +
    facet_grid(order~., scales = "free_y", space = "free", switch = "y") 
    
p4
# pdf("nano_species_heatmap.pdf", height=7, width=8) ; plot(p4) ; dev.off()
```

# Sorted syn
## Read data
```{r}
source(here("3_Phyloseq output and figures", "init_files", "read_petb_2.0.R"))


long_sorted <- long_sorted %>%
  mutate(cycle_name = recode(Cycle., "1" = "SA-Sc",
                                    "2" = "SA1",
                                    "3" = "ST1",
                                    "4" = "ST2",
                                    "5" = "SA2")) %>%
  mutate(cycle_name = fct_relevel(cycle_name, "ST1", "ST2", "SA-Sc", "SA1", "SA2"))
```
## prepare data frame
```{r}

long_sorted_1 <- long_sorted %>%
     filter(vial_type=="light")


# pivot longer dataframe
otu_rel_abund <- long_sorted_1 %>%
    group_by(file_code) %>%
    mutate(rel_abund = n_reads / sum(n_reads)) %>%
    ungroup() %>%
    select(-n_reads) %>%
    pivot_longer(cols = asv_code:Subclade,
                 names_to = "level",
                 values_to = "taxon") 
```

## summary for results 
```{r}
taxon_rel_abund <- otu_rel_abund %>%
    filter(level == "Clade") %>% #choose level here 
    group_by(cycle_name, file_code, taxon) %>%
    summarise(rel_abund = 100*sum(rel_abund)) %>%
    pivot_wider(names_from = taxon, values_from = rel_abund) %>% #from here is to replace any file_code with NA as 0
    replace(., is.na(.), 0) %>%
    pivot_longer(cols= !cycle_name:file_code,
                 names_to = "taxon",
                 values_to = "rel_abund")


# summary <- taxon_rel_abund %>%
#   # filter(cycle_name %in% c("SA2", "SA3")) %>%
#   group_by(taxon) %>%
#   summarise(median = median(rel_abund),
#             mean = mean(rel_abund),
#             max = mean(rel_abund),
#             min = min(rel_abund))
# summary
```


## subclade - select above 5%
```{r}
taxon_rel_abund <- otu_rel_abund %>%
    filter(level == "Subclade") %>% #choose level here 
    group_by(water_mass, cycle_name, file_code, taxon) %>%
    summarise(rel_abund = 100*sum(rel_abund)) %>%
    pivot_wider(names_from = taxon, values_from = rel_abund) %>% #from here is to replace any file_code with NA as 0
    replace(., is.na(.), 0) %>%
    pivot_longer(cols= !water_mass:file_code,
                 names_to = "taxon",
                 values_to = "rel_abund")

## summary for results
# summary <- taxon_rel_abund %>%
#   group_by(taxon, water_mass) %>%
#   summarise(median = median(rel_abund),
#             mean = mean(rel_abund))
# summary

taxon_pool <- taxon_rel_abund %>%
    group_by(cycle_name, taxon) %>%
    summarise(median = median(rel_abund)) %>% #calculate median proportion of reads per sp per cycle
    group_by(taxon) %>%
    summarise(pool = max(median) <= 5, #species with a max median proportion of reads in any cycle < 5% is under Others
              median = max(median),
              .groups = "drop")

#get class name for facet grid grouping with class
df <- long_sorted_1 %>%
    select(Clade, Subclade) %>%
    distinct()


#heatmap
p5 <- inner_join(taxon_rel_abund, taxon_pool, by="taxon") %>%
    mutate(taxon=if_else(pool, "Other", taxon)) %>% 
    group_by(cycle_name, file_code, taxon) %>%
    summarise(rel_abund = sum(rel_abund), #add Others taxon together
              .groups = "drop") %>%
    group_by(cycle_name, taxon) %>%
   summarise(rel_abund = median(rel_abund)) %>% # for median reads per taxon per cycle
   left_join(df , by=c("taxon" = "Subclade")) %>% # get class levels for sp.
   filter(taxon != "Other") %>%
    ggplot(aes(x=cycle_name, y = taxon, fill =rel_abund)) +
    geom_tile() +
    # geom_text(aes(label=format(round(rel_abund,1), nsmall = 1)), size=2)+
    scale_fill_distiller(palette = "Oranges",
                        name = "Median<br>Relative<br>Abund (%)",
                       expand = c(0,0),
                       na.value = "white",
                       direction = 1,
                       trans = "log",
                       breaks = c(0, 0.1, 1, 10, 100),
                       limits = c(NA,100))+
    theme_classic()+
    theme(text=element_text(size=10),
              strip.text.y.left = element_text(angle = 0),
              panel.spacing = unit(0,"lines"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              strip.background = element_blank(),
              axis.line = element_blank(),
              axis.ticks = element_blank(),
              legend.title = element_markdown(),
              strip.placement = "outside") +
    scale_x_discrete(position = "top") +
    xlab(NULL)+
    ylab(NULL) +
    facet_grid(Clade~., scales = "free_y", space = "free", switch = "y") +
    ggtitle("Syn - subclade (log scale)")
    
p5
# pdf("syn_heatmap.pdf", height=2, width=5) ; plot(p) ; dev.off()
```
## Descriptive stats
```{r}
#no. of samples per cycle
sample_no <- taxon_rel_abund %>%
  select(cycle_name_2, file_code) %>%
  distinct() %>%
  group_by(cycle_name_2) %>%
  summarise(n = n())
sample_no
```


# Figure 8 - Heatmap
```{r}
library(patchwork)
plot <- (p1 |plot_spacer() | p5) / (p2|p3) + plot_annotation(tag_levels = 'A') + plot_layout( heights = c(1, 1.75))

plot
pdf("heatmap_test v2.pdf", height=6, width=12) ; plot(plot) ; dev.off()

library(patchwork)
plot <- (p1 |p2) / (p3|p5) + plot_annotation(tag_levels = 'A')# + plot_layout( heights = c(1, 1.75))

plot
pdf("heatmap_tests.pdf", height=6, width=12) ; plot(plot) ; dev.off()

```
