---
title: "NMDS_PERMANOVA"
author: "Denise Ong"
date: "11/5/2022"
output: html_document
---
Community analysis - alpha diversity, NMDS, PERMANOVA, beta-disper

# Initialize
```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(here)
library(vegan)
source(here("phyloseq", "init_files", "init.R")) #for libraries
source(here("phyloseq", "init_files", "init_markdown.R")) #for markdown libraries
source(here("phyloseq", "init_files", "define_functions.R")) #functions to 1.transform to dataframe, 2.normalise and 3.create treemaps 
source(here("phyloseq", "trophic_mode", "trophic_function.R"))  #function to assign trophic level for phyloseq objects

#Colours for cycle
colours_cycle <- readxl::read_excel(here("phyloseq", "init_files", "colours_cycle.xlsx"))
colours_cycle <- structure(colours_cycle$colour,.Names=colours_cycle$cycle)

set.seed(1)
```

```{r}
# convert the sample_data() within a phyloseq object to a vegan compatible data object
pssd2veg <- function(physeq) {
  sd <- sample_data(physeq)
  return(as(sd,"data.frame"))
}

# convert the otu_table() within a phyloseq object to a vegan compatible data object
psotu2veg <- function(physeq) {
  OTU <- otu_table(physeq)
  if (taxa_are_rows(OTU)) {
    OTU <- t(OTU)
  }
  return(as(OTU, "matrix"))
}
```

# Sorted pico vs nano - 18SV4
## load data
CTD, sorted pico and sorted nano
filter out heterotrophs
```{r}
source(here("phyloseq", "init_files", "read_18s_filt_sort.R"))

ps_filt_sort <- phyloseq_assign_trophic(ps_filt_sort) %>%
    subset_taxa(trophic_mode_2 != "heterotrophic") #remove heterotrophs

#select only sorted samples, light incubation
sample <- data.frame(ps_filt_sort@sam_data) %>%
    filter(sample_type %in% c("Pico", "Nano")) %>%
    filter(vial %in% c("A", "B", "C"))

otu <- psotu2veg(ps_filt_sort)

#sample list of filt samples
sample_list <- sample %>%
  select(sample_name)

# subset removed groups and samples from the plot
otu_sub <- otu %>%
  as.data.frame() 
otu_sub$sample_name <-rownames(otu_sub)

otu_sub <- otu_sub %>% 
  inner_join(., sample_list) %>% 
  filter(!(sample_name %in% c("nano-12", "pico-15", "pico-01-30"))) #remove few sorted samples

rownames(otu_sub) <- otu_sub$sample_name

otu_sub <- otu_sub %>%
  select(-sample_name) %>%
  as.matrix()

```

### checks before plotting - ok, ignore for now.
```{r}
#to check for every sample number of reads and if all asvs have reads -  ok, ignore code now.

otu_long <- otu_sub %>%
  as.data.frame() 
otu_long$file_code <-rownames(otu_long)
otu_long <- otu_long %>%
    pivot_longer(cols = !file_code, names_to = "otu", values_to = "reads") %>%
  # group_by(file_code)  %>%
  #  summarise(sum = sum(reads)) %>%
    group_by(otu) %>%
    summarise(sum = sum(reads))

```

## NMDS
```{r}
set.seed(1)
dist <- avgdist(otu_sub, sample = 8000, dmethod = "bray") #rarify to control for uneven number of reads per sample.
nmds_sub<- metaMDS(dist)

nmds_sample_sub <- scores(nmds_sub) %>%
    as.tibble(rownames="sample_name") %>%
    inner_join(., sample, by = "sample_name")

goodness(nmds_sub)
stressplot(nmds_sub)

nmds_sub$stress

#plot overall
nmds_euks <- nmds_sample_sub %>%
    mutate(cycle_name = fct_relevel(cycle_name_2, "ST1", "ST2", "SA1","SA2","SA3" )) %>%
    mutate(water_mass = recode(cycle_name, "ST1" = "ST",
                                           "ST2" = "ST",
                                           "SA1" = "SA",
                                           "SA2" = "SA",
                                           "SA3" = "SA")) %>%
    ggplot(aes(x=NMDS1, y= NMDS2, shape = sample_type)) +
    geom_point(aes(colour = cycle_name)) +
    stat_ellipse(aes(colour = water_mass))+
    scale_colour_manual(values = colours_cycle) +
    theme_classic() +
    theme(legend.position = "top")

nmds_euks

# pdf("nmds_ellipse.pdf", height=6, width=6) ; plot(p) ; dev.off()


```

## PERMANOVA
```{r}
dist_df <- as.data.frame(as.matrix(dist))
dist_df$sample_name <-rownames(dist_df) #row names to column
rownames(dist_df) <- c() #remove row names
meta_dist <-  inner_join(sample, dist_df, by = "sample_name") #join with the metdata
all_dist <- dist_df %>%
    select(all_of(.[["sample_name"]])) %>%
    as.dist() #something to change to a distance matrix for adonis

all_test <- adonis(all_dist~sample_type, data=meta_dist)

all_test #significant difference between the three sample types

meta_dist %>% group_by(sample_type, sample) %>% count(sample)

bd <- betadisper(all_dist, meta_dist$sample_type)
anova(bd)
permutest(bd)
meta_dist %>% count(sample_type)


```

## Alpha diversity
```{r}

shared <- ps_filt_sort %>%
  phyloseq_normalize_median() %>%
  phyloseq_transform_to_long() %>%
  filter(sample_type %in% c("Pico", "Nano")) %>%
  filter(vial %in% c("A", "B", "C")) %>%
  filter(!(sample_name %in% c("nano-12", "pico-15", "pico-01-30"))) %>%
  select(asv_code, sample_name, n_reads) %>%
  group_by(sample_name) %>%
  mutate(total = sum(n_reads)) %>%
  filter(total != 0) %>%
  ungroup() %>%
  select(-total)

rand <-  shared %>%
  uncount(n_reads) %>%
  mutate(sample_name = sample(sample_name)) %>%
  count(asv_code, sample_name, sample_name="value")

#### functions to calculate diversity ####
richness <- function(x) {
  sum(x>0)
}

shannon <- function(x){
  rabund <- x[x>0]/sum(x)
  -sum(rabund*log(rabund))
}

simpson<- function(x){
  n<- sum(x)
  sum(x*(x-1)/(n*(n-1)))
}

#### ####

shared %>%
  group_by(sample_name) %>%
  summarise(sobs = richness(n_reads),
            shannon = shannon(n_reads),
            simpson = simpson(n_reads),
            invsimpson = 1/simpson,
            n = sum(n_reads)) %>%
  pivot_longer(cols = c(sobs, shannon, invsimpson, simpson),
               names_to = "metric") %>%
  ggplot(aes(x=n, y=value)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~metric,nrow=4, scales="free_y")

div <- shared %>%
  group_by(sample_name) %>%
  summarise(sobs = specnumber(n_reads),
            shannon = diversity(n_reads, index = "shannon"),
            simpson = diversity(n_reads, index = "simpson"),
            invsimpson = 1/simpson,
            n = sum(n_reads))
div_meta <-  inner_join(div, sample)

summary <- div_meta %>%
    group_by(sample_type, water_mass) %>%
    summarise(shannon_mean = round(mean(shannon), digits = 2),
              shannon_sd = round(sd(shannon), digits = 2),
              simpson_mean = round(mean(simpson), digits = 2),
              simpson_sd = round(sd(simpson), digits = 2),
              asv_mean = round(mean(sobs), digits = 2),
              asv_sd = round(sd(sobs), digits = 2),
              n=n()
              ) %>%
    unite("shannon", shannon_mean, shannon_sd, sep = " ± ") %>%
    unite("simpson", simpson_mean, simpson_sd, sep = " ± ") %>%
    unite("asv", asv_mean, asv_sd, sep = " ± ") 
summary

```

### Statistical tests for alpha diversity

```{r}
# Pico vs nano

for (i in c("shannon", "simpson", "sobs")) {
t_test_df<- div_meta %>%
  select(water_mass, i, sample_type) %>%
  rename(div = i)

summary <- t_test_df %>%
  group_by(sample_type) %>%
  summarise(count = n(),
    mean = mean(div, na.rm = TRUE),
    sd = sd(div, na.rm = TRUE))
print(summary)
 p <- t_test_df %>%
   ggplot(aes(x=sample_type, y= div)) + geom_point() +geom_boxplot() +ggtitle(i)
 print(p)

#Because p value <0.05, data is not normally distributed. Use the wilcoxon test
p<-with(t_test_df, shapiro.test(div[sample_type == "Pico"]))
print(p)
p<-with(t_test_df, shapiro.test(div[sample_type == "Nano"]))
print(p)
# res <- t.test(div ~ sample_type, data = t_test_df,
#                    var.equal = TRUE)
res <- wilcox.test(div ~ sample_type, data = t_test_df,
                   exact = FALSE)
print(res)
}
```


```{r}
# Within each group, check between water mass
library("ggpubr")

for (i in c("shannon", "simpson", "sobs")) {
t_test_df<- div_meta %>%
  filter(sample_type == "Nano") %>%
  # filter(sample_type == "Pico") %>%
  select(water_mass, i, sample_type) %>%
  rename(div = i)

summary <- t_test_df %>%
  group_by(water_mass) %>%
  summarise(count = n(),
    mean = mean(div, na.rm = TRUE),
    sd = sd(div, na.rm = TRUE))
print(summary)
 p <- t_test_df %>%
   ggplot(aes(x=water_mass, y= div)) + geom_point() +geom_boxplot() +ggtitle(i)
 print(p)

#Because p value <0.05, data is not normally distributed. Use the wilcoxon test
p<-with(t_test_df, shapiro.test(div[water_mass == "ST"]))
print(p)
p<-with(t_test_df, shapiro.test(div[water_mass == "SA"]))
print(p)
# res <- t.test(div ~ water_mass, data = t_test_df,
#                    var.equal = TRUE)
res <- wilcox.test(div ~ water_mass, data = t_test_df,
                   exact = FALSE)
print(res)
}
```


# Sorted Syn - petB
```{r}
source(here("phyloseq", "init_files", "read_petb_2.0.R"))

#select only sorted samples, light incubation
sample <- data.frame(ps_sorted@sam_data) %>%
    filter(vial %in% c("A", "B", "C"))

otu <- psotu2veg(ps_sorted)


#sample list of filt samples
sample_list <- sample %>%
  select(sample_name)

# subset removed groups and samples from the plot
otu_sub <- otu %>%
  as.data.frame() 
otu_sub$sample_name <-rownames(otu_sub)

otu_sub <- otu_sub %>% 
  inner_join(., sample_list) #select only corresponding depths for filt samples

rownames(otu_sub) <- otu_sub$sample_name

otu_sub <- otu_sub %>%
  select(-sample_name) %>%
  as.matrix()

```

##checks before plotting - ok, ignore for now.
```{r}
#to check for every sample number of reads and if all asvs have reads -  ok, ignore code now.

otu_long <- otu_sub %>%
  as.data.frame() 
otu_long$file_code <-rownames(otu_long)
otu_long <- otu_long %>%
    pivot_longer(cols = !file_code, names_to = "otu", values_to = "reads") %>%
   # group_by(file_code)  %>%
  #  summarise(sum = sum(reads)) %>%
    group_by(otu) %>%
    summarise(sum = sum(reads))

```


## NMDS
```{r}
set.seed(1)
dist <- avgdist(otu_sub, sample = 8000, dmethod = "bray") #rarify to control for uneven number of reads per sample.
nmds_sub<- metaMDS(dist)

nmds_sample_sub <- scores(nmds_sub) %>%
    as.tibble(rownames="sample_name") %>%
    inner_join(., sample, by = "sample_name")

goodness(nmds_sub)
stressplot(nmds_sub)

nmds_sub$stress

#plot overall
nmds_syn <- nmds_sample_sub %>%
    mutate(cycle_name = fct_relevel(cycle_name, "ST1", "ST2", "SA1","SA2","SA3" )) %>%
    ggplot(aes(x=NMDS1, y= NMDS2)) +
    geom_point(aes(colour = cycle_name), shape=15) +
  stat_ellipse(aes(colour = water_mass))+
    scale_colour_manual(values = colours_cycle) +
  theme_classic() +
  theme(legend.position = "top")

nmds_syn

# pdf("nmds_ellipse.pdf", height=6, width=6) ; plot(p) ; dev.off()


```


## PERMANOVA
```{r}
dist_df <- as.data.frame(as.matrix(dist))
dist_df$sample_name <-rownames(dist_df) #row names to column
rownames(dist_df) <- c() #remove row names
meta_dist <-  inner_join(sample, dist_df, by = "sample_name")  #join with the metdata

all_dist <- dist_df %>%
    select(all_of(.[["sample_name"]])) %>%
    as.dist() #something to change to a distance matrix for adonis

all_test <- adonis(all_dist~SAMPLE*water_mass, 
                   # strata = meta_dist$ctd_depth,
                   data=meta_dist)

all_test 

bd <- betadisper(all_dist, meta_dist$water_mass)
anova(bd)
permutest(bd)
meta_dist %>% count(water_mass)

```


## Alpha diversity
```{r}

shared <- ps_sorted %>%
  phyloseq_normalize_median() %>%
  phyloseq_transform_to_long() %>%
  select(asv_code, sample_name, n_reads) %>%
  group_by(sample_name) %>%
  mutate(total = sum(n_reads)) %>%
  filter(total != 0) %>%
  ungroup() %>%
  select(-total)

rand <-  shared %>%
  uncount(n_reads) %>%
  mutate(sample_name = sample(sample_name)) %>%
  count(asv_code, sample_name, sample_name="value")

#functions to calculate diversity
richness <- function(x) {
  sum(x>0)
}

shannon <- function(x){
  rabund <- x[x>0]/sum(x)
  -sum(rabund*log(rabund))
}

simpson<- function(x){
  n<- sum(x)
  sum(x*(x-1)/(n*(n-1)))
}

shared %>%
  group_by(sample_name) %>%
  summarise(sobs = richness(n_reads),
            shannon = shannon(n_reads),
            simpson = simpson(n_reads),
            invsimpson = 1/simpson,
            n = sum(n_reads)) %>%
  pivot_longer(cols = c(sobs, shannon, invsimpson, simpson),
               names_to = "metric") %>%
  ggplot(aes(x=n, y=value)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~metric,nrow=4, scales="free_y")

div <- shared %>%
  group_by(sample_name) %>%
  summarise(sobs = specnumber(n_reads),
            shannon = diversity(n_reads, index = "shannon"),
            simpson = diversity(n_reads, index = "simpson"),
            invsimpson = 1/simpson,
            n = sum(n_reads))
div_meta <-  inner_join(div, sample)

div_meta
```

### Statistical tests for alpha diversity
```{r}
library("ggpubr")

for (i in c("shannon", "simpson", "sobs")) {
t_test_df<- div_meta %>%
  # filter(sample_type == "Nano") %>%
  select(water_mass, i) %>%
  rename(div = i)

summary <- t_test_df %>%
  group_by(water_mass) %>%
  summarise(count = n(),
    mean = mean(div, na.rm = TRUE),
    sd = sd(div, na.rm = TRUE))
print(summary)
 p <- t_test_df %>%
   ggplot(aes(x=water_mass, y= div)) + geom_point() +geom_boxplot() +ggtitle(i)
 print(p)

#Because p value <0.05, data is not normally distributed. Use the wilcoxon test
p<-with(t_test_df, shapiro.test(div[water_mass == "ST"]))
print(p)
p<-with(t_test_df, shapiro.test(div[water_mass == "SA"]))
print(p)
# res <- t.test(div ~ water_mass, data = t_test_df,
#                    var.equal = TRUE)
res <- wilcox.test(div ~ water_mass, data = t_test_df,
                   exact = FALSE)
print(res)
}
```

# Table S5 - Alpha diversity indices
```{r}
#summary table
syn_summary_wm <- div_meta %>%
    group_by(water_mass) %>%
    summarise(shannon_mean = round(mean(shannon), digits = 2),
              shannon_sd = round(sd(shannon), digits = 2),
              simpson_mean = round(mean(simpson), digits = 2),
              simpson_sd = round(sd(simpson), digits = 2),
              asv_mean = round(mean(sobs), digits = 2),
              asv_sd = round(sd(sobs), digits = 2),
              n=n()
              ) %>%
    unite("shannon", shannon_mean, shannon_sd, sep = " ± ") %>%
    unite("simpson", simpson_mean, simpson_sd, sep = " ± ") %>%
    unite("asv", asv_mean, asv_sd, sep = " ± ") %>%
  mutate(sample_type = "Syn")

syn_summary_wm

summary_all <- full_join(summary, syn_summary_wm) %>%
  arrange(factor(water_mass, levels = c("ST", "SA"))) %>%
  arrange(factor(sample_type, levels = c("Syn", "Pico", "Nano")))
summary_all

writexl::write_xlsx(summary_all, "phyloseq/paper_c14/figs_paper/sorted_alphadiv.xlsx")
library(xtable)
print(xtable(summary_all),
      comment = FALSE,
      include.rownames=FALSE)
```

# Figure 6- NMDS 
```{r}
library(patchwork)
p<-nmds_syn+nmds_euks + plot_layout(guides = 'collect') +plot_annotation(tag_levels = 'A') &theme(legend.position = "top")
p
pdf("nmds_all.pdf", height=5.5, width=8) ; plot(p) ; dev.off()
```
