---
title: "18S filt vs sort"
author: "Denise Ong"
date: "10/27/2022"
output: html_document
---
Supplementary figure S14 for filtered samples.

# Initialize
```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(here)
here()
source(here("3_Phyloseq output and figures", "init_files", "init.R")) #for libraries
source(here("3_Phyloseq output and figures", "init_files", "init_markdown.R")) #for markdown libraries
source(here("3_Phyloseq output and figures", "init_files", "define_functions.R")) #functions to 1.transform to dataframe, 2.normalise and 3.create treemaps
source(here("3_Phyloseq output and figures", "trophic_mode", "trophic_function.R"))  #function to assign trophic level for phyloseq objects
source(here("3_Phyloseq output and figures", "init_files", "colours_18sv4.R")) # predetermined colours
source(here("3_Phyloseq output and figures", "init_files", "colours_petb.R")) # predetermined colours
# source(here("3_Phyloseq output and figures", "init_files", "colours_random.R")) #random colours
```

# Eukaryotes 18SV4 
## read data
```{r}
source(here("3_Phyloseq output and figures", "init_files", "read_18s_filt_sort.R"))

filt_samples_list <- junk %>%
  filter(sample_type == "filt") %>%
  filter(depth_category %in% c("2", "5")) %>% # select surface and DCM depths only.
  filter(station != "108")
# writexl::write_xlsx(filt_samples_list, "filt_samples_list.xlsx")

ps_filt_sort <- phyloseq_assign_trophic(ps_filt_sort) %>%
    subset_taxa(trophic_mode_2 != "heterotrophic") #remove heterotrophs
# Do normalizations and transformation
ps_filt_sort <- phyloseq_normalize_median(ps_filt_sort)
long_filt_sort <- phyloseq_transform_to_long(ps_filt_sort) %>%
  mutate(cycle_name_2 = dplyr::recode(cycle_name_2, "SA1" = "SA-Sc")) %>%
  mutate(cycle_name_2 = dplyr::recode(cycle_name_2, "SA2" = "SA1")) %>%
  mutate(cycle_name_2 = dplyr::recode(cycle_name_2, "SA3" = "SA2")) %>%
  mutate(cycle_name_2 = fct_relevel(cycle_name_2, "ST1", "ST2", "SA-Sc", "SA1", "SA2")) %>%
  mutate(sample = fct_relevel(sample, "SUR", "DCM"))
```


check overlapping samples
```{r}
sample <- data.frame(ps_filt_sort@sam_data) %>%
    select(ctd_cast, station, depth, sample_type, cycle_name_2) %>%
    distinct() %>%
    pivot_wider(names_from = depth, values_from = sample_type)
sample
```

## check photosynthetic in filt not included in sorted
```{r, eval=FALSE}
long_sorted_asvhash <- long_filt_sort %>%
    filter(sample_type != "filt") %>%
    select(asv_code) 

long_filt<- long_filt_sort %>%
    filter(sample_type == "filt")

long_filt_anti <- anti_join(long_filt, long_sorted_asvhash) %>%
  distinct() %>%
  # mutate(in_sorted = "no") %>%
    mutate(in_sorted = case_when(class == "Dinophyceae" ~ "dinophyceae",
                               TRUE~ "no"))  %>%
  filter(class != "Syndiniales") %>%
  filter(species != "Dinophyceae_XXX_sp.") %>%
  filter(trophic_mode == "photosynthetic") 

long_filt_sort <- inner_join(long_filt, long_sorted_asvhash) %>%
    distinct() %>%
  mutate(in_sorted = "yes")


long_filt_sort_join <- full_join(long_filt_anti, long_filt_sort) %>%
  filter(class != "Syndiniales") %>%
  filter(species != "Dinophyceae_XXX_sp.") %>%
  filter(trophic_mode == "photosynthetic") 

# check proportion of photosynthetic not in sorted
p  <- long_filt_sort_join %>%
  ggplot(aes(x=station, y=n_reads, fill=in_sorted, colour = in_sorted)) +
  geom_bar(position="fill", stat="identity", size=0.5) + 
        theme(text=element_text(size=10),
              axis.text.x = element_text(angle = 45, hjust = 1))+
        facet_grid(.~cycle_name_2, scales = "free_x") 
p

# check proportion of dinophyceae
p  <- long_filt_anti %>%
  ggplot(aes(x=station, y=n_reads, fill=trophic_mode_2, colour = trophic_mode_2)) +
  geom_bar(position="fill", stat="identity", size=0.5) + 
        theme(text=element_text(size=10),
              axis.text.x = element_text(angle = 45, hjust = 1))+
        facet_grid(.~cycle_name_2, scales = "free_x") 
p

```

## Pico vs filt
```{r}
# select only ASVs which are in sorted light incubated Pico samples 
long_pico_asvhash <- long_filt_sort %>%
    filter(sample_type == "Pico") %>%
    filter(vial %in% c("A", "B", "C")) %>%
    select(asv_code)

long_pico <- long_filt_sort %>%
    filter(sample_type == "Pico") %>%
    filter(class != "Dinophyceae") #remove Dinophyceae

long_filt<- long_filt_sort %>%
    filter(sample_type == "filt") %>%
    filter(class != "Dinophyceae")

long_filt_pico <- inner_join(long_filt, long_pico_asvhash) %>%
    distinct()
```

to calculate relative abundance in filtered sample
```{r}
# pivot longer dataframe
otu_rel_abund <- long_filt_pico %>%
    filter(class != "Dinophyceae")  %>% # remove dinophyceae reads
    filter(species != "Gephyrocapsa_oceanica") %>%
    group_by(file_code) %>%
    mutate(rel_abund = n_reads / sum(n_reads)) %>%
    ungroup() %>%
    select(-n_reads) %>%
    pivot_longer(cols = asv_code:species,
                 names_to = "level",
                 values_to = "taxon")  %>%
    filter(depth_category %in% c("2", "5")) %>% # select surface and DCM depths only.
    filter(station != "108") %>% # remove station 108 because it is in SA1-b
    filter(station != "371")  %>% # remove because SA2 only has one incubation in the same filt sample
    mutate(depth_cat_2 = case_when(depth_category == '2' ~ "SUR",
                                    depth_category == "5" ~ "DCM",
                                    TRUE ~ depth_category)) 
taxon_rel_abund <- otu_rel_abund %>%
    filter(level == "class") %>% #choose level here 
    group_by(water_mass, cycle_name_2, station, depth_cat_2, file_code, taxon) %>%
    summarise(rel_abund = 100*sum(rel_abund)) %>%
    pivot_wider(names_from = taxon, values_from = rel_abund, values_fill = 0) %>% #from here is to replace any file_code with NA as 0\
    # select(water_mass:file_code, Bacillariophyta:Trebouxiophyceae) %>%
    pivot_longer(cols= !water_mass:file_code,
                 names_to = "taxon",
                 values_to = "rel_abund")


taxon_rel_abund <- otu_rel_abund %>%
    filter(level == "class") %>% #choose level here 
    group_by(water_mass, cycle_name_2, station, depth_cat_2, file_code, taxon) %>%
    summarise(rel_abund = 100*sum(rel_abund)) %>%
    pivot_wider(names_from = taxon, values_from = rel_abund, values_fill = 0) %>% #from here is to replace any file_code with NA as 0\
    # select(water_mass:file_code, Bacillariophyta:Trebouxiophyceae) %>%
    pivot_longer(cols= !water_mass:file_code,
                 names_to = "taxon",
                 values_to = "rel_abund")

check <- taxon_rel_abund %>%
    group_by(taxon) %>%
    summarise(median= median(rel_abund))
#summary for supplementary
summary_filt_pico <- taxon_rel_abund %>%
    group_by(taxon, cycle_name_2) %>%
    summarise(median = round(median(rel_abund),2),
              IQR = round(IQR(rel_abund),2)) %>%
    unite("median_IQR", median:IQR, sep = " (") %>%
    mutate(edit = ")") %>%
    unite("median_IQR", median_IQR:edit, sep = "") %>%
    mutate(type = "filt") %>%
    select(taxon, cycle_name_2, type, median_IQR) %>%
    unite("cycle_name_2", cycle_name_2:type, sep="_") %>%
    pivot_wider(names_from = taxon, values_from = median_IQR) %>%
    t() %>%
    as.data.frame() %>%
    rownames_to_column(var = "Class")
```
to calculate relative abundance in sorted sample
```{r}
otu_rel_abund <- long_pico %>%
    filter(class != "Dinophyceae") %>% #remove Dinophyceae
    filter(species != "Gephyrocapsa_oceanica") %>%
    filter(vial %in% c("A", "B", "C")) %>%
    group_by(file_code) %>%
    mutate(rel_abund = n_reads / sum(n_reads)) %>%
    ungroup() %>%
    select(-n_reads) %>%
    pivot_longer(cols = asv_code:species,
                 names_to = "level",
                 values_to = "taxon") 

taxon_rel_abund <- otu_rel_abund %>%
    filter(level == "class") %>% #choose level here 
    group_by(water_mass, cycle_name_2, station,file_code, taxon) %>%
    summarise(rel_abund = 100*sum(rel_abund)) %>%
    pivot_wider(names_from = taxon, values_from = rel_abund, values_fill = 0) %>% #from here is to replace any file_code with NA as 0\
    # select(water_mass:file_code, Chryspophyceae:`MOCH-2`) %>%
    pivot_longer(cols= !water_mass:file_code,
                 names_to = "taxon",
                 values_to = "rel_abund")

check <- taxon_rel_abund %>%
    group_by(taxon) %>%
    summarise(median= median(rel_abund))

#summary for supplementary
summary_sort_pico <- taxon_rel_abund %>%
    group_by(taxon, cycle_name_2) %>%
    summarise(median = round(median(rel_abund),2),
              IQR = round(IQR(rel_abund),2)) %>%
    unite("median_IQR", median:IQR, sep = " (") %>%
    mutate(edit = ")") %>%
    unite("median_IQR", median_IQR:edit, sep = "") %>%
    mutate(type = "sort") %>%
    select(taxon, cycle_name_2, type, median_IQR) %>%
    unite("cycle_name_2", cycle_name_2:type, sep="_") %>%
    pivot_wider(names_from = taxon, values_from = median_IQR) %>%
    t() %>%
    as.data.frame() %>%
    rownames_to_column(var = "Class") %>%
    full_join(check, by = join_by(Class == taxon)) #order by median abundance


summary_pico <- full_join(summary_sort_pico, summary_filt_pico, by = join_by(Class))
writexl::write_xlsx(summary_pico, "summary_pico.xlsx")
```


additional checks
```{r}
filt_pico_merge <-full_join(long_pico,long_filt_pico) %>%
  distinct() %>%
    filter(class != "Dinophyceae")

# compare directly sorted vs filtered. 
p <- filt_pico_merge %>%
  filter(sample %in% c("SUR", "DCM")) %>%
  filter(station != "108") %>% 
    filter(station != "371") %>%# remove station 108 because it is in SA1-b
  # unite("cycle_station", cycle_name_2, station, sep = "_") %>%
  mutate(class = fct_lump_n(class, 10, other_level = "Others")) %>%
  ggplot(aes(x=sample_type, y=n_reads, fill=class, colour = class)) + 
        geom_bar(position="fill", stat="identity", size=0.5) + 
        theme(text=element_text(size=10),
              axis.text.x = element_text(angle = 90, hjust = 1))+
        facet_grid(sample~cycle_name_2, scales = "free_x", space = "free") +
        ggtitle("Sorted pico - class level") +
        scale_color_manual(values=colours_18s_class, limits = force) +
        scale_fill_manual(values=colours_18s_class, limits = force) +
  ylab("Proportion of reads")
p
```

## Nano vs filt
```{r}
long_nano_asvhash <- long_filt_sort %>%
    filter(sample_type == "Nano") %>%
    filter(vial %in% c("A", "B", "C")) %>%
    select(asv_code)

long_nano <- long_filt_sort %>%
    filter(sample_type == "Nano") 

long_filt<- long_filt_sort %>%
    filter(sample_type == "filt")

long_filt_nano <- inner_join(long_filt, long_nano_asvhash) %>%
    distinct()
```

to calculate relative abundance and summary for results
```{r}
# pivot longer dataframe
otu_rel_abund <- long_filt_nano %>%
    filter(class!= "Mamiellophyceae") %>%
    filter(class!= "Chloropicophyceae") %>%
    group_by(file_code) %>%
    mutate(rel_abund = n_reads / sum(n_reads)) %>%
    ungroup() %>%
    select(-n_reads) %>%
    pivot_longer(cols = asv_code:species,
                 names_to = "level",
                 values_to = "taxon") %>%
    filter(depth_category %in% c("2", "5")) %>% # select surface and DCM depths only.
    filter(station != "108") %>% # remove station 108 because it is in SA1-b
    filter(station != "371")  %>% # remove because SA2 only has one incubation in the same filt sample
    mutate(depth_cat_2 = case_when(depth_category == '2' ~ "SUR",
                                    depth_category == "5" ~ "DCM",
                                    TRUE ~ depth_category))

taxon_rel_abund <- otu_rel_abund %>%
    filter(level == "class") %>% #choose level here 
    group_by(water_mass, cycle_name_2, station, depth_category, file_code, taxon) %>%
    summarise(rel_abund = 100*sum(rel_abund)) %>%
    pivot_wider(names_from = taxon, values_from = rel_abund, values_fill = 0) %>% #from here is to replace any file_code with NA as 0
    # select(water_mass:file_code, Bacillariophyta:`Prasino-Clade-VIII`) %>%
    pivot_longer(cols= !water_mass:file_code,
                 names_to = "taxon",
                 values_to = "rel_abund")


check <- taxon_rel_abund %>%
    group_by(taxon) %>%
    summarise(median= median(rel_abund))
#summary for supplementary
summary_filt_nano <- taxon_rel_abund %>%
    group_by(taxon, cycle_name_2) %>%
    summarise(median = round(median(rel_abund),2),
              IQR = round(IQR(rel_abund),2)) %>%
    unite("median_IQR", median:IQR, sep = " (") %>%
    mutate(edit = ")") %>%
    unite("median_IQR", median_IQR:edit, sep = "") %>%
    mutate(type = "filt") %>%
    select(taxon, cycle_name_2, type, median_IQR) %>%
    unite("cycle_name_2", cycle_name_2:type, sep="_") %>%
    pivot_wider(names_from = taxon, values_from = median_IQR) %>%
    t() %>%
    as.data.frame() %>%
    rownames_to_column(var = "Class")
```

```{r}
otu_rel_abund <- long_nano %>%
    filter(class!= "Mamiellophyceae") %>%
    filter(class!= "Chloropicophyceae") %>%
    filter(vial %in% c("A", "B", "C")) %>%
    group_by(file_code) %>%
    mutate(rel_abund = n_reads / sum(n_reads)) %>%
    ungroup() %>%
    select(-n_reads) %>%
    pivot_longer(cols = asv_code:species,
                 names_to = "level",
                 values_to = "taxon") 

taxon_rel_abund <- otu_rel_abund %>%
    filter(level == "class") %>% #choose level here 
    group_by(water_mass, cycle_name_2, station,file_code, taxon) %>%
    summarise(rel_abund = 100*sum(rel_abund)) %>%
    pivot_wider(names_from = taxon, values_from = rel_abund, values_fill = 0) %>% #from here is to replace any file_code with NA as 0
    pivot_longer(cols= !water_mass:file_code,
                 names_to = "taxon",
                 values_to = "rel_abund")

check <- taxon_rel_abund %>%
    group_by(taxon) %>%
    summarise(median= median(rel_abund))

#summary for supplementary
summary_sort_nano <- taxon_rel_abund %>%
    group_by(taxon, cycle_name_2) %>%
    summarise(median = round(median(rel_abund),2),
              IQR = round(IQR(rel_abund),2)) %>%
    unite("median_IQR", median:IQR, sep = " (") %>%
    mutate(edit = ")") %>%
    unite("median_IQR", median_IQR:edit, sep = "") %>%
    mutate(type = "sort") %>%
    select(taxon, cycle_name_2, type, median_IQR) %>%
    unite("cycle_name_2", cycle_name_2:type, sep="_") %>%
    pivot_wider(names_from = taxon, values_from = median_IQR) %>%
    t() %>%
    as.data.frame() %>%
    rownames_to_column(var = "Class") %>%
    full_join(check, by = join_by(Class == taxon))


summary_nano <- full_join(summary_sort_nano, summary_filt_nano, by = join_by(Class))
writexl::write_xlsx(summary_nano, "summary_nano.xlsx")
```

additional checks
```{r}
filt_nano_merge <-full_join(long_nano,long_filt_nano) %>%
  distinct() %>%
  filter(class!= "Mamiellophyceae") %>%
  filter(class!= "Chloropicophyceae")

# compare directly sorted vs filtered. 
p <- filt_nano_merge %>%
  filter(sample %in% c("SUR", "DCM")) %>%
  filter(station != "108") %>% # remove station 108 because it is in SA1-b
  filter(station != "371") %>%
  # unite("cycle_station", cycle_name_2, station, sep = "_") %>%
  mutate(class = fct_lump_n(class, 10, other_level = "Others")) %>%
  ggplot(aes(x=sample_type, y=n_reads, fill=class, colour = class)) + 
        geom_bar(position="fill", stat="identity", size=0.5) + 
        theme(text=element_text(size=10),
              axis.text.x = element_text(angle = 90, hjust = 1))+
        facet_grid(sample~cycle_name_2, scales = "free_x", space = "free") +
        ggtitle("Sorted nano - class level") +
        scale_color_manual(values=colours_18s_class, limits = force) +
        scale_fill_manual(values=colours_18s_class, limits = force) +
  ylab("Proportion of reads")
p
```


#Syn - petB
## read data
```{r}
source(here("3_Phyloseq output and figures", "init_files", "read_petb_2.0.R"))

filt_samples_list <- data.frame(ps_filt_sort@sam_data) %>%
  filter(sample_type == "filt") %>%
  # filter(depth_category %in% c("2", "5")) %>% # select surface and DCM depths only.
  filter(station != "108") %>%
  select(station, depth, depth_category)
long_filt_ong <-  full_join(long_filt_ong, filt_samples_list, by = join_by(station, depth)) %>%
    filter(!(is.na(file_code))) 
```

```{r}
long_syn_asvhash <- long_sorted %>%
    filter(vial %in% c("A", "B", "C")) %>%
    select(asv_code)

long_filt_syn <- inner_join(long_filt_ong, long_syn_asvhash) %>%
    distinct()
```

to calculate relative abundance
```{r}
# pivot longer dataframe
otu_rel_abund <- long_filt_syn %>%
  mutate(cycle_name = recode(cycle, "1" = "SA-Sc",
                                    "2" = "SA1",
                                    "3" = "ST1",
                                    "4" = "ST2",
                                    "5" = "SA2")) %>%
    mutate(cycle_name = fct_relevel(cycle_name, "ST1", "ST2", "SA-Sc", "SA1", "SA2")) %>%
    group_by(file_code) %>%
    mutate(rel_abund = n_reads / sum(n_reads)) %>%
    ungroup() %>%
    select(-n_reads) %>%
    pivot_longer(cols = asv_code:Subclade,
                 names_to = "level",
                 values_to = "taxon") %>%
    filter(depth_category %in% c("2", "5")) %>% # select surface and DCM depths only.
    filter(station != "108") %>% # remove station 108 because it is in SA1-b
    filter(station != "371")  %>% # remove because SA2 only has one incubation in the same filt sample
    mutate(depth_cat_2 = case_when(depth_category == '2' ~ "SUR",
                                    depth_category == "5" ~ "DCM",
                                    TRUE ~ depth_category)) 

taxon_rel_abund <- otu_rel_abund %>%
    filter(level == "Subclade") %>% #choose level here 
    group_by(water_mass, cycle_name, station, depth_category, file_code, taxon) %>%
    summarise(rel_abund = 100*sum(rel_abund)) %>%
    pivot_wider(names_from = taxon, values_from = rel_abund, values_fill = 0) %>% #from here is to replace any file_code with NA as 0
    # select(water_mass:file_code, CRD1:IIf) %>%
    pivot_longer(cols= !water_mass:file_code,
                 names_to = "taxon",
                 values_to = "rel_abund") %>%
    arrange(desc(rel_abund))

check <- taxon_rel_abund %>%
    group_by(taxon) %>%
    summarise(median= median(rel_abund))

#summary for supplementary
summary_filt_syn <- taxon_rel_abund %>%
    group_by(taxon, cycle_name) %>%
    summarise(median = round(median(rel_abund),2),
              IQR = round(IQR(rel_abund),2)) %>%
    unite("median_IQR", median:IQR, sep = " (") %>%
    mutate(edit = ")") %>%
    unite("median_IQR", median_IQR:edit, sep = "") %>%
    mutate(type = "filt") %>%
    select(taxon, cycle_name, type, median_IQR) %>%
    unite("cycle_name", cycle_name:type, sep="_") %>%
    pivot_wider(names_from = taxon, values_from = median_IQR) %>%
    t() %>%
    as.data.frame() %>%
    rownames_to_column(var = "Subclade")

```

```{r}
otu_rel_abund <- long_sorted %>%
    filter(vial %in% c("A", "B", "C")) %>%
    group_by(file_code) %>%
    mutate(rel_abund = n_reads / sum(n_reads)) %>%
    ungroup() %>%
    select(-n_reads) %>%
    pivot_longer(cols = asv_code:Subclade,
                 names_to = "level",
                 values_to = "taxon") 

taxon_rel_abund <- otu_rel_abund %>%
    filter(level == "Subclade") %>% #choose level here 
    group_by(water_mass, cycle_name, file_code, taxon) %>%
    summarise(rel_abund = 100*sum(rel_abund)) %>%
    pivot_wider(names_from = taxon, values_from = rel_abund, values_fill = 0) %>% #from here is to replace any file_code with NA as 0
    pivot_longer(cols= !water_mass:file_code,
                 names_to = "taxon",
                 values_to = "rel_abund")

check <- taxon_rel_abund %>%
    group_by(taxon) %>%
    summarise(median= median(rel_abund))

#summary for supplementary
summary_sort_syn <- taxon_rel_abund %>%
    group_by(taxon, cycle_name) %>%
    summarise(median = round(median(rel_abund),2),
              IQR = round(IQR(rel_abund),2)) %>%
    unite("median_IQR", median:IQR, sep = " (") %>%
    mutate(edit = ")") %>%
    unite("median_IQR", median_IQR:edit, sep = "") %>%
    mutate(type = "sort") %>%
    select(taxon, cycle_name, type, median_IQR) %>%
    unite("cycle_name", cycle_name:type, sep="_") %>%
    pivot_wider(names_from = taxon, values_from = median_IQR) %>%
    t() %>%
    as.data.frame() %>%
    rownames_to_column(var = "Subclade") %>%
    full_join(check, by = join_by(Subclade == taxon))


summary_syn <- full_join(summary_sort_syn, summary_filt_syn, by = join_by(Subclade))
writexl::write_xlsx(summary_syn, "summary_syn.xlsx")
```

