---
title: "Group-specific biomass calculations"
author: "Denise Ong"
date: "11/14/2022"
output: html_document
---
Calculations for cell diameter and carbon biomass based on raw FSC values only from live shipboard analysis.

Data used for cell concentration to calculate group-specific NPP is based on preserved samples for Syn, and live shipboard analysis for pico and nano euks. For euks, about 30% of the euks are lost when preserved.

I am using MD_PhotoAdapt_Biomass_Rates_17MAR20.xlsx data, initial values only. 
U9133 (cycle 2, exp 5) has no counts - also for the NPP calculations. Taking the average of the counts.

MD7: Pico is split into pico 1 and pico 2 - I am taking the mean of the two.
MD8 does not have readings. U9128 for initials

# Initialise
```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(here)
source(here("0_init_files", "init.R")) #for libraries
source(here("0_init_files", "init_markdown.R")) #for markdown libraries

colours_cycle <- readxl::read_excel(here("0_init_files", "colours_cycle.xlsx"), sheet = "cycle")
colours_cycle <- structure(colours_cycle$colour,.Names=colours_cycle$cycle)

colours_exp <- readxl::read_excel(here("0_init_files", "colours_cycle.xlsx"), sheet = "exp")
colours_exp <- structure(colours_exp$colour,.Names=colours_exp$exp)
```

# Read data
```{r, message=FALSE}
#use the "complete data TAN1810" sheet for the experiment number
C14_hot <- readxl::read_excel(here("0_data_used", "TAN1810_14C-Pico_25052020_QA_1.3.xlsx"), sheet = "Complete data TAN1810") %>%
  dplyr::select (`Cycle#`, `EXP#`, `CTD#`, `STN`, `SAMPLE`, `DEPTH`) %>%   
     dplyr::rename(cycle = `Cycle#`,
            exp = `EXP#`,
            CTD = `CTD#`,
            depth_cat = `SAMPLE`, 
            depth_m = `DEPTH`,
            station = `STN`)%>%
    distinct() %>%
  filter(!is.na(cycle))

# For live shipboard samples - syn, pico, nano
# match CTD number to MD number
md_meta <- readxl::read_excel(here("0_data_used", "MD_PhotoAdapt_Biomass_Rates_17MAR20_v1.1.xlsx"), sheet = "Summary of PhotoAdaption ") %>% 
  select(Station, Cast, Cycle, MD) %>%
  rename(station = Station,
         CTD = Cast,
         cycle = Cycle) %>%
  distinct()

md_meta$CTD <- sub("^", "U", md_meta$CTD) # add U before CTD number

#loop to read all sheets except md1. missing MD8. Reading only initials for FSC of Syn, pico, nano
md_comb<- data.frame()
for (i in c(2:7, 9:18)) {
md <- readxl::read_excel(here("0_data_used", "MD_PhotoAdapt_Biomass_Rates_17MAR20_v1.1.xlsx"), sheet = (str_c("MD", i)), skip = 2, n_max = 8) %>%
  select('Depth','SYN MEAN FS':'NANO MEAN FS') %>%
  mutate(MD = i) %>%
  rename(depth_m = 'Depth')
 md_comb <- bind_rows(md_comb, md)
}
md1 <- readxl::read_excel(here("0_data_used", "MD_PhotoAdapt_Biomass_Rates_17MAR20_v1.1.xlsx"), sheet = "MD1", skip = 2, n_max = 8) %>%
  select('Depth','SYN MEAN FS':'NANO MEAN FS') %>%
  mutate(MD = 1) %>%
  rename(depth_m = 'Depth')

md_lv <- full_join(md_comb, md1) %>%
  arrange(MD) %>%
  full_join(md_meta, by = "MD") %>% # add cycle number
  rename(syn_LV = 'SYN MEAN FS',
         pico = 'PICO-E MEAN FS',
         pico1 = 'PICO1 MEAN FS',
         pico2 = 'PICO2 MEAN FS',
         nano = 'NANO MEAN FS') %>%
  mutate(pico = case_when(MD==7 ~(pico1+pico2)/2, #calculate mean for pico MD7
                              TRUE ~ pico)) 

md_all <- md_lv %>%
  mutate(MD = str_c("MD", MD)) %>%
  # full_join(sv) %>%
  full_join(C14_hot) %>%
  mutate(cycle =  case_when (MD == "EOC1" ~ 1, # fill in missing values
                             MD == "EOC2" ~ 2,
                             MD == "EOC3" ~ 4,
                             MD == "EOC4" ~ 4,
                             MD == "EOC5" ~ 5,
                             MD == "MD1" ~ 1,
                             MD == "MD12" ~ 3,
                             MD == "MD8" ~2,
                             TRUE ~ cycle)) %>%
  mutate(exp = case_when (CTD == "U9128" ~ 4,
                          TRUE ~ exp)) %>%
  mutate(cycle_name = dplyr::recode(cycle, "1" = "SA-Sc", # add cycle name
                                    "2" = "SA1",
                                    "3" = "ST1",
                                    "4" = "ST2",
                                    "5" = "SA2")) %>%
  mutate(water_mass = dplyr::recode(cycle, "1" = "SA", # water mass assignment
                                    "2" = "SA",
                                    "3" = "ST",
                                    "4" = "ST",
                                    "5" = "SA")) %>%
  mutate(water_mass = fct_relevel(water_mass, "ST", "SA")) %>%
  mutate(depth_cat = case_when(depth_m == 12 ~ "SUR", # all surface exp are at 12 m
                               depth_m == 40 ~ "DCM", # assuming that cycles 1-4 DCM is at 40 m, although not accurate because cycle 3 one is at 25m and one is at 40 m. 
                               depth_m == 70 & cycle ==5 ~ "DCM",
                               TRUE ~ depth_cat)) %>% # assign depth levels for groups without
  mutate(depth_cat = fct_relevel(depth_cat, "SUR", "DCM")) %>%
  pivot_longer(cols = c("syn_LV":"nano"), values_to = "fsc", names_to = "population") %>%
  filter(!(population == "syn_LV" & MD == "MD3")) %>% # remove this because FSC values after 5 m is 0 
  filter(MD != "EOC3") %>% # only 2 readings (5 m and 12 m)
  filter(!(population == "syn_LV" & MD == "MD13" & depth_m == 100)) %>% # skews the values by alot. Calculated cell diameter is 4um.
  filter(!(is.na(fsc))) %>%
  arrange(cycle, MD, population, depth_m) 
  
```

# FSC raw values
syn MD3 cycle 1, fsc values after 5m is 0 - Removed. 
Overall, fsc values at surface is higher than DCM. 
```{r}
for(i in c("syn_LV", "pico", "nano")){
  md_select <-  md_all %>%
    filter(!(is.na(depth_cat))) %>%
    filter(!(is.na(fsc))) %>%
    filter(population ==i)
  p <-  md_all %>%
    filter(!(is.na(fsc))) %>%
    filter(population ==i) %>%
     ggplot(aes(x=fsc, y=depth_m)) +
      geom_path() +
      geom_point() +
    geom_point(data=md_select, aes(x=fsc, y=depth_m, color='red'),
               size=3)+
      scale_y_reverse() +
      facet_wrap(.~cycle_name + MD, nrow=4, scales = "free") +
  ggtitle(str_c("FSC raw - ", i)) +
    theme_minimal() +
    theme(legend.position = "none") 
 print(p)
# pdf(here(str_c("figures/3_fcm/fsc_raw_", i, ".pdf")), height=8, width=8) ; plot(p) ; dev.off()

}
```

# Calculate cell diameter and biomass
calculations based on bead size.
diam = um
vol = um3
biomass = pgC/cell
```{r}
cell_size <- md_all %>%
  ungroup() %>%
  mutate(diam = case_when(population == "syn_LV" ~ (224333+fsc)/235871,
                          population == "pico" ~ (224333+fsc)/235871,
                          population == "nano" ~(224333+fsc)/235871)) %>%
  mutate(vol = 4/3*3.14*(diam/2)^3) %>% #assuming spherical shape
  mutate(biomass = case_when(population == "pico" ~ (0.216*(vol^(0.939))),
                             population == "nano" ~ (0.216*(vol^(0.939))),
                             population == "syn_LV" ~ vol*265/1000)) 
writexl::write_xlsx(cell_size, path = here("Output", "Raw data", "fsc_cell size.xlsx"))
```




