---
title: "Growth rates figures"
author: "Denise Ong"
date: "11/17/2022"
output: html_document
---
Figures and statistical tests for cell diameter and growth rates.

# Initialise
```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(here)
source(here("0_init_files", "init.R")) #for libraries
source(here("0_init_files", "init_markdown.R")) #for markdown libraries

colours_cycle <- readxl::read_excel(here("0_init_files", "colours_cycle.xlsx"), sheet = "cycle")
colours_cycle <- structure(colours_cycle$colour,.Names=colours_cycle$cycle)

colours_exp <- readxl::read_excel(here("0_init_files", "colours_cycle.xlsx"), sheet = "exp")
colours_exp <- structure(colours_exp$colour,.Names=colours_exp$exp)
```

# Read data - only including cell size data with corresponding pp rates.
PP rates for exp 4 and 5 (except syn SV exp 5) do not have corresponding cell size data. Using the average biomass calculations for SA2.
```{r}
size_raw <-  readxl::read_excel(here("Output", "Raw data", "fsc_cell size.xlsx")) %>%
  separate(population, c("population", "fsc_type")) %>%
  mutate(population = dplyr::recode(population, "syn" = "Syn",
                             "pico" = "Pico",
                             "nano" = "Nano")) %>%
  mutate(cycle = as.character(cycle)) %>%
  mutate(exp = as.character(exp))

rates <- readxl::read_excel(here("Output", "Raw data", "Rates_cell_npp_method2.xlsx")) %>%
  select(-MD) %>%
  rename(depth_cat = depth,
         CTD = ctd)

join <- full_join(size_raw, rates)
join_sub <- join %>%
  filter(!(is.na(fsc)))


# assign average biomass values for pp rates for exp 4 and 5
sub <- join %>%
  filter(is.na(fsc)) %>%
  select(!(fsc_type:biomass))

sa2 <- join %>%
  filter(cycle == 2) %>% 
  filter(!(is.na(fsc))) %>%
  select(depth_m:biomass) %>%
  distinct() %>%
  filter(!(is.na(depth_cat))) %>%
  group_by(cycle, population, depth_cat, depth_m, fsc_type) %>%
  summarise(fsc = mean(fsc),
            diam = mean(diam),
            vol = mean(vol),
            biomass = mean(biomass))

check <- left_join(sub, sa2) %>%
  filter(!(exp == 4 & population == "Syn" & fsc_type == "SV"))

# join back again
join_rates <- full_join(check, join_sub) %>%
  filter(!(is.na(pp))) %>%
  unite(pop_fsc, c(population, fsc_type), sep = "_", remove = FALSE) %>%
  mutate(water_mass = fct_relevel(water_mass, "ST", "SA")) %>%
  mutate(depth_cat = fct_relevel(depth_cat, "SUR", "DCM")) %>%
  mutate(population = fct_relevel(population, "Syn", "Pico", "Nano")) %>%
  filter(pop_fsc!="Syn_SV") # remove preserved syn

#check how many measured cell size not included in exp calculations.
check <- join_rates %>%
select(diam) %>%
  distinct()

```

# calculate growth rates and normalise pp rates
growth rate = day-1
convert pp to fgC/cell/day
convert biomass to fgC/cell
```{r}
join_rates <- join_rates %>%
  mutate(biomass_fg = 1000*biomass) %>%
  mutate(growth_rate = (pp*24)/(biomass_fg))# multiply by 24 to change from per hour to per day. 
```


# Table 2 - summary for paper - pp, npp, cell conc, growth rate
```{r}

summary <- join_rates %>%
  select(population, cycle_name, pp, npp, conc_ml, growth_rate) %>%
  pivot_longer(pp:growth_rate, names_to = "data", values_to = "value") %>%
  group_by(data, population, cycle_name) %>%
  summarize(mean = round(mean(value), 2),
            sd = round(sd(value),2), 
            # n = n()
            ) %>%
   unite("mean_sd", mean:sd, sep = " ± ") %>%
  pivot_wider(names_from = "cycle_name", values_from = "mean_sd") %>%
  arrange(factor(data, levels = c('pp', 'npp', 'conc_ml', 'growth_rate'))) %>%
  select(data, population, ST1, ST2, SA1, SA2, SA3)
summary


writexl::write_xlsx(summary, path = "group_summary.xlsx")

```

# Diameter
after subset, no difference between ST and SA. 
```{r}
# check group diameters
summary <- join_sub %>%
  filter(!(is.na(depth_cat))) %>%
  filter(!(is.na(exp))) %>%
  unite(pop_fsc, c(population, fsc_type), sep = "_", remove = FALSE) %>%
  select(water_mass, cycle_name, pop_fsc, diam) %>%
  distinct() %>%
  group_by(pop_fsc) %>%
  summarize(mean = round(mean(diam), 2),
            sd = round(sd(diam),2), 
            n = n(),
            min = round(min(diam),2),
            max = round(max(diam),2)) %>%
   unite("mean_sd", mean:sd, sep = " ± ") 
summary

# including exp 4 and 5
p <- join_rates %>%
  select(water_mass, cycle_name, exp, depth_cat, pop_fsc, diam) %>%
  distinct() %>%
  ggplot()+
  geom_jitter(aes(x=cycle_name, y=diam), width = 0.2) +
  geom_boxplot(aes(x=water_mass, y=diam))+
  facet_grid(pop_fsc~water_mass, scales = "free") +
  scale_colour_manual(values = colours_cycle) 
p
# pdf("figures/3_fcm/diam_surdcm_group.pdf", height=7, width=6) ; plot(p) ; dev.off()

# surface vs DCM
for(i in c("Syn_LV", "Pico_NA", "Nano_NA")){
  p <-  join_rates %>%
    select(water_mass, cycle_name,exp,  pop_fsc,depth_cat, diam) %>%
    distinct() %>%
    filter(pop_fsc ==i) %>%
    ggplot() +
      geom_jitter(aes(x=cycle_name, y=diam), width = 0.2) +
      geom_boxplot(aes(x=water_mass, y=diam))+
      facet_grid(water_mass~depth_cat, scales = "free") +
      scale_colour_manual(values = colours_cycle) +
    ggtitle(i)
  print(p)
# pdf(here(str_c("figures/3_fcm/diam_sur dcm_", i, ".pdf")), height=4, width=6) ; plot(p) ; dev.off()

}
```


# Growth rate
plots for checking
```{r}
#Plot all
p <- join_rates %>%
  ggplot()+
  geom_boxplot(aes(x=pop_fsc, y=growth_rate))+
  #geom_jitter(aes(x=pop_fsc, y=growth_rate), width = 0.2) +
  # facet_grid(.~water_mass, scales = "free") +
  theme_minimal() +
  scale_y_continuous(trans='log10') + annotation_logticks(sides = "l")
p

# by depth
for (i in c("Syn", "Pico", "Nano")) {
  p <- join_rates %>%
  filter(pop_fsc!="Syn_SV") %>%
  filter(population == i) %>%
    ungroup() %>%
    complete(exp, depth_cat) %>%
  ggplot() +
    geom_point(aes(x=depth_cat, y=growth_rate))+
    facet_grid(.~exp, scales = "free_x", space = "free") +
    # scale_y_continuous(trans='log10') +
    # annotation_logticks(sides = "l")+
    ylab("Growth rate (day-1)") +
    ggtitle(i)
  print(p)
  # pdf(here(str_c("pp_depth_", i, ".pdf")), height=5, width=8) ; plot(p) ; dev.off()
}


```


## Figure 5 - Growth rates by water mass
```{r}
p <- join_rates %>%
  # filter(growth_rate <2)%>%
  ggplot(aes(x=population, y=growth_rate, fill = water_mass))+
  # geom_violin(aes(x=population, y=growth_rate))+
  geom_boxplot(outlier.shape = NA, alpha = 0.8)+
  geom_point(colour = "black",  alpha = 0.5, size = 1, position = position_jitterdodge(dodge.width = 0.75, jitter.width = 0.2))+
  # geom_jitter(, width = 0.1, alpha =  0.3, position=position_dodge(1)) +
  # scale_y_continuous(trans='log10') + annotation_logticks(sides = "l") +
  ylab("Growth rate (day-1)") +
  xlab(NULL)+
  theme(strip.background = element_blank(),
        legend.position="top") +
  scale_fill_manual(values = c("#C62828", "#1565C0"))

p

pdf("growth rate_not log.pdf", height=6, width=6) ; plot(p) ; dev.off()


```

```{r}
p <- join_rates %>%
    ggplot() +
      geom_jitter(aes(x=cycle_name, y=growth_rate, colour = depth_cat), width = 0.2, alpha = 0.5, size = 1.5)+
      geom_boxplot(aes(x=water_mass, y= growth_rate, fill=water_mass))+
      scale_colour_manual(values =  c("#009292", "#DB6D00"))+
      scale_fill_manual(values = c("#C62828", "#1565C0")) +
      facet_grid(population~water_mass, scales = "free", space = "free_x") +
  theme_bw()+
      # theme_linedraw() +
      theme(legend.position = "top",
            panel.spacing = unit(1, "lines"))+
      # scale_y_continuous(trans='log10') +
      # annotation_logticks(sides = "l", outside = TRUE)+
      # coord_cartesian(clip = "off")+
      ylab("Growth rate (day-1)") +
      xlab("Cycle")
p

pdf("growth rate by cycle and depth_not log.pdf", height=6, width=6) ; plot(p) ; dev.off()

```

#Statistical tests

## Diameter

### Surface vs DCM
Not significantly different between surface and DCM.
```{r}
library("ggpubr")

for (i in c("Syn", "Pico", "Nano")) {
t_test_df<- stats_rates %>%
  filter(population == i) %>%
  dplyr::select(water_mass,exp, depth_cat, diam) %>%
  distinct()

summary <- t_test_df %>%
  group_by(depth_cat) %>%
  summarise(count = n(),
    mean = mean(diam, na.rm = TRUE),
    sd = sd(diam, na.rm = TRUE))
print(summary)
 p <- t_test_df %>%
   ggplot(aes(x=depth_cat, y=diam)) + 
   geom_point() +
   geom_boxplot()  +ggtitle(i)
 print(p)
 
 print(i)

#Because p value <0.05, data is not normally distributed. Use the wilcoxon test
p<-with(t_test_df, shapiro.test(diam[depth_cat == "SUR"]))
print(p)
p<-with(t_test_df, shapiro.test(diam[depth_cat == "DCM"]))
print(p)
var <- var.test(diam ~ depth_cat, data = t_test_df)
print(var)
res <- wilcox.test(diam ~ depth_cat, data = t_test_df,
                   exact = FALSE)
print(res)
}
```

### ST vs SA
Not significantly different between ST and SA
```{r}
library("ggpubr")
 
for (i in c("Syn", "Pico", "Nano")) {
t_test_df<- join_rates %>%
  filter(population == i) %>%
  dplyr::select(water_mass, exp, depth_cat, diam) %>%
  distinct()

summary <- t_test_df %>%
  group_by(water_mass) %>%
  summarise(count = n(),
    mean = mean(diam, na.rm = TRUE),
    sd = sd(diam, na.rm = TRUE))
print(summary)
 p <- t_test_df %>%
   ggplot(aes(x=water_mass, y=diam)) + geom_point() +geom_boxplot()   +ggtitle(i)
 print(p)
 
 print(i)

#Because p value <0.05, data is not normally distributed. Use the wilcoxon test
p<-with(t_test_df, shapiro.test(diam[water_mass == "ST"]))
print(p)
p<-with(t_test_df, shapiro.test(diam[water_mass == "SA"]))
print(p)
var <- var.test(diam ~ water_mass, data = t_test_df)
print(var)
res <- wilcox.test(diam ~ water_mass, data = t_test_df,
                   paired = FALSE, exact = FALSE)
print(res)
}
```

## growth rate
### Between groups
```{r}

hist(stats_rates$growth_rate)
qqnorm(stats_rates$growth_rate) 
shapiro.test(stats_rates$growth_rate)

hist(log10(stats_rates$growth_rate))
qqnorm(log10(stats_rates$growth_rate))
shapiro.test(log10(stats_rates$growth_rate)) # meets assumptions for ANOVA


lm <- lm(log10(growth_rate) ~ population, data=stats_rates)
anova <- Anova(lm, type = "3")
pairwise.t.test(log10(stats_rates$growth_rate), stats_rates$population, p.adj='bonferroni')


library(lsmeans)
library(multcompView)

leastsquare = lsmeans(lm,
                      pairwise ~ population,
                      adjust = "tukey")
d <- summary(lsmeans(lm, ~population))
ggplot(d, aes(population)) +
  geom_line(aes(y = lsmean, group = 1)) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.2) +
  geom_point(aes(y = lsmean), size = 3, shape = 21, fill = "white") +
  labs(x = "Time", y = "ls mean", title = "ls mean result over time") +
  theme_bw()
```
### In each water mass, compare the patterns again
```{r}

for (i in c("SA")) {
  df <- stats_rates %>%
    filter(water_mass == i)
  hist(log10(df$growth_rate))
  qqnorm(log10(df$growth_rate))
  print(shapiro.test(log10(df$growth_rate)))
  
  lm <- lm(log10(growth_rate) ~ population, data=df)
  print(Anova(lm, type = "3"))
  print(pairwise.t.test(log10(df$growth_rate), df$population, p.adj='bonferroni'))
}

#ST is not normally distributed. Use kruskal wallis test.
for(i in c("ST")) {
  df <- stats_rates %>%
    filter(water_mass == i)
  hist(log10(df$growth_rate))
  qqnorm(log10(df$growth_rate))
  print(shapiro.test(log10(df$growth_rate)))
  print(kruskal.test(growth_rate~population, data = df)) # not significant
  lm <- lm(log10(growth_rate) ~ population, data=df)
  print(Anova(lm, type = "3"))
  print(pairwise.t.test(log10(df$growth_rate), df$population, p.adj='bonferroni'))
  
}

```


### SUR vs DCM
```{r}
library("ggpubr")

for (i in c("Syn", "Pico", "Nano")) {
t_test_df<- stats_rates %>%
  filter(population == i) %>%
  dplyr::select(water_mass, exp, depth_cat, growth_rate)

summary <- t_test_df %>%
  group_by(depth_cat) %>%
  summarise(count = n(),
    mean = mean(growth_rate, na.rm = TRUE),
    sd = sd(growth_rate, na.rm = TRUE),
median =  median(growth_rate, na.rm = TRUE))
print(summary)
 p <- t_test_df %>%
   ggplot(aes(x=depth_cat, y=growth_rate)) + geom_point() +geom_violin(alpha = 0.1)+
   #scale_y_continuous(trans='log10') + annotation_logticks(sides = "l")  
   ggtitle(i) 
 print(p)
 print(i)

#Because p value <0.05, data is not normally distributed. Use the wilcoxon test
p<-with(t_test_df, shapiro.test(growth_rate[depth_cat == "SUR"]))
print(p)
p<-with(t_test_df, shapiro.test(growth_rate[depth_cat == "DCM"]))
print(p)
var <- var.test(growth_rate ~ depth_cat, data = t_test_df)
print(var)
res <- wilcox.test(growth_rate ~ depth_cat, data = t_test_df,
                   paired = FALSE,exact = FALSE)
print(res)
t_test <- t.test(growth_rate ~ depth_cat, data = t_test_df, var.equal = TRUE)
print(t_test)
}

#only for syn
for (i in c("ST", "SA")) {
t_test_df<- stats_rates %>%
  filter(population == "Syn") %>%
  filter(water_mass == i) %>%
  dplyr::select(water_mass, exp, depth_cat, growth_rate)

summary <- t_test_df %>%
  group_by(depth_cat) %>%
  summarise(count = n(),
    mean = mean(growth_rate, na.rm = TRUE),
    sd = sd(growth_rate, na.rm = TRUE))
print(summary)
 p <- t_test_df %>%
   ggplot(aes(x=depth_cat, y=growth_rate)) + geom_point() +geom_boxplot()+ 
   #scale_y_continuous(trans='log10') + annotation_logticks(sides = "l")  
   +ggtitle(i) +facet_grid(.~water_mass)
 print(p)
 print(i)

#Because p value <0.05, data is not normally distributed. Use the wilcoxon test
p<-with(t_test_df, shapiro.test(growth_rate[depth_cat == "SUR"]))
print(p)
p<-with(t_test_df, shapiro.test(growth_rate[depth_cat == "DCM"]))
print(p)
var <- var.test(growth_rate ~ depth_cat, data = t_test_df)
print(var)
res <- wilcox.test(growth_rate ~ depth_cat, data = t_test_df,
                   paired = FALSE,exact = FALSE)
print(res)
t_test <- t.test(growth_rate ~ depth_cat, data = t_test_df, var.equal = TRUE)
print(t_test)
}

```

### T vs SA
ST is different from SA for syn and pico.
```{r}
library("ggpubr")

for (i in c("Syn", "Pico", "Nano")) {
t_test_df<- stats_rates %>%
  filter(population == i) %>%
  dplyr::select(water_mass, exp,  growth_rate)

summary <- t_test_df %>%
  group_by(water_mass) %>%
  summarise(count = n(),
    mean = mean(growth_rate, na.rm = TRUE),
    sd = sd(growth_rate, na.rm = TRUE))
print(summary)
 p <- t_test_df %>%
   ggplot(aes(x=water_mass, y=growth_rate)) + geom_point() +geom_boxplot()+ scale_y_continuous(trans='log10') + annotation_logticks(sides = "l")  +ggtitle(i)
 print(p)

 print(i)
 
#Because p value <0.05, data is not normally distributed. Use the wilcoxon test
p<-with(t_test_df, shapiro.test(growth_rate[water_mass == "ST"]))
print(p)
p<-with(t_test_df, shapiro.test(growth_rate[water_mass == "SA"]))
print(p)
var <- var.test(growth_rate ~ water_mass, data = t_test_df)
print(var)
res <- wilcox.test(growth_rate ~ water_mass, data = t_test_df,
                   paired = FALSE,exact = FALSE)
print(res)
t_test <- t.test(growth_rate ~ water_mass, data = t_test_df, var.equal = TRUE)
print(t_test)
}
```

# Summary for paper
```{r}
summary <- join_rates %>%
  group_by(pop_fsc) %>%
  summarize(#median = round(median(npp), 2),
            mean = round(mean(growth_rate), 2),
            sd = round(sd(growth_rate),2), n = n()) %>%
   unite("mean_sd", mean:sd, sep = " ± ") 
summary


summary <- stats_rates %>%
  group_by(population, water_mass) %>%
  summarize(median = round(median(npp), 2),
            mean = round(mean(growth_rate), 2),
            sd = round(sd(growth_rate),2), n = n()) %>%
   unite("mean_sd", mean:sd, sep = " ± ") 
summary
```
