---
title: "Growth rates figures"
author: "Denise Ong"
date: "11/17/2022"
output: html_document
---

Calculate growth rates.
Figures and statistical tests for cell diameter and growth rates.

# Initialise
```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(here)
source(here("0_init_files", "init.R")) #for libraries
source(here("0_init_files", "init_markdown.R")) #for markdown libraries

colours_cycle <- readxl::read_excel(here("0_init_files", "colours_cycle.xlsx"), sheet = "cycle")
colours_cycle <- structure(colours_cycle$colour,.Names=colours_cycle$cycle)

colours_exp <- readxl::read_excel(here("0_init_files", "colours_cycle.xlsx"), sheet = "exp")
colours_exp <- structure(colours_exp$colour,.Names=colours_exp$exp)
```

# Read data - only including cell size data with corresponding pp rates.
PP rates for exp 4 and 5 (except syn SV exp 5) do not have corresponding cell size data. Using the average biomass calculations for SA1.
```{r}
size_raw <-  readxl::read_excel(here("Output", "Raw data", "fsc_cell size.xlsx")) %>%
  separate(population, c("population", "fsc_type")) %>%
  mutate(population = dplyr::recode(population, "syn" = "Syn",
                             "pico" = "Pico",
                             "nano" = "Nano")) %>%
  mutate(cycle = as.character(cycle)) %>%
  mutate(exp = as.character(exp))

rates <- readxl::read_excel(here("Output", "Raw data", "Rates_cell_npp_method2.xlsx")) %>%
  select(-MD) %>%
  rename(depth_cat = depth,
         CTD = ctd)

join <- full_join(size_raw, rates)
join_sub <- join %>%
  filter(!(is.na(fsc)))


# assign average biomass values for pp rates for exp 4 and 5
sub <- join %>%
  filter(is.na(fsc)) %>%
  select(!(fsc_type:biomass))

sa1 <- join %>%
  filter(cycle == 2) %>% 
  filter(!(is.na(fsc))) %>%
  select(depth_m:biomass) %>%
  distinct() %>%
  filter(!(is.na(depth_cat))) %>%
  group_by(cycle, population, depth_cat, depth_m, fsc_type) %>%
  summarise(fsc = mean(fsc),
            diam = mean(diam),
            vol = mean(vol),
            biomass = mean(biomass))

check <- left_join(sub, sa1) %>%
  filter(!(exp == 4 & population == "Syn" & fsc_type == "SV"))

# join back again
join_rates <- full_join(check, join_sub) %>%
  filter(!(is.na(pp))) %>%
  unite(pop_fsc, c(population, fsc_type), sep = "_", remove = FALSE) %>%
  mutate(water_mass = fct_relevel(water_mass, "ST", "SA")) %>%
  mutate(depth_cat = fct_relevel(depth_cat, "SUR", "DCM")) %>%
  mutate(population = fct_relevel(population, "Syn", "Pico", "Nano")) %>%
  filter(pop_fsc!="Syn_SV") # remove preserved syn

#check how many measured cell size not included in exp calculations.
join_rates %>% select(diam) %>%
  distinct() %>%
  summarise(n = n())

```

# calculate growth rates and normalise pp rates
growth rate = day-1
convert pp to fgC/cell/day
convert biomass to fgC/cell
```{r}
join_rates <- join_rates %>%
  mutate(biomass_fg = 1000*biomass) %>%
  mutate(growth_rate = (pp*24)/(biomass_fg)) %>% # multiply by 24 to change from per hour to per day. 
  select(-MD, -pico1, -pico2) 
# writexl::write_xlsx(join_rates, path = here("Output", "Raw data", "Growth rates all.xlsx"))
```

## supplementary plot
by water mass and depth
```{r}

p <- join_rates %>%
  ggplot() +
  geom_jitter(aes(x=fct_relevel(exp,"1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11"), y=growth_rate, shape = depth_cat), width = 0.2, alpha = 0.5) +
  facet_grid(population~fct_relevel(cycle_name, "ST1", "ST2", "SA-Sc", "SA1", "SA2"), scales = "free", space = "free_x") +
      theme(panel.spacing.x = unit(0,"lines"),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            strip.background = element_blank(),
            strip.text.y.right = element_text(angle = 0),
            panel.border = element_rect(colour = "black", fill = NA),
            legend.position = "top")+
  # scale_y_continuous(trans='log10') +
  # annotation_logticks(sides = "l")+
  ylab("Growth rate (day-1)") +
  xlab("Cycle")
p
pdf("growth rate by cycle and depth_not log_v2.pdf", height=6, width=6) ; plot(p) ; dev.off()


# p <- join_rates %>%
#     ggplot() +
#       geom_jitter(aes(x=cycle_name, y=growth_rate, colour = depth_cat), width = 0.2, alpha = 0.5, size = 1.5)+
#       geom_boxplot(aes(x=water_mass, y= growth_rate, fill=water_mass))+
#       scale_colour_manual(values =  c("#009292", "#DB6D00"))+
#       scale_fill_manual(values = c("#C62828", "#1565C0")) +
#       facet_grid(population~water_mass, scales = "free", space = "free_x") +
#   theme_bw()+
#       # theme_linedraw() +
#       theme(legend.position = "top",
#             panel.spacing = unit(1, "lines"))+
#       # scale_y_continuous(trans='log10') +
#       # annotation_logticks(sides = "l", outside = TRUE)+
#       # coord_cartesian(clip = "off")+
#       ylab("Growth rate (day-1)") +
#       xlab("Cycle")
# p


```

```{r}
join_rates <- join_rates %>%
  filter(exp!=7)
```


# Table 2 - summary for paper - pp, npp, cell conc, growth rate
```{r}

summary <- join_rates %>%
  select(population, cycle_name, pp, npp, conc_ml, growth_rate) %>%
  pivot_longer(pp:growth_rate, names_to = "data", values_to = "value") %>%
  group_by(data, population, cycle_name) %>%
  summarize(mean = round(mean(value), 2),
            sd = round(sd(value),2), 
            # n = n()
            ) %>%
   unite("mean_sd", mean:sd, sep = " ± ") %>%
  pivot_wider(names_from = "cycle_name", values_from = "mean_sd") %>%
  arrange(factor(data, levels = c('pp', 'npp', 'conc_ml', 'growth_rate'))) %>%
  select(data, population, ST1, ST2, 'SA-Sc', SA1, SA2)
summary


# writexl::write_xlsx(summary, path = here("Output", "Tables", "Table 2_group summary.xlsx"))

```

# Summary for text
```{r}
summary <- join_rates %>%
  group_by(population) %>%
  summarize(#median = round(median(npp), 2),
            mean = round(mean(growth_rate), 2),
            sd = round(sd(growth_rate),2), n = n()) %>%
   unite("mean_sd", mean:sd, sep = " ± ") 
summary


summary <- join_rates %>%
  group_by(population, cycle_name) %>%
  summarize(median = round(median(npp), 2),
            mean = round(mean(growth_rate), 2),
            sd = round(sd(growth_rate),2), n = n()) %>%
   unite("mean_sd", mean:sd, sep = " ± ") 
summary
```

# Diameter
## Plots for checking
after subset, no difference between ST and SA. 
```{r}
# check group diameters
summary <- join_sub %>%
  filter(!(is.na(depth_cat))) %>%
  filter(!(is.na(exp))) %>%
  unite(pop_fsc, c(population, fsc_type), sep = "_", remove = FALSE) %>%
  select(water_mass, cycle_name, pop_fsc, diam) %>%
  distinct() %>%
  group_by(pop_fsc) %>%
  summarize(mean = round(mean(diam), 2),
            sd = round(sd(diam),2), 
            n = n(),
            min = round(min(diam),2),
            max = round(max(diam),2)) %>%
   unite("mean_sd", mean:sd, sep = " ± ") 
summary

# including exp 4 and 5
p <- join_rates %>%
  select(water_mass, cycle_name, exp, depth_cat, pop_fsc, diam) %>%
  distinct() %>%
  ggplot()+
  geom_jitter(aes(x=cycle_name, y=diam), width = 0.2) +
  geom_boxplot(aes(x=water_mass, y=diam))+
  facet_grid(pop_fsc~water_mass, scales = "free") +
  scale_colour_manual(values = colours_cycle) 
p
# pdf("figures/3_fcm/diam_surdcm_group.pdf", height=7, width=6) ; plot(p) ; dev.off()

# surface vs DCM
for(i in c("Syn_LV", "Pico_NA", "Nano_NA")){
  p <-  join_rates %>%
    select(water_mass, cycle_name,exp,  pop_fsc,depth_cat, diam) %>%
    distinct() %>%
    filter(pop_fsc ==i) %>%
    ggplot() +
      geom_jitter(aes(x=cycle_name, y=diam), width = 0.2) +
      geom_boxplot(aes(x=water_mass, y=diam))+
      facet_grid(water_mass~depth_cat, scales = "free") +
      scale_colour_manual(values = colours_cycle) +
    ggtitle(i)
  print(p)
# pdf(here(str_c("figures/3_fcm/diam_sur dcm_", i, ".pdf")), height=4, width=6) ; plot(p) ; dev.off()

}
```

## Statistical tests
###Surface vs DCM
Not significantly different between surface and DCM.
syn: W = 32, p-value = 0.4256
pico: W = 25, p-value = 1
nano: W = 45, p-value = 0.1155
```{r}
library("ggpubr")

# i = "Syn"
i = "Pico"
# i = "Nano"
t_test_df<- join_rates %>%
  filter(population == i) %>%
  dplyr::select(water_mass,exp, depth_cat, diam) %>%
  distinct()

summary <- t_test_df %>%
  group_by(depth_cat) %>%
  summarise(count = n(),
    mean = mean(diam, na.rm = TRUE),
    sd = sd(diam, na.rm = TRUE))
print(summary)
 p <- t_test_df %>%
   ggplot(aes(x=depth_cat, y=diam)) + 
   geom_point() +
   geom_boxplot()  +ggtitle(i)
 print(p)

#Because p value <0.05, data is not normally distributed. Use the wilcoxon test
p<-with(t_test_df, shapiro.test(diam[depth_cat == "SUR"]))
print(p)
p<-with(t_test_df, shapiro.test(diam[depth_cat == "DCM"]))
print(p)
var <- var.test(diam ~ depth_cat, data = t_test_df)
print(var)
res <- wilcox.test(diam ~ depth_cat, data = t_test_df,
                   exact = FALSE)
print(res)

```

### ST vs SA
Not significantly different between ST and SA

syn: W = 36, p-value = 0.198
pico:W = 15, p-value = 0.2442
nano: W = 40, p-value = 0.1737
```{r}
library("ggpubr")
 
for (i in c("Syn", "Pico", "Nano")) {
t_test_df<- join_rates %>%
  filter(population == i) %>%
  dplyr::select(water_mass, exp, depth_cat, diam) %>%
  distinct()

summary <- t_test_df %>%
  group_by(water_mass) %>%
  summarise(count = n(),
    mean = mean(diam, na.rm = TRUE),
    sd = sd(diam, na.rm = TRUE))
print(summary)
 p <- t_test_df %>%
   ggplot(aes(x=water_mass, y=diam)) + geom_point() +geom_boxplot()   +ggtitle(i)
 print(p)
 
 print(i)

#Because p value <0.05, data is not normally distributed. Use the wilcoxon test
p<-with(t_test_df, shapiro.test(diam[water_mass == "ST"]))
print(p)
p<-with(t_test_df, shapiro.test(diam[water_mass == "SA"]))
print(p)
var <- var.test(diam ~ water_mass, data = t_test_df)
print(var)
res <- wilcox.test(diam ~ water_mass, data = t_test_df,
                   paired = FALSE, exact = FALSE)
print(res)
}
```

# Growth rate
## Statistical tests
### SUR vs DCM
syn: W = 101, p-value = 0.06563
t = -2.0221, df = 37, p-value = 0.05044
depth_cat | Predicted |       95% CI
SUR       |      0.08 | [0.06, 0.11]
DCM       |      0.14 | [0.09, 0.21]

pico: W = 155, p-value = 0.5636
t = -0.81659, df = 38, p-value = 0.4193
depth_cat | Predicted |       95% CI
SUR       |      0.16 | [0.11, 0.23]
DCM       |      0.20 | [0.12, 0.34]

nano: W = 171, p-value = 0.5133
t = -0.35411, df = 40, p-value = 0.7251
depth_cat | Predicted |       95% CI
SUR       |      0.23 | [0.17, 0.30]
DCM       |      0.25 | [0.16, 0.38]

```{r}
library("ggpubr")

# i = "Syn"
# i = "Pico"
i = "Nano"

t_test_df<- join_rates %>%
  filter(population == i) %>%
  dplyr::select(water_mass, exp, depth_cat, growth_rate)

summary <- t_test_df %>%
  group_by(depth_cat) %>%
  summarise(count = n(),
    mean = mean(growth_rate, na.rm = TRUE),
    sd = sd(growth_rate, na.rm = TRUE),
    median =  median(growth_rate, na.rm = TRUE))
print(summary)
 p <- t_test_df %>%
   ggplot(aes(x=depth_cat, y=growth_rate)) + geom_point() +geom_violin(alpha = 0.1)#+ facet_grid(.~water_mass) +
   # scale_y_continuous(trans='log10') + annotation_logticks(sides = "l") +
   ggtitle(i) 
 print(p)

# checks, all groups meet assumptions of normality and homoskedasticity
hist(log10(t_test_df$growth_rate))
qqnorm(log10(t_test_df$growth_rate))
shapiro.test(log10(t_test_df$growth_rate))

lm <- lm(log10(growth_rate) ~ depth_cat, data=t_test_df)
plot(lm) 

# check for variance. If variance sig different, use var.equal = FALSE
var.test(log10(growth_rate) ~ depth_cat, t_test_df)

# parametric
t.test(log10(growth_rate) ~ depth_cat, data =  t_test_df, var.equal = TRUE)

lm_emm <- ggemmeans(lm, 
                    terms = c("depth_cat"),
                    type = "fixed")
lm_emm
plot(lm_emm)

res <- wilcox.test(growth_rate ~ depth_cat, data = t_test_df, paired = FALSE,exact = FALSE)
print(res)


```


### Overall between groups
Anova Table (Type III tests)

Response: log10(growth_rate)
            Sum Sq  Df F value    Pr(>F)    
(Intercept) 36.343   1 191.060 < 2.2e-16 ***
population   2.706   2   7.113  0.001183 ** 
Residuals   23.967 126 
population | Predicted |       95% CI
Syn        |      0.12 | [0.09, 0.16]
Pico       |      0.20 | [0.15, 0.27]
Nano       |      0.26 | [0.19, 0.35]
```{r}
hist(join_rates$growth_rate)
qqnorm(join_rates$growth_rate)
shapiro.test(join_rates$growth_rate)

hist(log10(join_rates$growth_rate))
qqnorm(log10(join_rates$growth_rate))
shapiro.test(log10(join_rates$growth_rate)) # meets assumptions for ANOVA


lm <- lm(log10(growth_rate) ~ population, data=join_rates)

# plot(lm)
Anova(lm, type = "3")

library(ggeffects)
lm_emm <- ggemmeans(lm, 
                    terms = c("population"),
                    type = "fixed")

lm_emm
plot(lm_emm)

#post-hoc test
library(emmeans)
lm.emm <- emmeans(lm, ~ population)
contrast(lm.emm, "consec", simple = "each", combine = TRUE, adjust = "mvt")
```
```{r, eval=FALSE}
# because of unbalanced design, subset 16 samples randomly for each group first 
# no difference when subset. Use all samples.
# set.seed(1)
# set.seed(78674)
# set.seed(234154)
# set.seed(3413)
# set.seed(76876)
set.seed(53888)
sub <- join_rates %>%
  group_by(population, water_mass) %>%
  sample_n(14)

# interaction plot population and water mass?
lm1 <- lm(log10(growth_rate) ~ population*water_mass, data=sub)
lm2 <- lm(log10(growth_rate) ~ population+water_mass, data=sub) # choose this model
lm3 <- lm(log10(npp) ~ population, data=sub)
# lm4 <- lm(log10(npp) ~ water_mass, data=sub)

anova(lm3, lm2)
# choose lm2
# plot(lm2)
Anova(lm1, type = "3") # p< 0.0001
summary(lm1)

lm_emm <- ggemmeans(lm1,
                    terms = c("population", "water_mass"),
                    type = "fixed")

lm_emm
print(plot(lm_emm) + scale_y_continuous(trans='log10') +
      annotation_logticks(sides = "l", outside = TRUE))

#post-hoc test
# library(emmeans)
lm.emm <- emmeans(lm1, ~ population*water_mass)
print(contrast(lm.emm, "consec", simple = "each", combine = TRUE, adjust = "mvt"))
```

Multiple R-squared:  0.3106,	Adjusted R-squared:  0.2806 
F-statistic: 10.36 on 5 and 115 DF,  p-value: 3.203e-08

water_mass = ST

population | Predicted |       95% CI
Syn        |      0.19 | [0.13, 0.28]
Pico       |      0.24 | [0.16, 0.36]
Nano       |      0.20 | [0.13, 0.29]

 water_mass = SA

population | Predicted |       95% CI
Syn        |      0.07 | [0.05, 0.09]
Pico       |      0.15 | [0.11, 0.20]
Nano       |      0.25 | [0.19, 0.34]


```{r}
# test interaction factors
lm1 <- lm(log10(growth_rate) ~ population*water_mass, data=join_rates) # choose this model
# lm2 <- lm(log10(growth_rate) ~ population+water_mass, data=join_rates) 

# anova(lm1, lm2) # choose lm1

# plot(lm1)
Anova(lm1, type = "3") # because no interaction, use type 2
summary(lm1)

#post-hoc test
# library(emmeans)
lm.emm <- emmeans(lm1, ~ population*water_mass)
emm_df<-as.data.frame(contrast(lm.emm, "consec", simple = "each", combine = TRUE, adjust = "mvt"))

#plot predicted means and CI
lm_emm <- ggemmeans(lm1,
                    terms = c("population", "water_mass"),
                    type = "fixed")

lm_emm

emm.npp.df <- as.data.frame(lm_emm)

plot(lm_emm) + scale_y_continuous(trans='log10') +
      annotation_logticks(sides = "l", outside = TRUE)


growth_mixed <- ggplot() +
        # Raw data
        geom_half_violin(aes(x = population,
                            y = growth_rate, 
                            split = water_mass, 
                            fill=water_mass),
                        nudge = 0.001,
                        alpha=0.5,
                        data = join_rates) +
        # Line of model predicted means
        geom_pointrange(aes(x = x,
                            y = predicted,
                            ymin = conf.low,
                            ymax = conf.high,
                            colour = group),
                        size = 0.6,
                        lwd = 1.3, 
                        position = position_dodge(width = 0.25),
                        data = lm_emm) +
        labs(x = "Population",
             y = "Growth rate (day-1)",
             colour = "Water mass",
             shape = "Cycle") +
      scale_shape_manual(values=c(0,1, 15, 17, 19)) +
      scale_fill_manual(values = c("#C62828", "#1565C0"))+ 
      scale_colour_manual(values = c("#C62828", "#1565C0"))+
      scale_y_continuous(trans='log10') +
      annotation_logticks(sides = "l") +
      annotate("text", x = 2.75, y = 1, label = "ANOVA, R2=0.28,\n Population X Water mass: F(5,115) = 10.36,\n p = 0.001", size= 3)
growth_mixed
```

```{r}
#load pp and npp plots from group rates figures_4.0.Rmd file
# p <- ((pp_mixed|npp_mixed)/ (st|sa)) + 
#   plot_annotation(tag_levels = 'A') +  
#   plot_layout(guides = 'collect')  &
#   theme(legend.position='top')  & 
#   theme(axis.text = element_text(size = 12), 
#         legend.text = element_text(size = 12), 
#         legend.title = element_text(size = 12), 
#         axis.title = element_text(size = 12)
#         )
# p

# pdf("pp_npp_growth rate_mean and CI.pdf", height=9.5, width=9) ; plot(p) ; dev.off()

p <- (pp_mixed|npp_mixed)/(growth_mixed|plot_spacer())+ 
  plot_annotation(tag_levels = 'A') +  
  plot_layout(guides = 'collect')  &
  theme(legend.position='top')  & 
  theme(axis.text = element_text(size = 12), 
        legend.text = element_text(size = 12), 
        legend.title = element_text(size = 12), 
        axis.title = element_text(size = 12)
        )
p

pdf("pp_npp_growth rate_mean and CI_V2.pdf", height=9.5, width=9) ; plot(p) ; dev.off()

```


<!--
### Main figure and stats - In each water mass, compare between groups

```{r}
# ST
df <- join_rates %>%
  filter(water_mass == "ST")
hist(log10(df$growth_rate))
qqnorm(log10(df$growth_rate))
print(shapiro.test(log10(df$growth_rate)))
lm <- lm(log10(growth_rate) ~ population, data=df)
plot(lm)
Anova(lm, type = "3")
summary(lm)
lm_emm <- ggemmeans(lm, 
                    terms = c("population"),
                    type = "fixed")
lm_emm
plot(lm_emm)

#post-hoc test
lm.emm <- emmeans(lm, ~ population)
contrast(lm.emm, "consec", simple = "each", combine = TRUE, adjust = "mvt")

```
```{r}
library(gghalves)
st <- ggplot() +
        # Raw data
       geom_half_violin(aes(x = population,
                            y = growth_rate),
                        side = 'l',
                        nudge = 0.05,
                        fill = "#C62828",
                        alpha = 0.5,
                        data = df) +
        # geom_half_point_panel(aes(x = population,
        #                     y = npp),
        #                 side = 'l', alpha = 0.6,
        #                 size = 1, range_scale = 0.5,
        #                 data = pp_rates) +
        # Line of model predicted means
        geom_pointrange(aes(x = x,
                            y = predicted,
                            ymin = conf.low,
                            ymax = conf.high),
                        size = 0.6,
                        lwd = 1.3, 
                        position = position_dodge(width = 1),
                        colour = "#C62828",
                        data = lm_emm) +
        labs(x = "Population",
             y = "Growth rate (day-1)",
             # colour = "Water mass",
             # shape = "Cycle",
             # title = "Subtropical"
             ) +
      # scale_shape_manual(values=c(0,1, 15, 17, 19)) +
      # scale_colour_manual(values = c("#C62828", "#1565C0"))+ 
      scale_y_continuous(trans='log10',
                         limits = c(0.01, 3.5)
                         ) +
      annotation_logticks(sides = "l") +
  annotate("text", x = 3, y = 0.015, label = "ANOVA; R2=-0.03,\n  F{2, 47} = 0.14, p = 0.87")

st
```

```{r}
#SA
df <- join_rates %>%
  filter(water_mass == "SA")
hist(log10(df$growth_rate))
qqnorm(log10(df$growth_rate))
print(shapiro.test(log10(df$growth_rate)))

lm <- lm(log10(growth_rate) ~ population, data=df)
plot(lm)
Anova(lm, type = "3")
summary(lm)

lm_emm <- ggemmeans(lm, 
                    terms = c("population"),
                    type = "fixed")
lm_emm
plot(lm_emm)

#post-hoc test
lm.emm <- emmeans(lm, ~ population)
contrast(lm.emm, "consec", simple = "each", combine = TRUE, adjust = "mvt")
```
```{r}
sa <- ggplot() +
        # Raw data
       geom_half_violin(aes(x = population,
                            y = growth_rate),
                        side = 'l',
                        nudge = 0.05,
                        fill = "#1565C0",
                        alpha = 0.5,
                        data = df) +
       #  geom_half_point_panel(aes(x = population,
       #                      y = growth_rate),
       #                  side = 'l', alpha = 0.6,
       #                  size = 1, range_scale = 0.5,
       #                  data = df) +
        # Line of model predicted means
        geom_pointrange(aes(x = x,
                            y = predicted,
                            ymin = conf.low,
                            ymax = conf.high),
                        size = 0.6,
                        lwd = 1.3, 
                        position = position_dodge(width = 1),
                        colour = "#1565C0",
                        data = lm_emm) +
        labs(x = "Population",
             y = "Growth rate (day-1)",
             # title = "Subantarctic"
             ) +
      scale_y_continuous(trans='log10',
                         limits = c(0.01, 3.5)
                         ) +
      annotation_logticks(sides = "l") +
  annotate("text", x = 3, y = 0.015, label = "ANOVA; R2=0.35,\n F{2, 76} = 21.73, p < 0.0001")

sa
```

### ST vs SA
ST is different from SA for syn and pico.

Syn: t = 5.9525, df = 40, p-value = 5.512e-07

water_mass | Predicted |       95% CI
ST         |      0.28 | [0.19, 0.41]
SA         |      0.07 | [0.05, 0.09]

pico: t = 2.7745, df = 41, p-value = 0.008287
water_mass | Predicted |       95% CI
ST         |      0.33 | [0.21, 0.53]
SA         |      0.15 | [0.10, 0.21]

nano:t = 0.2725, df = 18.577, p-value = 0.7882
water_mass | Predicted |       95% CI
ST         |      0.28 | [0.17, 0.45]
SA         |      0.25 | [0.18, 0.36]
```{r}
# i = "Syn"
i = "Pico"
# i = "Nano"

t_test_df<- join_rates %>%
  filter(population == i) %>%
  dplyr::select(water_mass, exp,  growth_rate)

summary <- t_test_df %>%
  group_by(water_mass) %>%
  summarise(count = n(),
    mean = mean(growth_rate, na.rm = TRUE),
    sd = sd(growth_rate, na.rm = TRUE))
print(summary)
 p <- t_test_df %>%
   ggplot(aes(x=water_mass, y=growth_rate)) + geom_point() +geom_boxplot()+ scale_y_continuous(trans='log10') + annotation_logticks(sides = "l")  +ggtitle(i)
 print(p)
 
# checks, all groups meet assumptions of normality and homoskedasticity
hist(log10(t_test_df$growth_rate))
qqnorm(log10(t_test_df$growth_rate))
shapiro.test(log10(t_test_df$growth_rate))

lm <- lm(log10(growth_rate) ~ water_mass, data=t_test_df)
plot(lm) 

# check for variance. If variance sig different, use var.equal = FALSE
var.test(log10(growth_rate) ~ water_mass, t_test_df)

# parametric
t.test(log10(growth_rate) ~ water_mass, data =  t_test_df, var.equal = FALSE)

lm_emm <- ggemmeans(lm, 
                    terms = c("water_mass"),
                    type = "fixed")
lm_emm
plot(lm_emm)
```
-->

