---
title: "20220412_sorted euks NMDS"
author: "Denise Ong"
date: "4/12/2022"
output: html_document
---

Do a NMDS to reflect overall community. 
By rarefifying the data, compare with normalising.
Removing Syndiniales and other samples that are excluded.
use sequence hash to merge? 
remove 

# Initialize
```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(here)
library(vegan)
source(here("phyloseq", "init_files", "init.R")) #for libraries
source(here("phyloseq", "init_files", "init_markdown.R")) #for markdown libraries
source(here("phyloseq", "init_files", "define_functions.R")) #functions to 1.transform to dataframe, 2.normalise and 3.create treemaps 

#Colours for cycle
colours_cycle <- readxl::read_excel(here("phyloseq", "init_files", "colours_cycle copy.xlsx"))
colours_cycle <- structure(colours_cycle$colour,.Names=colours_cycle$cycle)
```

```{r}
# convert the sample_data() within a phyloseq object to a vegan compatible data object
pssd2veg <- function(physeq) {
  sd <- sample_data(physeq)
  return(as(sd,"data.frame"))
}

# convert the otu_table() within a phyloseq object to a vegan compatible data object
psotu2veg <- function(physeq) {
  OTU <- otu_table(physeq)
  if (taxa_are_rows(OTU)) {
    OTU <- t(OTU)
  }
  return(as(OTU, "matrix"))
}
```


# Read phyloseq for pico and nano
```{r}
source(here("phyloseq", "init_files", "read_pico.R"))
source(here("phyloseq", "init_files", "read_nano.R"))

#merge sorted pico and nano ps, because they have the same ASV code
ps_merge<- merge_phyloseq(ps_pico, ps_nano)

long_merge <- phyloseq_transform_to_long(ps_merge) 

sample <- data.frame(ps_merge@sam_data) %>%
    select(sorting_population, sample_name, water_mass, cycle_name_2, cycle_exp, vial_type, SAMPLE) %>%
    rename(depth = SAMPLE)
otu <- psotu2veg(ps_merge)

```

#checks before plotting - ok, ignore for now.
```{r}
#to check for every sample number of reads and if all asvs have reads -  ok, ignore code now.

otu_long <- otu %>%
  as.data.frame() 
otu_long$file_code <-rownames(otu_long)
otu_long <- otu_long %>%
    pivot_longer(cols != file_code, names_to = "otu", values_to = "reads") %>%
    group_by(file_code)  %>%
    summarise(sum = sum(reads)) %>%
    group_by(otu) %>%
    summarise(sum = sum(reads))

```


#plot overall
```{r}
#rarify
set.seed(1)
dist <- avgdist(otu, sample = 9000, dmethod = "bray") #rarify to control for uneven number of reads per sample.
nmds<- metaMDS(dist)

nmds_sample <- scores(nmds) %>%
    as.tibble(rownames="sample_name") %>%
    inner_join(., sample, by = "sample_name")

#plot overall - 
p <- nmds_sample %>%
    mutate(cycle_name = fct_relevel(cycle_name_2, "ST1", "ST2", "SA1","SA2","SA3" )) %>%
    ggplot(aes(x=NMDS1, y= NMDS2, colour = cycle_name, shape = depth)) +
    geom_point() +
    scale_colour_manual(values = colours_cycle)+
  ggtitle("NMDS rarify reads")
p


#pdf(here("phyloseq", "paper_c14", "20220417", "sorted_euks_nmds_rarify.pdf"), height=6, width=8) ; plot(p) ; dev.off()

#normalise
ps_merge_norm <- phyloseq_normalize_median(ps_merge)
otu_norm <- psotu2veg(ps_merge_norm)

dist_norm <- vegdist(otu_norm, method = "bray")
nmds_norm <- metaMDS(dist_norm)

nmds_sample_norm <- scores(nmds_norm) %>%
    as.tibble(rownames="sample_name") %>%
    inner_join(., sample, by = "sample_name")

p <- nmds_sample_norm %>%
    mutate(cycle_name = fct_relevel(cycle_name_2, "ST1", "ST2", "SA1","SA2","SA3" )) %>%
    ggplot(aes(x=NMDS1, y= NMDS2, colour = cycle_name, shape = depth)) +
    geom_point() +
    scale_colour_manual(values = colours_cycle) +
  ggtitle("NMDS normalise reads")
p

#pdf(here("phyloseq", "paper_c14", "20220417", "sorted_euks_nmds_norm.pdf"), height=6, width=8) ; plot(p) ; dev.off()

```


# subset removed groups and samples from the plot
##remove heterotrophs (added 20220417)
```{r}
source(here("phyloseq", "trophic_mode", "trophic_function.R"))  
ps_merge_1 <- phyloseq_assign_trophic(ps_merge)
long_merge_1 <- phyloseq_transform_to_long(ps_merge_1) 
#list of ASVs for heterotrophs
het_list <- long_merge_1  %>%
  filter(trophic_mode  == "heterotrophic") %>%
  select(asv_code) %>%
  distinct()

otu_long <- otu %>%
  as.data.frame() 
otu_long$file_code <-rownames(otu_long)
otu_sub <- otu_long %>%
    pivot_longer(cols = !file_code, names_to = "otu", values_to = "reads") %>%
    anti_join(het_list, by = c("otu" = "asv_code")) %>%
    pivot_wider(names_from = otu, values_from = reads) %>%
  as.matrix()

```


##  Remove sample nano-12, pico-15 and pico-1-30
```{r}
otu_sub <- otu %>%
  as.data.frame() 
otu_sub$file_code <-rownames(otu_sub)

otu_sub <- otu_sub %>% 
  filter(!(file_code %in% c("nano-12", "pico-15", "pico-01-30"))) %>%
  select(-file_code) %>%
  as.matrix()
```

```{r}
set.seed(1)
dist <- avgdist(otu_sub, sample = 9000, dmethod = "bray") #rarify to control for uneven number of reads per sample.
nmds_sub<- metaMDS(dist)

nmds_sample_sub <- scores(nmds_sub) %>%
    as.tibble(rownames="sample_name") %>%
    inner_join(., sample, by = "sample_name")

goodness(nmds_sub)
stressplot(nmds_sub)

nmds_sub$stress

#plot overall
p <- nmds_sample_sub %>%
    mutate(cycle_name = fct_relevel(cycle_name_2, "ST1", "ST2", "SA1","SA2","SA3" )) %>%
    ggplot(aes(x=NMDS1, y= NMDS2, colour = cycle_name, shape = depth)) +
    geom_point() +
  ggalt::geom_encircle(aes(group = sorting_population), expand = 0, s_shape=1,
                  alpha = 0.2, color = "black", size=2) +
    scale_colour_manual(values = colours_cycle) +
   theme(legend.position="bottom")

p
 pdf(here("phyloseq", "figures_GRC", "sorted_euks_nmds.pdf"), height=7, width=6) ; plot(p) ; dev.off()
```

```{r}
#plot by depth and vial - to check influence of depth.
p <- nmds_sample_sub %>%
    mutate(cycle_name = fct_relevel(cycle_name_2, "ST1", "ST2", "SA1","SA2","SA3" )) %>%
    ggplot(aes(x=NMDS1, y= NMDS2, colour = depth, shape = vial_type)) +
    geom_point() +
  facet_grid(cycle_exp~sorting_population, scales = "free_x")+
  ggtitle("NMDS sorted euks by depth and vial") 
p

pdf(here("phyloseq", "paper_c14", "20220411", "sorted_euks_nmds_depth and vial.pdf"), height=6, width=9) ; plot(p) ; dev.off()


```


# compare against normalising the data - looks the same (ignore this)
```{r}
#normalise ASVs
asv_norm <- t(otu_sub)
taxa <- data.frame(tax_table(ps_merge))
samples <- data.frame(sample_data(ps_merge))

taxa <- as.matrix(taxa)

otus= otu_table(asv_norm, taxa_are_rows = TRUE)
taxa = tax_table(taxa)
samples = sample_data(samples)
ps <- merge_phyloseq(otus, taxa, samples)

ps_norm <- phyloseq_normalize_median(ps)
otu_norm <- psotu2veg(ps_norm)

set.seed(4)
dist_norm <- vegdist(otu_norm, method = "bray")
nmds_norm <- metaMDS(dist_norm)


nmds_sample_sub_norm <- scores(nmds_norm) %>%
    as.tibble(rownames="sample_name") %>%
    inner_join(., sample, by = "sample_name")

p <- nmds_sample_sub_norm %>%
    mutate(cycle_name = fct_relevel(cycle_name_2, "ST1", "ST2", "SA1","SA2","SA3" )) %>%
    ggplot(aes(x=NMDS1, y= NMDS2, colour = cycle_name, shape = depth)) +
    geom_point() +
    scale_colour_manual(values = colours_cycle) +
  ggtitle("NMDS normalise reads")
p
```

#include CTD data.
```{r}
source(here("phyloseq", "init_files", "read_18s_filt_sort.R"))

ps_filt_sort <- phyloseq_assign_trophic(ps_filt_sort) %>%
    subset_taxa(trophic_mode_2 != "heterotrophic") #remove heterotrophs

#select only surface and dcm samples
sample <- data.frame(ps_filt_sort@sam_data) %>%
  filter(sample %in% c("SUR", "DCM"))

otu <- psotu2veg(ps_filt_sort)

#sample list of filt samples
sample_list <- sample %>%
  select(sample_name)

# subset removed groups and samples from the plot
otu_sub <- otu %>%
  as.data.frame() 
otu_sub$sample_name <-rownames(otu_sub)

otu_sub <- otu_sub %>% 
  inner_join(., sample_list) %>% #select only corresponding depths for filt samples
  filter(!(sample_name %in% c("nano-12", "pico-15", "pico-01-30"))) #remove few sorted samples

rownames(otu_sub) <- otu_sub$sample_name

otu_sub <- otu_sub %>%
  select(-sample_name) %>%
  as.matrix()

```

#checks before plotting - ok, ignore for now.
```{r}
#to check for every sample number of reads and if all asvs have reads -  ok, ignore code now.

otu_long <- otu_sub %>%
  as.data.frame() 
otu_long$file_code <-rownames(otu_long)
otu_long <- otu_long %>%
    pivot_longer(cols = !file_code, names_to = "otu", values_to = "reads") %>%
   # group_by(file_code)  %>%
  #  summarise(sum = sum(reads)) %>%
    group_by(otu) %>%
    summarise(sum = sum(reads))

```

```{r}
set.seed(1)
dist <- avgdist(otu_sub, sample = 8000, dmethod = "bray") #rarify to control for uneven number of reads per sample.
nmds_sub<- metaMDS(dist)

nmds_sample_sub <- scores(nmds_sub) %>%
    as.tibble(rownames="sample_name") %>%
    inner_join(., sample, by = "sample_name")

goodness(nmds_sub)
stressplot(nmds_sub)

nmds_sub$stress

#plot overall
p <- nmds_sample_sub %>%
    mutate(cycle_name = fct_relevel(cycle_name_2, "ST1", "ST2", "SA1","SA2","SA3" )) %>%
    mutate(sample = fct_relevel(sample, "SUR", "DCM") ) %>%
    ggplot(aes(x=NMDS1, y= NMDS2, colour = cycle_name, shape = sample)) +
    geom_point() +
    scale_colour_manual(values = colours_cycle) 
p

pdf(here("phyloseq", "paper_c14", "20220417", "nmds_sorted_filt_2.pdf"), height=6, width=9) ; plot(p) ; dev.off()


```