---
title: "18S barchart"
author: "Denise Ong"
date: "11/6/2022"
output: html_document
---

Plotting a median range chart and heatmap using median relative abundance. 
Only using light incubated samples. Initials and dark are controls. already checked previously.
# Initialize
```{r, echo=TRUE, message=FALSE, warning=FALSE}
library(here)
source(here("phyloseq", "init_files", "init.R")) #for libraries
source(here("phyloseq", "init_files", "init_markdown.R")) #for markdown libraries
source(here("phyloseq", "init_files", "define_functions.R")) #functions to 1.transform to dataframe, 2.normalise and 3.create treemaps 
source(here("phyloseq", "init_files", "colours_18sv4.R")) # predetermined colours
#source(here("phyloseq", "init_files", "colours_random.R")) #random colours
source(here("phyloseq", "trophic_mode", "trophic_function.R"))  # Function to assign trophic function for phyloseq objects. 2 
source(here("phyloseq", "init_files", "top_taxa_func.R")) #Function to select top taxa
library(ggtext)
library(ggforestplot)
# Colours
colours_cycle <- readxl::read_excel(here("phyloseq", "init_files","colours_cycle.xlsx"))
colours_cycle <- structure(colours_cycle$colour,.Names=colours_cycle$cycle)
```

# Sorted pico
## Read data
remove heterotrophs and 2 samples 
also removing 
```{r}
source(here("phyloseq", "init_files", "read_pico.R"))

#Assign trophic mode and remove heterotrophics that are not dinophyceae
ps_pico <- phyloseq_assign_trophic(ps_pico) %>%
    subset_taxa(trophic_mode_2 != "heterotrophic")

# Do normalizations and transformation
ps_pico <- phyloseq_normalize_median(ps_pico)
long_pico <- phyloseq_transform_to_long(ps_pico) %>%
    filter(!(file_code %in% c("pico-01-30", "pico-15"))) %>% #filter problem samples
    filter(!(cycle_exp =="SA1_2" & SAMPLE == "DCM")) %>% #only intials in these experiments
    filter(!(cycle_exp == "SA1_1" & SAMPLE == "DCM")) 

#create levels for factors
long_pico <- long_pico %>%
    mutate(cycle_exp = fct_relevel(cycle_exp, "ST1_7", "ST1_8", "ST2_9", "ST2_10", "SA1_1", "SA1_2", "SA2_3", "SA2_4", "SA2_5", "SA2_6", "SA3_11")) %>%
    mutate(trophic_mode_2 = fct_relevel(trophic_mode_2, "photosynthetic", "mixotrophic", "dinophyceae")) %>%
    mutate(cycle_name_2 = fct_relevel(cycle_name_2, "ST1", "ST2", "SA1", "SA2", "SA3")) %>%
    mutate(vial_type = fct_relevel(vial_type, "initial", "light", "dark")) %>%
    mutate(depth_cat = fct_relevel(SAMPLE, "SUR", "DCM"))

```

## dataframe prep
```{r}

long_pico_1 <- long_pico %>%
   #  ps_top_asv(10) %>% #subset top 10 ASVs per sample
     filter(vial_type=="light")


# pivot longer dataframe
otu_rel_abund <- long_pico_1 %>%
    group_by(file_code) %>%
    mutate(rel_abund = n_reads / sum(n_reads)) %>%
    ungroup() %>%
    select(-n_reads) %>%
    pivot_longer(cols = asv_code:species,
                 names_to = "level",
                 values_to = "taxon") 
# for log scale
nseqs_per_sample <- otu_rel_abund %>%
    group_by(file_code) %>%
    summarise(N=sum(rel_abund)) %>%
    count(N) %>%
    pull(N)
# 
# lod <- 1/nseqs_per_sample

```

## Class - Median range plot -ignore
```{r, eval=FALSE}
#group by the level you want to plot, calculate percent abundance
taxon_rel_abund <- otu_rel_abund %>%
    filter(level == "class") %>% #choose level here 
        mutate(taxon = case_when(taxon == 'Mamiellophyceae' ~ 'Mamiellophyceae',
                                 taxon == 'Chloropicophyceae' ~ 'Chloropicophyceae',
                                 taxon == 'Prymnesiophyceae' ~ 'Prymnesiophyceae',
                              #   taxon == 'Bacillariophyta' ~ 'Bacillariophyta',
                                 #taxon == 'Bolidophyceae' ~ 'Bolidophyceae',
                                 taxon == 'Pelagophyceae' ~'Pelagophyceae',
                                 taxon == "Chrysophyceae" ~ 'Chrysophyceae',
                              #   taxon == 'Dinophyceae' ~'Dinophyceae',
                                 TRUE~ 'Other')) %>%
     group_by(cycle_name_2, file_code, taxon) %>%
    summarise(rel_abund = 100*sum(rel_abund), .groups = "drop") %>%
    pivot_wider(names_from = taxon, values_from = rel_abund) %>% #from here is to replace any file_code with NA as 0
    replace(., is.na(.), 0) %>%
    pivot_longer(cols=Chrysophyceae:Chloropicophyceae,
                 names_to = "taxon",
                 values_to = "rel_abund")


#test <- taxon_rel_abund %>%
#    group_by(file_code) %>%
#    summarise(total = sum(rel_abund))

# Range plot
p <- taxon_rel_abund %>%
    group_by(file_code, cycle_name_2, taxon) %>%
    summarise(rel_abund = sum(rel_abund), 
              .groups = "drop") %>%
    mutate(taxon=fct_relevel(taxon,
                             "Other",
                        #     "Dinophyceae",
                         #    "Bacillariophyta",
                             "Chrysophyceae",
                             "Pelagophyceae",
                             "Prymnesiophyceae",
                             "Chloropicophyceae",
                             "Mamiellophyceae")) %>%
    mutate(cycle_name_2 = fct_relevel(cycle_name_2, "SA3", "SA2", "SA1", "ST2", "ST1")) %>%
    # mutate(rel_abund = if_else(rel_abund == 0,
    #                            2/3 *lod,
    #                            rel_abund)) %>%
    ggplot(aes(x=rel_abund, y=taxon, colour = cycle_name_2)) +
    #geom_point()+
    #geom_jitter(position = position_jitterdodge(jitter.width = 0.4, dodge.width=0.8), pch=21, stroke=0, size=1.8)+
     stat_summary(fun.data=median_hilow, geom = "pointrange", #75% confidence interval
                  fun.args = list(conf.int=0.75),
                  position=position_dodge(width=0.4),
                  size=0.25) +
    # stat_summary(fun.data = mean_se, geom = "pointrange",
    #              position=position_dodge(width=0.4),
    #               size=0.25)+
    scale_colour_manual(values=colours_cycle, limits = force)+
    theme_classic()+
    xlab("Relative abundance (%)") +
    ylab("Class") +
    theme(legend.position = c(0.9,0.15),
          legend.background = element_rect(colour="black", fill = NA),
          panel.grid.major.x = element_line(colour="lightgrey", size=0.25))+
    guides(colour=guide_legend(title="Cycle")) #+
#    coord_trans(x="log10") +
#    scale_x_continuous(breaks=c(0.1, 1, 10, 50, 100))


p
#pdf("sorted_pico_class_no log.pdf", height=6, width=5) ; plot(p) ; dev.off()

```

## Class - median for each cycle more than >5%
```{r}
taxon_rel_abund <- otu_rel_abund %>%
    filter(level == "class") %>% #choose level here 
    group_by(water_mass, cycle_name_2, file_code, taxon) %>%
    summarise(rel_abund = 100*sum(rel_abund)) %>%
    pivot_wider(names_from = taxon, values_from = rel_abund) %>% #from here is to replace any file_code with NA as 0
    replace(., is.na(.), 0) %>%
    pivot_longer(cols= !water_mass:file_code,
                 names_to = "taxon",
                 values_to = "rel_abund")
## summary for results
summary <- taxon_rel_abund %>%
  filter(taxon %in% c("Mamiellophyceae", "Chrysophyceae","Pelagophyceae","Prymnesiophyceae", "Chloropicophyceae")) %>%
  filter(cycle_name_2 != "SA1") %>%
  group_by(taxon, water_mass) %>%
  summarise(median = median(rel_abund),
            mean = mean(rel_abund))
summary
```

```{r}
taxon_pool <- taxon_rel_abund %>%
    group_by(cycle_name_2, taxon) %>%
    summarise(median = median(rel_abund)) %>% #calculate median proportion of reads per sp per cycle
    group_by(taxon) %>%
    summarise(pool = max(median) <= 5, #species with a max median proportion of reads in any cycle < 5% is under Others
              median = max(median),
              .groups = "drop")

library(ggforestplot)
# median range plot
 p1.1<- inner_join(taxon_rel_abund, taxon_pool, by="taxon") %>%
    mutate(taxon=if_else(pool, "Other", taxon)) %>%
    group_by(file_code, cycle_name_2, taxon) %>%
    summarise(rel_abund = sum(rel_abund), 
              .groups = "drop") %>%
    mutate(cycle_name_2 = fct_relevel(cycle_name_2, "SA3", "SA2", "SA1", "ST2", "ST1")) %>%
    mutate(taxon=fct_reorder(taxon, rel_abund)) %>%
    ggplot(aes(x=rel_abund, y=taxon, colour = cycle_name_2)) +
      geom_stripes(odd = "#FFFFFFFF", even = "#EFEFEFFF") +
   stat_summary(fun.data=median_hilow, geom = "pointrange", #75% confidence interval
                  fun.args = list(conf.int=0.75),
                  position=position_dodge(width=0.4),
                  size=0.25) +
     scale_colour_manual(values=colours_cycle, limits = force)  +
    theme_classic()+
    xlab("Relative abundance (%)") +
    ylab(NULL) +
  theme_classic() +
    theme(legend.position = c(0.85,0.2),
          #legend.background =element_blank(),
          panel.grid.major.x = element_line(colour="lightgrey", size=0.25),
          legend.text = element_text(size = 8),
          legend.title = element_text(size = 8)) +
   xlim(0, 100) +
   guides(colour=guide_legend(title="Cycle")) 

 p
 
 # pdf("sorted_pico_class_test.pdf", height=5, width=5) ; plot(p) ; dev.off()
 
 #heatmap
 p1 <- inner_join(taxon_rel_abund, taxon_pool, by="taxon") %>%
    mutate(taxon=if_else(pool, "Other", taxon)) %>%
    group_by(file_code, cycle_name_2, taxon) %>%
    summarise(rel_abund = sum(rel_abund), 
              .groups = "drop") %>%
  group_by(cycle_name_2, taxon) %>%
  summarise(rel_abund = median(rel_abund)) %>% # for median reads per taxon per cycle
     mutate(taxon=fct_relevel(taxon,
                             "Other",
                             "Chrysophyceae",
                             "Pelagophyceae",
                             "Prymnesiophyceae",
                             "Mamiellophyceae")) %>%
  ggplot(aes(x=cycle_name_2, y = taxon, fill =rel_abund)) +
    geom_tile() +
   scale_fill_distiller(type = "seq", 
                        palette = "BuGn",
                        name = "Median<br>Relative<br>Abund (%)",
                       expand = c(0,0),
                       trans = "log",
        na.value = "#F7FCFD",
        direction = 1,
        breaks = c(0, 0.1, 1, 10, 100),
        limits = c(NA,100))+
  # scale_fill_gradient2(
  #       name = "Median<br>Relative<br>Abund (%)",
  #       low = "white",
  #       mid = "#C3D7A4",
  #       high = "#52854C",
  #       space = "Lab",
  #       trans = "log",
  #       limits = c(NA,100),
  #       expand = c(0,0),
  #       breaks = c(0, 0.1, 1, 10, 100),
  #       na.value = "white",
  #       guide = "colourbar",
  #       aesthetics = "fill") +
  theme_minimal() +
  theme(text=element_text(size=10),
              strip.text.y.left = element_text(angle = 0),
              panel.spacing.x = unit(0,"lines"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              strip.background = element_blank(),
          axis.line = element_blank(),
          axis.ticks = element_blank(),
          legend.title = element_markdown()) +
  scale_x_discrete(position = "top") +
  xlab(NULL)+
  ylab(NULL)
p1
# pdf("20221130_pico_class_heatmap.pdf", height=5, width=5) ; plot(p) ; dev.off()

```


## species - median range plot - ignore
```{r, eval=FALSE}
taxon_rel_abund <- otu_rel_abund %>%
    filter(level == "species") %>% #choose level here 
        mutate(taxon = case_when(taxon == 'Bathycoccus_prasinos' ~ 'Bathycoccus_prasinos',
                                 taxon == 'Micromonas_commoda_A2' ~ 'Micromonas_commoda_A2',
                                 taxon == 'Ostreococcus_lucimarinus' ~ 'Ostreococcus_lucimarinus',
                                 taxon == 'Micromonas_bravo_B1' ~ 'Micromonas_bravo_B1',
                                 taxon == 'Phaeocystis_antarctica' ~ 'Phaeocystis_antarctica',
                                 taxon == 'Chrysochromulina_sp.' ~'Chrysochromulina_sp.',
                                 taxon == "Aureococcus_anophagefferens" ~ 'Aureococcus_anophagefferens',
                                 taxon == 'Pelagomonadales_clade_B2_sp.' ~'Pelagomonadales_clade_B2_sp.',
                                 taxon == 'Pelagomonas_calceolata' ~'Pelagomonas_calceolata',
                                 taxon == 'Chrysophyceae_Clade-I_X_sp.' ~ 'Chrysophyceae_Clade-I_X_sp.',
                                # taxon == 'Chloroparvula_pacifica' ~'Chloroparvula_pacifica',
                                 TRUE~ 'Other')) %>%
     group_by(cycle_name_2, file_code, taxon) %>%
    summarise(rel_abund = 100*sum(rel_abund), .groups = "drop") %>%
    pivot_wider(names_from = taxon, values_from = rel_abund) %>% #from here is to replace any file_code with NA as 0
    replace(., is.na(.), 0) %>%
    pivot_longer(cols=Aureococcus_anophagefferens:Pelagomonas_calceolata,
                 names_to = "taxon",
                 values_to = "rel_abund")

#get class name for facet grid grouping with class
class_df <- long_pico_1 %>%
    select(class, species) %>%
    distinct()# %>%
    # #filter(species %in% c(#"Chloroparvula_pacifica",
    #                          "Chrysophyceae_Clade-I_X_sp.",
    #                          "Pelagomonas_calceolata",
    #                          "Pelagomonadales_clade_B2_sp.",
    #                          "Aureococcus_anophagefferens",
    #                          "Chrysochromulina_sp.",
    #                          "Phaeocystis_antarctica",
    #                          "Micromonas_bravo_B1",
    #                     "Ostreococcus_lucimarinus",
    #                     "Micromonas_commoda_A2",
    #                     "Bathycoccus_prasinos"))


p <- taxon_rel_abund %>%
    group_by(file_code, cycle_name_2, taxon) %>%
    summarise(rel_abund = sum(rel_abund), 
              .groups = "drop") %>%
    full_join(class_df , by=c("taxon" = "species")) %>%
    mutate(taxon=fct_relevel(taxon,
                             "Other",
                             #"Chloroparvula_pacifica",
                             "Chrysophyceae_Clade-I_X_sp.",
                             "Pelagomonas_calceolata",
                             "Pelagomonadales_clade_B2_sp.",
                             "Aureococcus_anophagefferens",
                             "Chrysochromulina_sp.",
                             "Phaeocystis_antarctica",
                            # "Micromonas_bravo_B1",
                        "Ostreococcus_lucimarinus",
                        "Micromonas_commoda_A2",
                        "Bathycoccus_prasinos")) %>%
    mutate(class = fct_relevel(class,
                               "Mamiellophyceae",
                               "Prymnesiophyceae",
                               "Pelagophyceae",
                               "Chrysophyceae",
                              # "Chloropicophyceae",
                               "NA")) %>%
    mutate(cycle_name_2 = fct_relevel(cycle_name_2, "SA3", "SA2", "SA1", "ST2", "ST1")) %>%
    # mutate(rel_abund = if_else(rel_abund == 0,
    #                            2/3 *lod,
    #                            rel_abund)) %>%
    ggplot(aes(x=rel_abund, y=taxon, colour = cycle_name_2)) +
    #geom_point()+
    #geom_jitter(position = position_jitterdodge(jitter.width = 0.4, dodge.width=0.8), pch=21, stroke=0, size=1.8)+
     stat_summary(fun.data=median_hilow, geom = "pointrange", #75% confidence interval
                  fun.args = list(conf.int=0.75),
                  position=position_dodge(width=0.4),
                  size=0.25) +
    # stat_summary(fun.data = mean_se, geom = "pointrange",
    #              position=position_dodge(width=0.4),
    #               size=0.25)+
    scale_colour_manual(values=colours_cycle, limits = force)+
    theme_classic()+
    xlab("Relative abundance (%)") +
    ylab(NULL) +
    facet_grid(class~., scales = "free_y", space = "free", switch = "y") +
    theme(legend.position = c(0.9,0.15),
          legend.background = element_rect(colour="black", fill = NA),
          panel.grid.major.x = element_line(colour="lightgrey", size=0.25),
          strip.placement = "outside",
          strip.text.y.left = element_text(angle = 0))+
    guides(colour=guide_legend(title="Cycle"))
            #    coord_trans(x="log10") +
#    scale_x_continuous(breaks=c(0.1, 1, 10, 50, 100))


p
#pdf("sorted_pico_species_no log.pdf", height=7, width=8) ; plot(p) ; dev.off()
```

## species - select median >5% for any cycle
```{r}
taxon_rel_abund <- otu_rel_abund %>%
    filter(level == "species") %>% #choose level here 
    group_by(cycle_name_2, file_code, taxon) %>%
    summarise(rel_abund = 100*sum(rel_abund)) %>%
    pivot_wider(names_from = taxon, values_from = rel_abund) %>% #from here is to replace any file_code with NA as 0
    replace(., is.na(.), 0) %>%
    pivot_longer(cols= !cycle_name_2:file_code,
                 names_to = "taxon",
                 values_to = "rel_abund")

## summary for results
summary <- taxon_rel_abund %>%
  filter(taxon %in% c("Mamiellophyceae", "Chrysophyceae","Pelagophyceae","Prymnesiophyceae")) %>%
  filter(cycle_name_2 == "SA1") %>%
  group_by(taxon) %>%
  summarise(median = median(rel_abund),
            mean = mean(rel_abund))
summary
```

```{r}
taxon_pool <- taxon_rel_abund %>%
    group_by(cycle_name_2, taxon) %>%
    summarise(median = median(rel_abund)) %>% #calculate median proportion of reads per sp per cycle
    group_by(taxon) %>%
    summarise(pool = max(median) <= 5, #species with a max median proportion of reads in any cycle < 5% is under Others
              median = max(median),
              .groups = "drop")

test <- inner_join(taxon_rel_abund, taxon_pool, by="taxon") %>%
    mutate(taxon=if_else(pool, "Other", taxon)) %>% 
    group_by(file_code, cycle_name_2, taxon) %>%
    summarise(rel_abund = sum(rel_abund), #add Others taxon together
              .groups = "drop") %>%
  group_by(taxon, cycle_name_2) %>%
  summarise(median = median(rel_abund),
            mean = mean(rel_abund))
test

#get class name for facet grid grouping with class
class_df <- long_pico_1 %>%
    select(class, species) %>%
    distinct()

#median range plot
 p2.2<- inner_join(taxon_rel_abund, taxon_pool, by="taxon") %>%
    mutate(taxon=if_else(pool, "Other", taxon)) %>% 
    group_by(file_code, cycle_name_2, taxon) %>%
    summarise(rel_abund = sum(rel_abund), #add Others taxon together
              .groups = "drop") %>%
   left_join(class_df , by=c("taxon" = "species")) %>% # get class levels for sp.
   mutate(cycle_name_2 = fct_relevel(cycle_name_2, "SA3", "SA2", "SA1", "ST2", "ST1")) %>%
  # mutate(class = fct_reorder(class, rel_abund)) %>%
      mutate(class = fct_relevel(class,
                               "Mamiellophyceae",
                               "Prymnesiophyceae",
                               "Pelagophyceae",
                               "Chrysophyceae",
                              # "Chloropicophyceae",
                               "NA")) %>%
   ggplot(aes(x=rel_abund, y=taxon, colour = cycle_name_2)) +
      geom_stripes(even = "#FFFFFFFF", odd = "#EFEFEFFF") +
     stat_summary(fun.data=median_hilow, geom = "pointrange", #75% confidence interval
                  fun.args = list(conf.int=0.75),
                  position=position_dodge(width=0.5),
                  size=0.25) +
    scale_colour_manual(values=colours_cycle, limits = force)+
    theme_classic()+
    xlab("Relative abundance (%)") +
    ylab(NULL) +
   theme_classic() +
    theme(legend.position = c(0.9,0.15),
          #legend.background =element_blank(),
          panel.grid.major.x = element_line(colour="lightgrey", size=0.25),
          strip.placement = "outside",
          strip.text.y.left = element_text(angle = 0),
          strip.background = element_blank()) +
    guides(colour=guide_legend(title="Cycle")) +
       facet_grid(class~., scales = "free_y", space = "free", switch = "y") 

p
# pdf("sorted_pico_species_median.pdf", height=7, width=8) ; plot(p) ; dev.off()

#heatmap
p2 <- inner_join(taxon_rel_abund, taxon_pool, by="taxon") %>%
    mutate(taxon=if_else(pool, "Other", taxon)) %>% 
    group_by(file_code, cycle_name_2, taxon) %>%
      summarise(rel_abund = sum(rel_abund), #add Others taxon together
              .groups = "drop") %>%
    group_by(cycle_name_2, taxon) %>%
   summarise(rel_abund = median(rel_abund)) %>% # for median reads per taxon per cycle
   left_join(class_df , by=c("taxon" = "species")) %>% # get class levels for sp.
      mutate(class = fct_relevel(class,
                               "Mamiellophyceae",
                               "Prymnesiophyceae",
                               "Pelagophyceae",
                               "Chrysophyceae",
                              # "Chloropicophyceae",
                               "NA")) %>%
     ggplot(aes(x=cycle_name_2, y = taxon, fill =rel_abund)) +
    geom_tile() +
  #geom_text(aes(label=format(round(rel_abund,1), nsmall = 1)), size=5)+
   scale_fill_distiller(type = "seq", 
                        palette = "BuGn",
                        name = "Median<br>Relative<br>Abund (%)",
                       expand = c(0,0),
                       trans = "log",
        na.value = "#F7FCFD",
        direction = 1,
        breaks = c(0, 0.1, 1, 10, 100),
        limits = c(NA,100))+
    # scale_fill_gradient2(
    #     name = "Median<br>Relative<br>Abund (%)",
    #     low = "white",
    #     mid = "#C3D7A4",
    #     high = "#52854C",
    #     space = "Lab",
    #     trans = "log",
    #     limits = c(NA, 100),
    #     expand = c(0,0),
    #     breaks = c(0, 0.1, 1, 10, 100),
    #     na.value = "white",
    #     guide = "colourbar",
    #     aesthetics = "fill") +
  theme_classic()+
    theme(text=element_text(size=10),
              strip.text.y.left = element_text(angle = 0),
              panel.spacing.y = unit(0,"lines"),
              # panel.grid.major = element_blank(),
              # panel.grid.minor = element_blank(),
              strip.background = element_blank(),
          axis.line = element_blank(),
          axis.ticks = element_blank(),
          legend.title = element_markdown(),
          strip.placement = "outside") +
  scale_x_discrete(position = "top") +
  xlab(NULL)+
  ylab(NULL) +
    facet_grid(class~., scales = "free_y", space = "free", switch = "y") 
    
p2
# pdf("20221130_pico_species_heatmap.pdf", height=5, width=7) ; plot(p) ; dev.off()

```

```{r}
#no. of samples per cycle
sample_no <- long_pico_1 %>%
  select(cycle_name_2, file_code) %>%
  distinct() %>%
  group_by(cycle_name_2) %>%
  summarise(n = n())
sample_no
```


# Sorted nano
## Read data
```{r}
source(here("phyloseq", "init_files", "read_nano.R"))

#assign trophic mode and remove heterotrophics that are not dinophyceae
ps_nano <- phyloseq_assign_trophic(ps_nano)%>%
    subset_taxa(trophic_mode_2 != "heterotrophic")

# Do normalizations and long form, remove few samples.
ps_nano <- phyloseq_normalize_median(ps_nano)
long_nano <- phyloseq_transform_to_long(ps_nano) %>%
    filter(!(file_code %in% c("nano-12"))) 

#order all the factors
long_nano <- long_nano %>%
  mutate(cycle_exp = fct_relevel(cycle_exp, "ST1_7", "ST1_8", "ST2_9", "ST2_10", "SA1_1", "SA1_2", "SA2_3", "SA2_4", "SA2_5", "SA2_6", "SA3_11")) %>%
  mutate(trophic_mode = fct_relevel(trophic_mode, "photosynthetic", "mixotrophic"))  %>%
  mutate(trophic_mode_2 = fct_relevel(trophic_mode_2, "photosynthetic", "mixotrophic",  "dinophyceae"))  %>%
  mutate(cycle_name_2 =  fct_relevel(cycle_name_2,"ST1", "ST2", "SA1", "SA2", "SA3"))  %>%
  mutate(vial_type = fct_relevel(vial_type, "initial", "light", "dark"))
```
## prepare data frame
```{r}

long_nano_1 <- long_nano %>%
   #  ps_top_asv(10) %>% #subset top 10 ASVs per sample
     filter(vial_type=="light")


# pivot longer dataframe
otu_rel_abund <- long_nano_1 %>%
    group_by(file_code) %>%
    mutate(rel_abund = n_reads / sum(n_reads)) %>%
    ungroup() %>%
    select(-n_reads) %>%
    pivot_longer(cols = asv_code:species,
                 names_to = "level",
                 values_to = "taxon") 
```

## class
```{r}
taxon_rel_abund <- otu_rel_abund %>%
    filter(level == "class") %>% #choose level here 
    group_by(cycle_name_2, file_code, taxon) %>%
    summarise(rel_abund = 100*sum(rel_abund)) %>%
    pivot_wider(names_from = taxon, values_from = rel_abund) %>% #from here is to replace any file_code with NA as 0
    replace(., is.na(.), 0) %>%
    pivot_longer(cols= !cycle_name_2:file_code,
                 names_to = "taxon",
                 values_to = "rel_abund")

## summary for results
summary <- taxon_rel_abund %>%
  filter(taxon %in% c("Dinophyceae", "Prymnesiophyceae", "Bacillariophyta", "Cryptophyceae", "Haptophyta_Clade_HAP3")) %>%
  # filter(cycle_name_2 %in% c("SA1")) %>%
  group_by(taxon, cycle_name_2) %>%
  summarise(median = median(rel_abund),
            mean = mean(rel_abund))
summary
```

```{r}
taxon_pool <- taxon_rel_abund %>%
    group_by(cycle_name_2, taxon) %>%
    summarise(median = median(rel_abund)) %>% #calculate median proportion of reads per sp per cycle
    group_by(taxon) %>%
    summarise(pool = max(median) <= 5, #species with a max median proportion of reads in any cycle < 5% is under Others
              median = max(median),
              .groups = "drop")

#get class name for facet grid grouping with class
# class_df <- long_nano_1 %>%
#     select(class, order) %>%
#     distinct()


#median range plot
 p<- inner_join(taxon_rel_abund, taxon_pool, by="taxon") %>%
    mutate(taxon=if_else(pool, "Other", taxon)) %>% 
    group_by(file_code, cycle_name_2, taxon) %>%
    summarise(rel_abund = sum(rel_abund), #add Others taxon together
              .groups = "drop") %>%
   # left_join(class_df , by=c("taxon" = "order")) %>% # get class levels for order
   mutate(cycle_name_2 = fct_relevel(cycle_name_2, "SA3", "SA2", "SA1", "ST2", "ST1")) %>%
  # mutate(class = fct_reorder(class, rel_abund)) %>%
      # mutate(class = fct_relevel(class,
      #                          "Mamiellophyceae",
      #                          "Prymnesiophyceae",
      #                          "Pelagophyceae",
      #                          "Chrysophyceae",
      #                         # "Chloropicophyceae",
      #                          "NA")) %>%
   ggplot(aes(x=rel_abund, y=taxon, colour = cycle_name_2)) +
      geom_stripes(even = "#FFFFFFFF", odd = "#EFEFEFFF") +
     stat_summary(fun.data=median_hilow, geom = "pointrange", #75% confidence interval
                  fun.args = list(conf.int=0.75),
                  position=position_dodge(width=0.5),
                  size=0.25) +
    scale_colour_manual(values=colours_cycle, limits = force)+
    theme_classic()+
    xlab("Relative abundance (%)") +
    ylab(NULL) +
   theme_classic() +
    theme(#legend.position = c(0.9,0.15),
          #legend.background =element_blank(),
          panel.grid.major.x = element_line(colour="lightgrey", size=0.25),
          strip.placement = "outside",
          strip.text.y.left = element_text(angle = 0),
          strip.background = element_blank()) +
   xlim(0,50)+
    guides(colour=guide_legend(title="Cycle")) 
       # facet_grid(class~., scales = "free_y", space = "free", switch = "y") 

p
# pdf("nano_order_median.pdf", height=7, width=8) ; plot(p3.1) ; dev.off()

```

## Order median range plot
select groups
```{r, eval=FALSE}
#group by the level you want to plot, calculate percent abundance
taxon_rel_abund <- otu_rel_abund %>%
    filter(level == "order") %>% #choose level here 
        mutate(taxon = case_when(taxon == 'Cryptomonadales' ~ 'Cryptomonadales',
                                 taxon == 'Dinophyceae_X' ~ 'Dinophyceae_X',
                                 taxon == 'Suessiales' ~ 'Suessiales',
                                 taxon == 'Gymnodiniales' ~ 'Gymnodiniales',
                                 taxon == 'Bolidophyceae' ~ 'Bolidophyceae',
                                 taxon == 'Peridiniales' ~'Peridiniales',
                                 taxon == 'Prorocentrales' ~ 'Prorocentrales',
                                 taxon == 'Isochrysidales' ~'Isochrysidales',
                                 taxon == 'Prymnesiales' ~'Prymnesiales',
                                 taxon == 'Phaeocystales' ~'Phaeocystales',
                                 taxon == 'Bacillariophyta_X' ~'Bacillariophyta_X',
                                 taxon == 'Chlorellales' ~'Chlorellales',
                                 TRUE~ 'Other')) %>%
     group_by(cycle_name_2, file_code, taxon) %>%
    summarise(rel_abund = 100*sum(rel_abund), .groups = "drop") %>%
    pivot_wider(names_from = taxon, values_from = rel_abund) %>% #from here is to replace any file_code with NA as 0
    replace(., is.na(.), 0) %>%
    pivot_longer(cols=Bacillariophyta_X:Chlorellales,
                 names_to = "taxon",
                 values_to = "rel_abund")
#get class name for facet grid grouping with class
class_df <- long_nano_1 %>%
    select(class, order) %>%
    distinct() %>%
    filter(order %in% c("Cryptomonadales",
                        "Dinophyceae_X",
                        "Suessiales",
                        "Gymnodiniales",
                        "Bolidophyceae",
                        "Peridiniales",
                        "Prorocentrales",
                        "Isochrysidales",
                        "Prymnesiales",
                        "Phaeocystales",
                        "Bacillariophyta_X",
                        "Chlorellales"
                        ))

p <- taxon_rel_abund %>%
    group_by(file_code, cycle_name_2, taxon) %>%
    summarise(rel_abund = sum(rel_abund), 
              .groups = "drop") %>%
    # mutate(taxon=fct_relevel(taxon,
    #                          "Other",
    #                     #     "Dinophyceae",
    #                      #    "Bacillariophyta",
    #                          "Chrysophyceae",
    #                          "Pelagophyceae",
    #                          "Prymnesiophyceae",
    #                          "Chloropicophyceae",
    #                          "Mamiellophyceae")) %>%
    mutate(cycle_name_2 = fct_relevel(cycle_name_2, "SA3", "SA2", "SA1", "ST2", "ST1")) %>%
    # mutate(rel_abund = if_else(rel_abund == 0,
    #                            2/3 *lod,
    #                            rel_abund)) %>%
    full_join(class_df , by=c("taxon" = "order")) %>%
    ggplot(aes(x=rel_abund, y=taxon, colour = cycle_name_2)) +
    #geom_point()+
    #geom_jitter(position = position_jitterdodge(jitter.width = 0.4, dodge.width=0.8), pch=21, stroke=0, size=1.8)+
     stat_summary(fun.data=median_hilow, geom = "pointrange", #75% confidence interval
                  fun.args = list(conf.int=0.50),
                  position=position_dodge(width=0.4),
                  size=0.25) +
    # stat_summary(fun.data = mean_se, geom = "pointrange",
    #              position=position_dodge(width=0.4),
    #               size=0.25)+
    scale_colour_manual(values=colours_cycle, limits = force)+
    theme_classic()+
    xlab("Relative abundance (%)") +
    ylab(NULL) +
    facet_grid(class~., scales = "free_y", space = "free", switch = "y") +
    theme(legend.position = c(0.9,0.15),
          legend.background = element_rect(colour="black", fill = NA),
          panel.grid.major.x = element_line(colour="lightgrey", size=0.25),
          strip.placement = "outside",
          strip.text.y.left = element_text(angle = 0))+
    guides(colour=guide_legend(title="Cycle")) #+
#    coord_trans(x="log10") +
#    scale_x_continuous(breaks=c(0.1, 1, 10, 50, 100))


p
 pdf("sorted_nano_order.pdf", height=7.5, width=6) ; plot(p) ; dev.off()
```


## Order - select above 5%
```{r}
taxon_rel_abund <- otu_rel_abund %>%
    filter(level == "order") %>% #choose level here 
    group_by(water_mass, cycle_name_2, file_code, taxon) %>%
    summarise(rel_abund = 100*sum(rel_abund)) %>%
    pivot_wider(names_from = taxon, values_from = rel_abund) %>% #from here is to replace any file_code with NA as 0
    replace(., is.na(.), 0) %>%
    pivot_longer(cols= !water_mass:file_code,
                 names_to = "taxon",
                 values_to = "rel_abund")

summary <- taxon_rel_abund %>%
  filter(taxon %in% c("Bacillariophyta_X", "Cryptomonadales",  "Haptophyta_Clade_HAP3_X")) %>%
  group_by(taxon, cycle_name_2) %>%
  summarise(median = median(rel_abund),
            mean = mean(rel_abund))
summary

```

```{r}
taxon_pool <- taxon_rel_abund %>%
    group_by(cycle_name_2, taxon) %>%
    summarise(median = median(rel_abund)) %>% #calculate median proportion of reads per sp per cycle
    group_by(taxon) %>%
    summarise(pool = max(median) <= 5, #species with a max median proportion of reads in any cycle < 5% is under Others
              median = max(median),
              .groups = "drop")

#get class name for facet grid grouping with class
class_df <- long_nano_1 %>%
    select(class, order) %>%
    distinct()

#median range plot
 p3.1<- inner_join(taxon_rel_abund, taxon_pool, by="taxon") %>%
    mutate(taxon=if_else(pool, "Other", taxon)) %>% 
    group_by(file_code, cycle_name_2, taxon) %>%
    summarise(rel_abund = sum(rel_abund), #add Others taxon together
              .groups = "drop") %>%
   left_join(class_df , by=c("taxon" = "order")) %>% # get class levels for order
   mutate(cycle_name_2 = fct_relevel(cycle_name_2, "SA3", "SA2", "SA1", "ST2", "ST1")) %>%
  # mutate(class = fct_reorder(class, rel_abund)) %>%
      # mutate(class = fct_relevel(class,
      #                          "Mamiellophyceae",
      #                          "Prymnesiophyceae",
      #                          "Pelagophyceae",
      #                          "Chrysophyceae",
      #                         # "Chloropicophyceae",
      #                          "NA")) %>%
   ggplot(aes(x=rel_abund, y=taxon, colour = cycle_name_2)) +
      geom_stripes(even = "#FFFFFFFF", odd = "#EFEFEFFF") +
     stat_summary(fun.data=median_hilow, geom = "pointrange", #75% confidence interval
                  fun.args = list(conf.int=0.75),
                  position=position_dodge(width=0.5),
                  size=0.25) +
    scale_colour_manual(values=colours_cycle, limits = force)+
    theme_classic()+
    xlab("Relative abundance (%)") +
    ylab(NULL) +
   theme_classic() +
    theme(#legend.position = c(0.9,0.15),
          #legend.background =element_blank(),
          panel.grid.major.x = element_line(colour="lightgrey", size=0.25),
          strip.placement = "outside",
          strip.text.y.left = element_text(angle = 0),
          strip.background = element_blank()) +
   xlim(0,50)+
    guides(colour=guide_legend(title="Cycle")) +
       facet_grid(class~., scales = "free_y", space = "free", switch = "y") 

p3.1
pdf("nano_order_median.pdf", height=7, width=8) ; plot(p3.1) ; dev.off()

#heatmap
p3 <- inner_join(taxon_rel_abund, taxon_pool, by="taxon") %>%
    mutate(taxon=if_else(pool, "Other", taxon)) %>% 
    group_by(cycle_name_2, file_code, taxon) %>%
    summarise(rel_abund = sum(rel_abund), #add Others taxon together
              .groups = "drop") %>%
    group_by(cycle_name_2, taxon) %>%
   summarise(rel_abund = median(rel_abund)) %>% # for median reads per taxon per cycle
   left_join(class_df , by=c("taxon" = "order")) %>% # get class levels for sp.
     ggplot(aes(x=cycle_name_2, y = taxon, fill =rel_abund)) +
    geom_tile() +
  # geom_text(aes(label=format(round(rel_abund,1), nsmall = 1)), size=3)+
  scale_fill_distiller(palette = "Purples",
                        name = "Median<br>Relative<br>Abund (%)",
                       expand = c(0,0),
      #  breaks = c(0, 0.1, 1, 10, 50),
        na.value = "white",
       direction = 1)+
    # scale_fill_gradient2(
    #     name = "Median<br>Relative<br>Abund (%)",
    #     low = "white",
    #     mid = "#C3D7A4",
    #     high = "#52854C",
    #     space = "Lab",
    # #    trans = "log",
    #     expand = c(0,0),
    #   #  breaks = c(0, 0.1, 1, 10, 50),
    #     na.value = "white",
    #     guide = "colourbar",
    #     aesthetics = "fill") +
  theme_classic()+
    theme(text=element_text(size=10),
              strip.text.y.left = element_text(angle = 0),
              panel.spacing = unit(0,"lines"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              strip.background = element_blank(),
          axis.line = element_blank(),
          axis.ticks = element_blank(),
          legend.title = element_markdown(),
          strip.placement = "outside") +
  scale_x_discrete(position = "top") +
  xlab(NULL)+
  ylab(NULL) +
    facet_grid(class~., scales = "free_y", space = "free", switch = "y") 
    
p3
pdf("nano order_heatmap.pdf", height=4, width=7) ; plot(p3) ; dev.off()
```

## Species - median  select 5% minimum
```{r}
taxon_rel_abund <- otu_rel_abund %>%
    filter(level == "species") %>% #choose level here 
    group_by(cycle_name_2, file_code, taxon) %>%
    summarise(rel_abund = 100*sum(rel_abund)) %>%
    pivot_wider(names_from = taxon, values_from = rel_abund) %>% #from here is to replace any file_code with NA as 0
    replace(., is.na(.), 0) %>%
    pivot_longer(cols= !cycle_name_2:file_code,
                 names_to = "taxon",
                 values_to = "rel_abund")

taxon_pool <- taxon_rel_abund %>%
    group_by(cycle_name_2, taxon) %>%
    summarise(median = mean(rel_abund)) %>% #calculate median proportion of reads per sp per cycle
    group_by(taxon) %>%
    summarise(pool = max(median) <= 5, #species with a max median proportion of reads in any cycle < 5% is under Others
              .groups = "drop")

#get class name for facet grid grouping with class
class_df <- long_nano_1 %>%
    select(class, order, species) %>%
    distinct()

#median range plot
 p4.1<- inner_join(taxon_rel_abund, taxon_pool, by="taxon") %>%
    mutate(taxon=if_else(pool, "Other", taxon)) %>% # replace dinophyceae XXX and haptophyta clade hap3 XXX
    group_by(file_code, cycle_name_2, taxon) %>%
    summarise(rel_abund = sum(rel_abund), #add Others taxon together
              .groups = "drop") %>%
   left_join(class_df , by=c("taxon" = "species")) %>% # get class levels for order
   mutate(cycle_name_2 = fct_relevel(cycle_name_2, "SA3", "SA2", "SA1", "ST2", "ST1")) %>%
   ggplot(aes(x=rel_abund, y=taxon, colour = cycle_name_2)) +
      geom_stripes(even = "#FFFFFFFF", odd = "#EFEFEFFF") +
     stat_summary(fun.data=median_hilow, geom = "pointrange", #75% confidence interval
                  fun.args = list(conf.int=0.75),
                  position=position_dodge(width=0.5),
                  size=0.25) +
    scale_colour_manual(values=colours_cycle, limits = force)+
    theme_classic()+
    xlab("Relative abundance (%)") +
    ylab(NULL) +
   theme_classic() +
    theme(#legend.position = c(0.9,0.15),
          #legend.background =element_blank(),
          panel.grid.major.x = element_line(colour="lightgrey", size=0.25),
          strip.placement = "outside",
          strip.text.y.left = element_text(angle = 0),
          strip.background = element_blank()) +
    guides(colour=guide_legend(title="Cycle")) +
       facet_grid(order~., scales = "free_y", space = "free", switch = "y") 

p4.1
pdf("nano_species_median.pdf", height=7, width=8) ; plot(p4.1) ; dev.off()

#heatmap
p4 <- inner_join(taxon_rel_abund, taxon_pool, by="taxon") %>%
    mutate(taxon=if_else(pool, "Other", taxon)) %>% 
    group_by(cycle_name_2, file_code, taxon) %>%
    summarise(rel_abund = sum(rel_abund), #add Others taxon together
              .groups = "drop") %>%
    group_by(cycle_name_2, taxon) %>%
   summarise(rel_abund = mean(rel_abund)) %>% # for median reads per taxon per cycle
   left_join(class_df , by=c("taxon" = "species")) %>% # get class levels for sp.
     ggplot(aes(x=cycle_name_2, y = taxon, fill =rel_abund)) +
    geom_tile() +
  # geom_text(aes(label=format(round(rel_abund,1), nsmall = 1)), size=3)+
  scale_fill_distiller(name = "Median<br>Relative<br>Abund (%)",
                       palette = "Purples",
                       expand = c(0,0),
                      na.value = "white",
                      direction = 1)+
    # scale_fill_gradient2(
    #     name = "Median<br>Relative<br>Abund (%)",
    #     low = "white",
    #     mid = "#C3D7A4",
    #     high = "#52854C",
    #     space = "Lab",
    # #    trans = "log",
    #     expand = c(0,0),
    #   #  breaks = c(0, 0.1, 1, 10, 50),
    #     na.value = "white",
    #     guide = "colourbar",
    #     aesthetics = "fill") +
  theme_classic()+
    theme(text=element_text(size=10),
              strip.text.y.left = element_text(angle = 0),
              panel.spacing.x = unit(0,"lines"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              strip.background = element_blank(),
          axis.line = element_blank(),
          axis.ticks = element_blank(),
          legend.title = element_markdown(),
          strip.placement = "outside") +
  scale_x_discrete(position = "top") +
  xlab(NULL)+
  ylab(NULL) +
    facet_grid(order~., scales = "free_y", space = "free", switch = "y") 
    
p4
pdf("nano_species_heatmap.pdf", height=7, width=8) ; plot(p4) ; dev.off()
```
```{r}
library(patchwork)
plot <- (p1 |p5) / (p2|p3) + plot_annotation(tag_levels = 'A') + plot_layout( heights = c(1, 1.75))

plot
pdf("test_heatmap.pdf", height=6, width=12) ; plot(plot) ; dev.off()

plot <- (p3 |p4) + plot_layout(guides = 'collect') +plot_annotation(tag_levels = 'A')

plot
pdf("pico_class species_heatmap.pdf", height=9, width=9) ; plot(plot) ; dev.off()
```

## Species - median  select 5% minimum
```{r}
taxon_rel_abund <- otu_rel_abund %>%
    filter(level == "species") %>% #choose level here 
    group_by(water_mass, cycle_name_2, file_code, taxon) %>%
    summarise(rel_abund = 100*sum(rel_abund)) %>%
    pivot_wider(names_from = taxon, values_from = rel_abund) %>% #from here is to replace any file_code with NA as 0
    replace(., is.na(.), 0) %>%
    pivot_longer(cols= !water_mass:file_code,
                 names_to = "taxon",
                 values_to = "rel_abund")

summary <- taxon_rel_abund %>%
  filter(taxon %in% c("Phaeocystis_antarctica")) %>%
  group_by(taxon, cycle_name_2) %>%
  summarise(median = median(rel_abund),
            mean = mean(rel_abund))
summary

taxon_pool <- taxon_rel_abund %>%
    group_by(cycle_name_2, taxon) %>%
    summarise(median = mean(rel_abund)) %>% #calculate median proportion of reads per sp per cycle
    group_by(taxon) %>%
    summarise(pool = max(median) <= 5, #species with a max median proportion of reads in any cycle < 5% is under Others
              .groups = "drop")

#get class name for facet grid grouping with class
class_df <- long_nano_1 %>%
    select(class, order, genus) %>%
    distinct()

#median range plot
 p4.1<- inner_join(taxon_rel_abund, taxon_pool, by="taxon") %>%
    mutate(taxon=if_else(pool, "Other", taxon)) %>% # replace dinophyceae XXX and haptophyta clade hap3 XXX
    group_by(file_code, cycle_name_2, taxon) %>%
    summarise(rel_abund = sum(rel_abund), #add Others taxon together
              .groups = "drop") %>%
   left_join(class_df , by=c("taxon" = "genus")) %>% # get class levels for order
   mutate(cycle_name_2 = fct_relevel(cycle_name_2, "SA3", "SA2", "SA1", "ST2", "ST1")) %>%
   ggplot(aes(x=rel_abund, y=taxon, colour = cycle_name_2)) +
      geom_stripes(even = "#FFFFFFFF", odd = "#EFEFEFFF") +
     stat_summary(fun.data=median_hilow, geom = "pointrange", #75% confidence interval
                  fun.args = list(conf.int=0.75),
                  position=position_dodge(width=0.5),
                  size=0.25) +
    scale_colour_manual(values=colours_cycle, limits = force)+
    theme_classic()+
    xlab("Relative abundance (%)") +
    ylab(NULL) +
   theme_classic() +
    theme(#legend.position = c(0.9,0.15),
          #legend.background =element_blank(),
          panel.grid.major.x = element_line(colour="lightgrey", size=0.25),
          strip.placement = "outside",
          strip.text.y.left = element_text(angle = 0),
          strip.background = element_blank()) +
    guides(colour=guide_legend(title="Cycle")) +
       facet_grid(order~., scales = "free_y", space = "free", switch = "y") 

p4.1
# pdf("nano_species_median.pdf", height=7, width=8) ; plot(p4.1) ; dev.off()

#heatmap
p4 <- inner_join(taxon_rel_abund, taxon_pool, by="taxon") %>%
    mutate(taxon=if_else(pool, "Other", taxon)) %>% 
    group_by(cycle_name_2, file_code, taxon) %>%
    summarise(rel_abund = sum(rel_abund), #add Others taxon together
              .groups = "drop") %>%
    group_by(cycle_name_2, taxon) %>%
   summarise(rel_abund = mean(rel_abund)) %>% # for median reads per taxon per cycle
   left_join(class_df , by=c("taxon" = "species")) %>% # get class levels for sp.
     ggplot(aes(x=cycle_name_2, y = taxon, fill =rel_abund)) +
    geom_tile() +
  # geom_text(aes(label=format(round(rel_abund,1), nsmall = 1)), size=3)+
  scale_fill_distiller(name = "Median<br>Relative<br>Abund (%)",
                       palette = "Purples",
                       expand = c(0,0),
                      na.value = "white",
                      direction = 1)+
    # scale_fill_gradient2(
    #     name = "Median<br>Relative<br>Abund (%)",
    #     low = "white",
    #     mid = "#C3D7A4",
    #     high = "#52854C",
    #     space = "Lab",
    # #    trans = "log",
    #     expand = c(0,0),
    #   #  breaks = c(0, 0.1, 1, 10, 50),
    #     na.value = "white",
    #     guide = "colourbar",
    #     aesthetics = "fill") +
  theme_classic()+
    theme(text=element_text(size=10),
              strip.text.y.left = element_text(angle = 0),
              panel.spacing.x = unit(0,"lines"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              strip.background = element_blank(),
          axis.line = element_blank(),
          axis.ticks = element_blank(),
          legend.title = element_markdown(),
          strip.placement = "outside") +
  scale_x_discrete(position = "top") +
  xlab(NULL)+
  ylab(NULL) +
    facet_grid(order~., scales = "free_y", space = "free", switch = "y") 
    
p4
pdf("nano_species_heatmap.pdf", height=7, width=8) ; plot(p4) ; dev.off()
```

# Sorted syn
## Read data
```{r}
source(here("phyloseq", "init_files", "read_petb_2.0.R"))

# #assign trophic mode and remove heterotrophics that are not dinophyceae
# ps_nano <- phyloseq_assign_trophic(ps_nano)%>%
#     subset_taxa(trophic_mode_2 != "heterotrophic")
long_sorted <- long_sorted %>%
  mutate(cycle_name = recode(Cycle., "1" = "SA1",
                                    "2" = "SA2",
                                    "3" = "ST1",
                                    "4" = "ST2",
                                    "5" = "SA3")) %>%
  mutate(cycle_name = fct_relevel(cycle_name, "ST1", "ST2", "SA1", "SA2", "SA3"))
```
## prepare data frame
```{r}

long_sorted_1 <- long_sorted %>%
   #  ps_top_asv(10) %>% #subset top 10 ASVs per sample
     filter(vial_type=="light")


# pivot longer dataframe
otu_rel_abund <- long_sorted_1 %>%
    group_by(file_code) %>%
    mutate(rel_abund = n_reads / sum(n_reads)) %>%
    ungroup() %>%
    select(-n_reads) %>%
    pivot_longer(cols = asv_code:Subclade,
                 names_to = "level",
                 values_to = "taxon") 
```

```{r}
taxon_rel_abund <- otu_rel_abund %>%
    filter(level == "Clade") %>% #choose level here 
    group_by(cycle_name, file_code, taxon) %>%
    summarise(rel_abund = 100*sum(rel_abund)) %>%
    pivot_wider(names_from = taxon, values_from = rel_abund) %>% #from here is to replace any file_code with NA as 0
    replace(., is.na(.), 0) %>%
    pivot_longer(cols= !cycle_name:file_code,
                 names_to = "taxon",
                 values_to = "rel_abund")

## summary for results
summary <- taxon_rel_abund %>%
  # filter(taxon %in% c("Mamiellophyceae", "Chrysophyceae","Pelagophyceae","Prymnesiophyceae")) %>%
  filter(cycle_name %in% c("SA2", "SA3")) %>%
  group_by(taxon) %>%
  summarise(median = median(rel_abund),
            mean = mean(rel_abund),
            max = mean(rel_abund),
            min = min(rel_abund))
summary
```


## subclade - select above 5%
```{r}
taxon_rel_abund <- otu_rel_abund %>%
    filter(level == "Subclade") %>% #choose level here 
    group_by(water_mass, cycle_name, file_code, taxon) %>%
    summarise(rel_abund = 100*sum(rel_abund)) %>%
    pivot_wider(names_from = taxon, values_from = rel_abund) %>% #from here is to replace any file_code with NA as 0
    replace(., is.na(.), 0) %>%
    pivot_longer(cols= !water_mass:file_code,
                 names_to = "taxon",
                 values_to = "rel_abund")

## summary for results
summary <- taxon_rel_abund %>%
  # filter(taxon %in% c("Mamiellophyceae", "Chrysophyceae","Pelagophyceae","Prymnesiophyceae")) %>%
  # filter(cycle_name_2 == "SA1") %>%
  group_by(taxon, water_mass) %>%
  summarise(median = median(rel_abund),
            mean = mean(rel_abund))
summary

taxon_pool <- taxon_rel_abund %>%
    group_by(cycle_name, taxon) %>%
    summarise(median = median(rel_abund)) %>% #calculate median proportion of reads per sp per cycle
    group_by(taxon) %>%
    summarise(pool = max(median) <= 5, #species with a max median proportion of reads in any cycle < 5% is under Others
              median = max(median),
              .groups = "drop")

#get class name for facet grid grouping with class
df <- long_sorted_1 %>%
    select(Clade, Subclade) %>%
    distinct()

#median range plot
 p5.1<- inner_join(taxon_rel_abund, taxon_pool, by="taxon") %>%
    mutate(taxon=if_else(pool, "Other", taxon)) %>% 
    group_by(file_code, cycle_name, taxon) %>%
    summarise(rel_abund = sum(rel_abund), #add Others taxon together
              .groups = "drop") %>%
   left_join(df , by=c("taxon" = "Subclade")) %>% # get class levels for order
   mutate(cycle_name = fct_relevel(cycle_name, "SA3", "SA2", "SA1", "ST2", "ST1")) %>%
   ggplot(aes(x=rel_abund, y=taxon, colour = cycle_name)) +
      geom_stripes(even = "#FFFFFFFF", odd = "#EFEFEFFF") +
     stat_summary(fun.data=median_hilow, geom = "pointrange", #75% confidence interval
                  fun.args = list(conf.int=0.75),
                  position=position_dodge(width=0.5),
                  size=0.25) +
    scale_colour_manual(values=colours_cycle, limits = force)+
    theme_classic()+
    xlab("Relative abundance (%)") +
    ylab(NULL) +
   theme_classic() +
    theme(#legend.position = c(0.9,0.15),
          #legend.background =element_blank(),
          panel.grid.major.x = element_line(colour="lightgrey", size=0.25),
          strip.placement = "outside",
          strip.text.y.left = element_text(angle = 0),
          strip.background = element_blank()) +
    guides(colour=guide_legend(title="Cycle"))  +
    facet_grid(Clade~., scales = "free_y", space = "free", switch = "y") 
p
#pdf("sorted_pico_species_median.pdf", height=7, width=8) ; plot(p) ; dev.off()

#heatmap
p5 <- inner_join(taxon_rel_abund, taxon_pool, by="taxon") %>%
   
    mutate(taxon=if_else(pool, "Other", taxon)) %>% 
    group_by(cycle_name, file_code, taxon) %>%
    summarise(rel_abund = sum(rel_abund), #add Others taxon together
              .groups = "drop") %>%
    group_by(cycle_name, taxon) %>%
   summarise(rel_abund = median(rel_abund)) %>% # for median reads per taxon per cycle
   left_join(df , by=c("taxon" = "Subclade")) %>% # get class levels for sp.
 
     ggplot(aes(x=cycle_name, y = taxon, fill =rel_abund)) +
    geom_tile() +
  scale_fill_distiller(palette = "Oranges",
                        name = "Median<br>Relative<br>Abund (%)",
                       expand = c(0,0),
        na.value = "white",
       direction = 1)+
  theme_classic()+
    theme(text=element_text(size=10),
              strip.text.y.left = element_text(angle = 0),
              panel.spacing = unit(0,"lines"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              strip.background = element_blank(),
          axis.line = element_blank(),
          axis.ticks = element_blank(),
          legend.title = element_markdown(),
          strip.placement = "outside") +
  scale_x_discrete(position = "top") +
  xlab(NULL)+
  ylab(NULL) +
    facet_grid(Clade~., scales = "free_y", space = "free", switch = "y") 
    
p
# pdf("syn_heatmap.pdf", height=2, width=5) ; plot(p) ; dev.off()
```
# Descriptive stats
```{r}
#no. of samples per cycle
sample_no <- taxon_rel_abund %>%
  select(cycle_name_2, file_code) %>%
  distinct() %>%
  group_by(cycle_name_2) %>%
  summarise(n = n())
sample_no
```

## Order heatmap -ignore
```{r, eval=FALSE}
#group by the level you want to plot, calculate percent abundance
taxon_rel_abund <- otu_rel_abund %>%
    filter(level == "order") %>% #choose level here 
        mutate(taxon = case_when(taxon == 'Cryptomonadales' ~ 'Cryptomonadales',
                                 taxon == 'Dinophyceae_X' ~ 'Dinophyceae_X',
                                 #taxon == 'Suessiales' ~ 'Suessiales',
                                 taxon == 'Gymnodiniales' ~ 'Gymnodiniales',
                                 taxon == 'Bolidophyceae' ~ 'Bolidophyceae',
                                 taxon == 'Peridiniales' ~'Peridiniales',
                                 taxon == 'Prorocentrales' ~ 'Prorocentrales',
                                 taxon == 'Isochrysidales' ~'Isochrysidales',
                                 taxon == 'Prymnesiales' ~'Prymnesiales',
                                 taxon == 'Phaeocystales' ~'Phaeocystales',
                                 taxon == 'Bacillariophyta_X' ~'Bacillariophyta_X',
                                 #taxon == 'Chlorellales' ~'Chlorellales',
                                 TRUE~ 'Other')) %>%
     group_by(cycle_name_2, file_code, taxon) %>%
    summarise(rel_abund = 100*sum(rel_abund), .groups = "drop") %>%
    pivot_wider(names_from = taxon, values_from = rel_abund) %>% #from here is to replace any file_code with NA as 0
    replace(., is.na(.), 0) %>%
    pivot_longer(cols=Bacillariophyta_X:Prymnesiales,
                 names_to = "taxon",
                 values_to = "rel_abund") %>%
  group_by(cycle_name_2, taxon) %>%
  summarise(rel_abund = median(rel_abund)) # for median reads per taxon per cycle

#get class name for facet grid grouping with class
class_df <- long_nano_1 %>%
    select(class, order) %>%
    distinct() %>%
    filter(order %in% c("Cryptomonadales",
                        "Dinophyceae_X",
                        #"Suessiales",
                        "Gymnodiniales",
                        "Bolidophyceae",
                        "Peridiniales",
                        "Prorocentrales",
                        "Isochrysidales",
                        "Prymnesiales",
                        "Phaeocystales",
                        "Bacillariophyta_X"
                      #  "Chlorellales"
                        ))
library(ggtext)
p <- taxon_rel_abund %>%
    full_join(class_df , by=c("taxon" = "order")) %>%
     mutate(taxon = fct_reorder(taxon, rel_abund, .desc = TRUE)) %>%
     ggplot(aes(x=cycle_name_2, y = taxon, fill =rel_abund)) +
    geom_tile() +
  #geom_text(aes(label=format(round(rel_abund,1), nsmall = 1)), size=5)+
    scale_fill_gradient2(
        name = "Median<br>Relative<br>Abund (%)",
        low = "white",
        mid = "#C3D7A4",
        high = "#52854C",
        space = "Lab",
       # trans = "log",
        expand = c(0,0),
        na.value = "white",
        guide = "colourbar",
        aesthetics = "fill",
        limits = c(0, NA)) +
  theme_classic()+
    theme(text=element_text(size=10),
              strip.text.y.left = element_text(angle = 0),
              panel.spacing.x = unit(0,"lines"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              strip.background = element_blank(),
          axis.line = element_blank(),
          axis.ticks = element_blank(),
          legend.title = element_markdown(),
          strip.placement = "outside") +
  scale_x_discrete(position = "top") +
  xlab(NULL)+
  ylab(NULL) +
    facet_grid(class~., scales = "free_y", space = "free", switch = "y") 
    
p
```

## Check the genus of Bacillariophyta -ignore
```{r}
check <- long_nano_1 %>%
  filter(order ==  "Bacillariophyta_X")

p <-   check %>%
  ggplot(aes(x=n_reads, y=fct_rev(file_code), color=genus, fill=genus)) + 
        geom_bar(position="fill", stat="identity") + 
        theme(panel.spacing.y = unit(0,"lines"),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              strip.background = element_blank(),
              panel.border = element_rect(colour = "black", fill = NA),
              strip.text.y.right = element_text(angle = 0))+
       facet_grid(cycle_name_2~., scales = "free_y", space = "free")+
       # scale_color_manual(values=colours_petb_subclade, limits = force) + 
      #  scale_fill_manual(values=colours_petb_subclade, limits = force) +
        xlab("Percent of reads")+
        ylab("Sample number")
p

```

